### 2020-11-13 why-negative-selection.Rmd
p.value to p in graphs

---
title: "Economic Theory of Fertility Explains Modern Natural Selection"
author: "David Hugh-Jones, Abdel Abdellaoui"
date: "`r Sys.Date()`"
abstract: "Natural selection has been documented in contemporary humans, but little is known about the mechanisms behind it. We test for natural selection on 33 polygenic scores over two generations, using data from UK Biobank. Selection effects appear consistent over time. Polygenic scores which predict lower earnings and education are selected for. Effects are larger among low-income and less educated people, younger parents, and people not living with a partner. Overall effect sizes are substantially bigger when we correct for ascertainment bias. The direction of natural selection is reversed among older parents, or controlling for age at first live birth. These patterns can be explained by economic theories of fertility, in which higher earnings may either increase or decrease fertility."
bibliography: "negative-selection.bib"
output: 
  bookdown::pdf_document2:
    toc: false
    latex_engine: xelatex
    number_sections: false
header-includes:
  - \usepackage{subfig}
  - \usepackage{setspace}\doublespacing
  - \usepackage{placeins}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}

library(drake)
library(magrittr)
library(dplyr)
library(forcats)
library(ggplot2)
library(tidyr)
library(santoku) # keep this after tidyr as it masks chop
library(purrr)
library(glue)
library(broom)
library(huxtable)

drake::loadd(rgs)
drake::loadd(score_names)
drake::loadd(famhist)

options(digits = 2)
options(dplyr.summarise.inform = FALSE) # shut up about summarize
knitr::opts_chunk$set(
        echo = FALSE, 
        warning = FALSE, 
        cache = FALSE, 
        error = TRUE,
        fig.height = 3.5
      )

huxtable::set_default_properties(latex_float = "h!")

theme_set(theme_minimal())
my_hline <- geom_hline(yintercept = 0, colour = "grey20", linetype = "dotted")
my_vline <- geom_vline(xintercept = 0, colour = "grey20", linetype = "dotted")
pval_palette <- c("FALSE" = "grey65", "TRUE" = "#FAC32D")

standard_ggplot <- function (
        dfr, 
        fill_col, 
        n_regs, 
        ...,
        score_col = quo(term), 
        order_idx = 1, 
        n_cats    = NULL
      ) {
  if (! missing(score_col)) score_col <- enquo(score_col)
  mfc <- missing(fill_col)
  fill_col <- if (mfc) quo(NULL) else enquo(fill_col)
  if (missing(n_cats)) {
    n_cats <- if (mfc) 1 else length(unique(pull(dfr, {{fill_col}})))
  }

  n_regs <- as.double(n_regs)

  dfr %>% mutate(
          !! score_col := pretty_names(!! score_col),
          !! score_col := fct_reorder(!! score_col, estimate, order_abs(order_idx))
        ) %>% 
        rename(p = p.value) %>% 
        ggplot(aes(estimate, {{score_col}}, fill = {{fill_col}}, ...,
          colour = p < 0.05/{{n_regs}})) +
          pval_geom_point() +
          my_vline + 
          pval_scale() +
          my_fill_scale(n = n_cats) +
          my_labs() +
          my_theme()
}

my_theme <- function (...) theme(axis.text.y = element_text(size = 7), ...)

pval_scale <- function (...) scale_color_manual(values = pval_palette, ...)

pval_geom_point <- function (size = 1.5, stroke = 0.5, ...) {
  geom_point(shape = 21, size = size, stroke = stroke, ...)
}

my_fill_scale <- function (n, aesthetics = "fill", ...) {
  if (n == 2) {
    scale_colour_manual(values = c("steelblue4", "#D63C3C"), 
          aesthetics = aesthetics, ...)
    # scale_color_brewer(type = "qual", palette = 3, aesthetics = aesthetics, ...)
  } else {
    scale_color_brewer(aesthetics = aesthetics, ...)
  }
}

my_labs <- function (x = "Effect size", y = "", ...) labs(x = x, y = y, ...)

# for use when reordering factors. Order by the nth score in a group
order_abs <- function (n = 1) {
  function (x) x[n] 
}

pretty_names <- function (names) {
  pretty <- c(
    ADHD_2017               = "ADHD",
    age_at_menarche         = "Age at menarche",
    age_at_menopauze        = "Age at menopause",
    agreeableness           = "Agreeableness",
    ai_substance_use        = "Age at smoking initiation",
    alcohol_schumann        = "Alcohol use",
    alzheimer               = "Alzheimer",
    autism_2017             = "Autism",
    bipolar                 = "Bipolar",
    bmi_combined            = "BMI",
    body_fat                = "Body Fat",
    caffeine                = "Caffeine",
    cannabis                = "Cannabis (ever vs. never)",
    cognitive_ability       = "Cognitive Ability",
    conscientiousness       = "Conscientiousness",
    coronary_artery_disease = "Coronary Artery Disease",
    cpd_substance_use       = "Cigarettes per day",
    diagram_T2D             = "Type 2 Diabetes",
    dpw_substance_use       = "Drinks per week",
    EA2_noUKB               = "Educ. attainment 2 (no UK)",
    EA3_excl_23andMe_UK     = "Educ. attainment 3 (no UK)",
    eating_disorder         = "Eating disorder",
    extraversion            = "Extraversion",
    height_combined         = "Height",
    hip_combined            = "Hip circumference",
    MDD_PGC2_noUKB          = "Major Depressive Disorder",
    neuroticism             = "Neuroticism",
    openness                = "Openness",
    sc_substance_use        = "Smoking cessation",
    SCZ2                    = "Schizophrenia",
    si_substance_use        = "Smoking initiation",
    wc_combined             = "Waist circumference",
    whr_combined            = "Waist-hip ratio"
  )

  pretty[names]
}

n_in_regs <- function (var, dv = "n_children") {
  # all PGS have the same NA pattern, so we just use one
  tbl <- table(famhist[
                  ! is.na(famhist[dv]) & ! is.na(famhist$whr_combined), 
                  var
                ])
  c(tbl)
}

add_n <- function (var, famhist_var = var, dv = "n_children") {
  var <- as.factor(var)
  n <- n_in_regs(famhist_var, dv)
  levels(var) <- sprintf("%s (N = %s)", levels(var), 
                         prettyNum(n[seq_len(nlevels(var))], big.mark = ","))
  # var_n <- sprintf("%s (N = %s)", var, prettyNum(n[var], big.mark = ","))
  # var_n[is.na(var)] <- NA
  var
}

# from https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R
# for ordering within facets
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}

```

\normalem

```{r NOTES-TODO, eval=FALSE}
"
# Open questions/TODO

* If we weight with the GHS, and plot the means of figure 1 using this w
  weighting, most things stay the same - but the change in EA is reversed! EA
  increases over time. Could this be because the sample includes many
  relatively educated older people? If so, then mean effects are *overestimated*
  by figure 1.
* Reweight other regressions (e.g. the omnibus regression in the appendix)?
* Look at geography, esp. in siblings regressions.
* We say we don't see change over time. But, we do see the effect of
  controlling for AFLB changing over generations.
* Beauchamp 2016 used the HRS. He reckons it's representative. Is it? How do
  his effect sizes compare?
  - he also worries about the quality/quantity tradeoff, which might mean that
    people with fewer children have more grandchildren (than you'd expect).
    As we have the two generations, we can test this. (Actually, there's a
    positive correlation of about 0.1)
* Another point is that negative selection only among poor will increase
    genetic inequality.
* When we have better definition of "partner", bring it across from the other
  project to define with_partner
* If you take "year of last job started" and see if it is after/before
  the first child, you can figure out whether mothers keep working.
  - better still, look at all jobs started and see if any are within e.g. 5 years of
  first child's birth
  - this could then be another way to get at income vs substitution effects. You'd
  predict negative selection effects among mothers who worked. Of course the
  choice to work is endogenous.
* Put mediating effects of pay and education in there (see explore-sex-vs-pay.R)
  - in a multivariate regression, most effects vanish once we control for these, including EA3

## LITERATURE NOTES

* Lundberg and Pollak JEP 2007:
    - several models where nonmarital fertility is more likely for poorer women
    - Burdett and Ermisch 2002, Willis 1999, Neal 2004
    - Unmarried parents may not be able to credibly commit to each other

* Happel et al. 1984
   - childbirth is chosen to minimize loss of earnings from interrupted work.
   - early childbirth if your skills decay a lot  while childrearing. (No point
     gaining them then losing them.) Late childbirth otherwise.
   - late childbirth if you have acquired skills before marriage.
   - Capital-market imperfections make it wise to delay childbirth (till man has
     high income to cope while wife is at home). More so for men's wages which
     rise faster.
   - empirically: (among married couples) 
     - high-skill occupation women have children later. Higher-earning men have 
     them earlier.

- Caucutt et al. 2001
  - high waged and high educated women (of recent cohorts) more likely to
    delay childbirth

- Cohen et al. 2013
  - analyse effect of Israeli subsidy
  - along the way, find that income affects fertility negatively at low income,
    positively at high income, supporting Becker and Tomes 76

- Becker and Tomes 1976
  - investment in children quality and quantity when quality must be equal
    across children; and some quality comes from an innate endowment.
  - "observed" quantity-of-children income elasticity rises as income rises
  - while "observed" quality-of-children elasticity falls as income rises, 
    b/c more of total quality must be paid for by parents themselves. 
  - the need to pay more for quality results in lower quantity, since total cost is 
    quantity X quality
  - something fishy about argument somehow...

- Lindo 2008
  - men who lose a job have children sooner, but fewer children in total
  - controls for all skills effects
  - so overall effect of income is positive, but timing changes

- Monstad et al. 2008
  - using an education reform to vary education
  - no evidence of a causal effect on fertility, but births start later
  * maybe the usual correlation reflects underlying skills. Your data could
    test this.

- Autor et al. 2017
  - the China shock (exogenous shock to income) 
  - they use Chinese import penetration by industry (instrumented using
    other countries) times industry employment shares; and separate it for
    males and females
  - shocks to male employment decreased fertility and marriage; 
    increased unmarried mothers
  - shocks to female employment *increase* marriage and motherhood
  * this roughly seems to be the consensus
  * Whereas, you find that male and female *skills* (as 'instrumented' by PGS)
    have the same effect on fertility - i.e. less skills = higher fertility

- Baudin et al. 2015
  * look at childlessness vs. fertility (extensive/intensive margin)
  * some people are too poor to afford children
  * 'a husband alleviates the time cost of having children'. But for some groups (see appendix)
    there is a +ve income/fertility gradient
  * fertility decreases with female educn due to subst effect
  * baby boom and bust in US: a decline in childlessness caused by poverty, then an increase
    caused by substn effect ('opportunity cost')
  * fertility decreases with educn both for single mothers and married women
    - slightly (25%) higher decrease per year of educn (Table 2) for single mothers

# WHAT MODEL FITS YOUR DATA
## ASSUMPTIONS
* Genetics produce human capital h. 
* h and education are compliments in producing earnings
* There are income and substitution effects on fertility:
  - income effects are stronger at high income levels (Becker and Tomes 1976)
    and among couples
* People seek partners when in the labour market. Women match a person, observe
  their quality, and decide whether to have a first child.
* Then they (may) have more children, either with the partner or not 
  (high quality partners stay around more)
* High quality partners increase household income

## RESULTS
* People with low h are more likely to leave education early and enter the labour market
  - because e and h are complements
  - this is part of the substitution effect, as they expect to spend some time
    raising children rather than being in work
* At low values of h, increases in h reduce fertility (substitution effect dominates)
  at higher values, smaller reductions or increases (income effect stronger)
  - because h predicts income
* In 1-parent households, increases in h reduce fertility (sub. effect), in 
  2-parent households, smaller reductions or increases (income effect stronger)
  - because h predicts income and because stronger income effect among couples
* People with low h start having children earlier (b/c leave educ earlier)
  - so part of the substitution effect comes via this
  - controlling for this the income effect dominates
  - hence increases in h reduce fertility less once we control for AFLB

## THOUGHTS
* Should we distinguish between men and women in couples?
  - within couples, male h should increase fertility (via earnings)
  - female h should decrease it (ditto?)
  - see explore-sex-vs-pay.R
"
```

Geneticists have long hypothesized that natural selection is taking
place in modern human populations. Recent work confirms this using
genome-wide analysis [@Barban_2016; @beauchamp2016genetic;
@kong2017selection; @sanjak2018evidence]. In particular, genetic
variants associated with higher educational attainment are being
selected against, although effect sizes appear small.

As yet we know little about the *mechanisms* behind natural selection in
contemporary populations. This paper uses data from UK Biobank to learn
more. We test for natural selection on 33 different polygenic scores. We
extend the analysis over two generations, using data on respondents'
number of siblings as well as their number of children to reconstruct
two generations of family size. This is interesting because consistent
natural selection over multiple generations could lead to substantive
effects in the long run. Most importantly, we examine the strength of
natural selection in different subgroups of the population, in order to
uncover patterns that can help illuminate the mechanisms involved.

We use data from UK Biobank. We compute 33 polygenic scores, normalized
to mean 0, variance 1. Figure \@ref(fig:plot-means-over-time) plots mean
polygenic scores by 5-year birth intervals. Several scores show
consistent increases or declines over this 30-year period, of the order
of 5% of a standard deviation.

```{r plot-means-over-time, fig.height = 5.5, fig.cap = "Mean polygenic scores by birth year in UK Biobank. Points are means for 5-year intervals. Lines are 95% confidence intervals. Green lines show a significant linear increase over time (p < 0.05/33). Red lines show a significant decrease."}

drake::loadd(pgs_over_time)

pgs_over_time %>% 
      mutate(score_name = pretty_names(score_name)) %>% 
      ggplot(aes(YOB, group = score_name, colour = Change)) + 
      geom_pointrange(aes(ymin = score_lo, y = score, ymax = score_hi), 
        fatten = 1.25) +
      facet_wrap(~score_name, ncol = 5) +
      my_hline +
      scale_color_manual(values = c("-" = "red", "+" = "darkgreen", "o" = "black")) +
      theme(
        strip.text  = element_text(size = 6), 
        axis.text.x = element_text(size = 7),
        legend.position = "none"
      ) +
      labs(y = "PGS", x = "Birth year (5 year intervals)")

```

These changes could reflect either natural selection within the UK
population, or ascertainment bias within the sample. Respondents are
higher income and better educated than the UK population, and they may
also differ on other unobserved characteristics [@10.1093/aje/kwx246].
Since richer and educated people also live longer, this bias might also
increase with age.

We next regress respondents'polygenic scores on their number of children
($y_i$):

```{=tex}
\begin{equation}
y_i = \alpha + \beta\mathrm{PGS}_i + \varepsilon_i
\label{eq:regression}
\end{equation}
```

```{r calc-sex-ratio}
female_ratio <- mean(famhist$sex == 0)
```

The "selection effect", $\beta$, measures the strength of natural
selection within the sample. In fact, since polygenic scores are
normalized to mean 0, $\beta$ is the expected polygenic score among
children of the sample [@beauchamp2016genetic]. To learn more about the
underlying mechanisms, we split the sample, starting with basic
demographic variables including education, income and sex. These are all
potential sources of ascertainment bias: as well as the ascertainment
for income and education, mentioned above, the sample sex ratio skews
`r female_ratio * 100`% female.

Figure \@ref(fig:plot-income-educ-level) plots selection effects for
each polygenic score, grouping respondents by age of completing
full-time education, and by household income. Effects are larger and
more significant for the lowest income category, and for the lowest
education category. Note that the overall effect is not a simple average
of the effect among the different subgroups, because polygenic scores
may also shift respondents between the subgroups. For example, a high
PGS for educational attainment may correlate with fewer children among
early school leavers, but may also increase the age at which a
respondent leaves school.

```{r plot-income-educ-level, fig.cap = "Selection effects by education and income", fig.subcap = c("Age left full-time education", "Household income"), fig.ncol = 1}

drake::loadd(res_edu)
n_regs <- as.double(nrow(res_edu))

res_edu %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs(1)),
        "Age left FTE" = fct_relevel(age_fte_cat, "< 16", "16-18", "> 18"),
        "Age left FTE" = add_n(`Age left FTE`, "age_fte_cat")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, colour = p < 0.05/{{n_regs}}, 
        fill = `Age left FTE`)) +
        my_vline +
        pval_geom_point() + 
        pval_scale() +
        my_fill_scale(n = 3, direction = -1) +
        my_labs() +
        my_theme()

drake::loadd(res_income)

n_regs <- as.double(nrow(res_income))
res_income %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs(1)),
        Income = factor(income_cat, 
          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", "> £100K")
        ),
        Income = add_n(Income, "income_cat")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, fill = Income, 
        colour = p < 0.05/{{n_regs}})) +
        my_vline +
        pval_geom_point(alpha = 0.8) + 
        my_fill_scale(n = 5, direction = -1) +
        pval_scale() +
        my_labs() +
        my_theme()     
```

```{r calc-income-edu-controlled}
drake::loadd(res_income_controlled, res_edu_controlled)
n_regs_ic <- nrow(res_income_controlled)
n_sig_ic <- sum(res_income_controlled$p.value < 0.05/n_regs_ic)

n_regs_educ <- nrow(res_edu_controlled)
n_sig_educ <- sum(res_edu_controlled$p.value < 0.05/n_regs_educ)
```

These results could be driven by age, if older respondents are poorer
and less educated, and also more subject to selection on polygenic
scores. However, if we rerun the regressions, interacting the polygenic
score with income category and also with a quadratic in age, the
interaction with income remains significant at 0.05/`r n_regs_ic` for
`r n_sig_ic` out of `r n_regs_ic` regressions. Similarly if we interact
the PGS with age of leaving full time education and a quadratic in age,
the interaction with age leaving FTE remains significant at
0.05/`r n_regs_educ` for `r n_sig_educ` out of `r n_regs_educ`
regressions.

Selection effects are also different between men and women (Appendix
Figure \@ref(fig:plot-sexes)). Differences are particularly large for
educational attainment, height and MDD. Several PGS for mental illness
and personality traits are more selected for (or less against) among
women, including major depressive disorder (MDD), schizophrenia and
neuroticism, while extraversion is more selected for among men. Waist
circumference and waist-hip ratio are less selected for among women, and
educational attainment is more selected against, though the difference
is only significant for EA3.

```{r calc-age-flb-cross}

drake::loadd(res_age_flb_cross)

cor_age_flb <- res_age_flb_cross %>% 
      mutate(category = gsub(":.*", "", term)) %>% 
      pivot_wider(score_name, names_from = category, values_from = estimate) %>%              {cor(.$`age_flb_cat10-22`, .$`age_flb_cat28-52`)}

```

We next focus on variables related to mate search and household
formation. We split males and females by lifetime number of sexual
partners, at the median value of 3 (Figure \@ref(fig:plot-n-partners)a).
For both sexes, selection effects are larger and more significant among
those with more than 3 partners. Results are unchanged if we control for
respondents' age (appendix Figure
\@ref(fig:plot-n-partners-controlled)). Splitting respondents by whether
they were living with a spouse or partner at the time of interview,
effects are larger among those not living with a spouse or partner
(Figure \@ref(fig:plot-n-partners)b). Lastly, we split female
respondents by age at first live birth (AFLB).[^1] There is evidence for
genetic effects on AFLB [@Barban_2016], and there is a close link
between this variable and number of children born. Figure
\@ref(fig:plot-n-partners)c shows effect sizes estimated separately for
each tercile of AFLB. Several effects are strikingly different across
terciles. ADHD and MDD are selected for amongst the youngest third of
mothers, but selected against among the oldest two-thirds. Educational
attainment is selected for among the oldest two-thirds of mothers, but
is not significantly selected among the youngest third. Similarly,
several PGS of body measurements are selected against only among older
mothers. The correlation between effect sizes for the youngest and
oldest terciles is `r cor_age_flb`. As before, recall that PGS may have
two effects: changing the number of children within categories, and
shifting respondents between categories.

[^1]: This information is unavailable for men.

```{r plot-n-partners, fig.cap = "Selection effects by number of sexual partners, household type, and age at first live birth", fig.subcap = c("Number of partners", "Household type", "Age at first live birth"), fig.ncol = 1, out.height = c("35%", "27%", "27%"), fig.align = "center"}

drake::loadd(res_partners)
res_partners %<>%       
      filter(grepl(":", term)) %>% 
      mutate(`N partners` = ifelse(grepl("TRUE:", term), "3 or less", "4 or more"))

n_regs <- as.double(nrow(res_partners))
res_partners %>% 
      mutate(
        score_name = pretty_names(score_name)
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(
        y      = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        x      = estimate, 
        fill   = `N partners`, 
        colour = p < 0.05/{{n_regs}}
      )) + 
        my_vline +
        pval_geom_point() +
        facet_wrap(vars(sex), scales = "free_y") +
        my_fill_scale(n = 2) +
        pval_scale() +
        my_theme(legend.position = "bottom") +
        my_labs(x = "", y = "") + 
        scale_y_reordered()

drake::loadd(res_with_partner)
res_with_partner %<>% filter(grepl(":", term))

n_regs <- nrow(res_with_partner)
hhtype_n <- n_in_regs("with_partner")

res_with_partner %>% 
      mutate(
        Household = ifelse(grepl("TRUE", term), "With partner", 
                      "Without partner"),
        with_p    = as.character(grepl("TRUE", term)),
        Household = sprintf("%s (N = %s)", Household, 
                      prettyNum(hhtype_n[with_p], big.mark = ","))
      ) %>% 
      standard_ggplot(fill = Household, n_regs = n_regs, score_col = score_name)

drake::loadd(res_age_flb_cross)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_cross))

res_age_flb_cross %>% 
      mutate(
        score_name = pretty_names(score_name),
        score_name = fct_reorder(score_name, estimate, order_abs(1)),
        term       = gsub("age_flb_cat(.*):.*", "\\1", term),
        term       = add_n(term, "age_flb_cat")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, score_name, fill = term, 
        colour = p < 0.05/{{n_regs}})) +
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 3, direction = -1, 
          guide = guide_legend(title = "Age at first live birth")) +
        my_labs() +
        my_theme()
```

```{r calc-corr-age-flb}

drake::loadd(res_age_flb)
drake::loadd(res_sex)

res_sex_comparison <- res_sex %>% 
      filter(sex == "Female") %>% 
      mutate(score_name = term) %>% 
      select(score_name, term, estimate:conf.high)

res_age_flb %<>% filter(term != "age_flb")

res_combined <- bind_rows(
        raw      = res_sex_comparison, 
        with_age = res_age_flb %>% select(score_name, term, estimate:conf.high), 
        .id      = "reg.type"
      ) %>% 
      arrange(score_name)

raw_flb_comparison <- res_combined %>% 
      pivot_wider(
        id_cols     = score_name, 
        names_from  = reg.type, 
        values_from = estimate
      )

cor_raw_flb <- cor(raw_flb_comparison$raw, raw_flb_comparison$with_age)
opp_signed <- sum(
        sign(raw_flb_comparison$raw) != sign(raw_flb_comparison$with_age)
      )

```

To investigate this further, we regress number of children on PGS
*controlling* for AFLB (appendix Figure \@ref(fig:plot-age-flb)). In
`r opp_signed` out of `r length(score_names)` cases, effects
change sign when controls are added. The correlation between effect
sizes controlling for AFLB, and raw effect sizes, is `r cor_raw_flb`.
Selection effects appear to come through two opposing channels: an
effect on AFLB, and an opposite-signed effect on number of children 
controlling for AFLB.

```{r calc-res-weighted}

drake::loadd(res_wt_flb_weights)
drake::loadd(res_wt_msoa_weights)
drake::loadd(res_wt_age_qual_weights)
drake::loadd(res_sibs_parent_weights)
drake::loadd(res_all)
drake::loadd(res_sex)

res_wt_combined <- bind_rows(
  None                     = res_all %>% 
                               filter(reg.type == "raw", dep.var == "N children") %>% 
                               select(-reg.type, -dep.var, -score_name),
  Geographical             = res_wt_msoa_weights,
  `Age/Qualification`      = res_wt_age_qual_weights,
  `Age/Qual/AFLB`  = res_wt_flb_weights,
  .id = "Weights"
)

res_wt_cb_wide <- res_wt_combined %>% 
                    tidyr::pivot_wider(
                      id_cols     = term,
                      names_from  = Weights,
                      values_from = estimate
                    ) %>% 
                    rowwise() %>% 
                    mutate(
                      consistent = all(c_across(-term) > 0) ||
                        all(c_across(-term) < 0)
                    ) %>% 
                    ungroup() %>% 
                    left_join(
                      res_wt_combined %>% 
                        filter(Weights == "None") %>% 
                        select(term, p.value),
                      by = "term"
                    ) %>% 
                    left_join(
                      res_sex %>% 
                        filter(sex == "Female") %>% 
                        select(term, estimate_females = estimate),
                      by = "term"
                    )

rel_sizes <- res_wt_cb_wide %>% 
                filter(
                  consistent, 
                  p.value < 0.05/length(score_names)
                ) %>% 
                mutate(
                  Geographical = Geographical/None,
                  `Age/Qualification` = `Age/Qualification`/None,
                  `Age/Qual/AFLB` = `Age/Qual/AFLB`/estimate_females
                ) %>% 
                select(-consistent, -None, -p.value, -estimate_females) %>% 
                mutate(
                  term = pretty_names(term)
                )

rel_sizes_summary <- rel_sizes %>% 
                summarize(
                  across(-term, ~ c(mean(.x), median(.x)))
                ) %>% 
                mutate(
                  term = c("Mean", "Median")
                )

mean_rel <- c(rel_sizes_summary[1, 1:3])
```

Since UK Biobank subjects are not representative of the wider population, our
results suggest that naïve estimates of natural selection are at risk of
ascertainment bias. To correct for this, we weight participants using data from
the 2011 Census and from the 2006 General Household Survey (GHS). We try three
alternative weighting schemes: geographical weighting by census MSOA, age and
presence/absence of a partner; weighting by age and highest educational
qualification; and age, highest qualification, and age at first live birth, for
women only. Mean effect sizes across all polygenic scores are increased by a
factor of `r mean_rel$Geographical` (geographical weighting), 
`r mean_rel[["Age/Qualification"]]` (age/qualification) or 
`r mean_rel[["Age/Qual/AFLB"]]` (age/qualification/AFLB). Estimates might be
further affected by weighting on other demographic variables. Since the Biobank
sample seems to be ascertained in ways that shrink estimates of selection
effects, we suspect that our weighted estimates are still conservative.

The UK Biobank data contains information on respondents' number of
siblings. Since respondents' polygenic scores are equal in expectation
to the mean scores of their parents, we can use this to look at selection 
effects in the parents' generation. We estimate equation (\ref{eq:regression})
using *number of siblings* (including the respondent) as the dependent variable.
The parents' generation has a further source of ascertainment bias: sampling
parents of respondents overweights parents who have many children. For instance,
parents of three children will have, on average, three times more children
represented in UK Biobank than parents of one child. Parents of no children will
by definition not be represented. To compensate, we reweight our preferred
weightings (Age/Qualification) by the inverse of *number of siblings*.

Selection effects are highly correlated across the two generations, and most
share the same sign (appendix Figure \@ref(fig:plot-res-sibs)). Effect size
estimates are larger for the parents' generation. We treat this result cautiously,
because when we split respondents up by year of birth, we find few differences
in effect sizes between early- and late-born respondents, for either generation.
In other words, since estimated effect sizes change when we change the dependent
variable, but do not change over time within either dependent variable, this
may be due to remaining ascertainment bias within the sample.

Although the direction of selection effects does not change between the generations,
there are other differences. After we residualize polygenic scores on the first
100 principal components of the array data, selection effects on *number of
siblings* are reduced by about ten percent on average, whereas effects on
*number of children* barely change (appendix Figure
\@ref(fig:plot-regs-controlled-pcs)). This could be because
earlier fertility is driven more by geographically clustered deprivation [e.g.
via an insurance motive, @rendall1998old], which may correlate with the
broad-scale genetic variation captured by principal components. Another
difference is that in the parents' generation, raw selection effects and effects
controlling for AFLB are positively correlated, whereas they are negatively
correlated in the respondents' generation.[^educ-exception] These differences
suggest that the mechanisms of natural selection may have changed over time,
even though the outcomes are broadly the same.

[^educ-exception]: Educational attainment is an exception. It is selected 
against in the parents' generation, but selected for if we control for AFLB,
which is the same pattern as for respondents.

Our empirical results can be explained by the economic
theory of fertility [@becker1960economic; @willis1973new; @becker1976child].
According to this, increases in a person's wage affect their fertility via two
opposing channels. There is an *income effect* by which children become more
affordable, like any other good. There is also a *substitution effect*: since
childrearing has a cost in time, the opportunity cost of childrearing increases
if one's market wage is higher. The income effect would lead higher earners to
have more children. The substitution effect would lead higher earners to have
fewer children. If so, then genetic variants which affect earnings potential in
the labour market may cause opposing effects on fertility. The income effect
will cause natural selection in favour of earnings-increasing variants. The
substitution effect will do the reverse.

```{r calc-res-earn-educ}

drake::loadd(res_ee_control)
drake::loadd(res_all)

res_ee_control %<>% filter(term == score_name)
n_05 <- sum(res_ee_control$p.value < 0.05)
n_05_33 <- sum(res_ee_control$p.value < 0.05/33)

```

This theory predicts that genetic variants will be selected for (or against) in
proportion to their effect on earnings. Figure
\@ref(fig:plot-cors-earnings-educ) plots selection effects on each PGS against
that PGS's correlation with earnings in a respondent's first job[^earnings-est],
and against its correlation with educational attainment, a predictor of lifetime
earnings. The raw relationships (left column) are strongly negative. If we plot
selection effects controlling for age at first live birth (right column), the
effect reverses and becomes positive. Thus, the labour market appears to
play an important role in natural selection. Substitution effects
dominate income effects overall, in accordance with the widespread association 
between income and lower fertility [@becker1960economic;@jones2006economic].

[^earnings-est]: Earnings are estimated from the 2007 Annual Survey of Hours and
Earnings, using the SOC 2000 job code.

We also re-estimate equation \@ref(eq:regression) for each
PGS, controlling for earnings in first job and education levels. Effect sizes
are generally reduced (appendix Figure \@ref(fig:plot-res-earn-educ)), and only
`r n_05_33` out of 33 PGS are significant at $p = 0.05/33$, suggesting that
earnings and education indeed mediate the effect of PGS on fertility.

```{r plot-cors-earnings-educ, fig.cap="Selection effects, with and without controls for age at first live birth, by correlations with earnings and educational attainment. Each point represents one PGS."}

drake::loadd(res_all)
drake::loadd(res_age_flb)

cors_income <- cor(famhist[score_names], famhist$first_job_pay, use = "pairwise")
# TODO: use age_left_edu when this includes college

cors_educ <-   cor(famhist[score_names], famhist$age_fulltime_edu, use = "pairwise")

effect_size <-  res_all %>% 
                  filter(dep.var == "N children", reg.type == "raw") %>% 
                  pull(estimate)

effect_size_flb <-  res_age_flb %>% 
                      filter(term != "age_flb") %>% 
                      pull(estimate)

dfr <- data.frame(
        Correlation  = rep(c("Earnings", "Education"), each = 33, 2),
        cor          = rep(c(cors_income[,1], cors_educ[,1]), 2),
        effect_size  = c(rep(effect_size, 2), rep(effect_size_flb, 2)),
        Controls     = rep(c("None", 
                        "Age at first live birth"), each = 66)
      )

dfr %>% 
      mutate(
        Controls = fct_relevel(Controls, "None")
      ) %>% 
      ggplot(aes(effect_size, cor, colour = Correlation)) + 
        geom_point() +
        facet_grid(Correlation ~ Controls, labeller = label_both) + 
        my_vline + 
        my_hline +
        labs(x = "Selection effect", y = "Correlation") +
        scale_colour_brewer(type = "qual", palette = 2) +
        theme(legend.position = "none")

```

This theory can explain why selection effects have opposite signs when AFLB is
controlled for. Suppose that people are aware of their endowments of human
capital, and that education and innate human capital are complements. They
choose how long to stay in education. After leaving education they enter the
labour market and form families. If so, then people with less human capital will
leave education earlier and have their first child earlier [cf.
@caucutt2002women; @monstad2008education]. This is a pure substitution effect: the income effect plays
no role in the decision to leave education, since people are not yet earning
wages.[^no-borrow] Whilst in the labour market, higher earners will have more
children (i.e. income effects dominate). These assumptions will lead to the
pattern we observe: earnings-increasing variants lead to later first births (and
fewer children overall), but more children for any given AFLB.

[^no-borrow]: This argument assumes that capital market imperfections hinder
people from borrowing against future income.

The theory can also explain why natural selection is weaker among higher-income
parents. @becker1976child show that if parents care about child quality as well
as quantity, under certain assumptions the income effect will be stronger at
higher income levels; this can lead to a U-shaped relationship between income
and fertility. Empirically, @cohen2013financial find that income decreases 
fertility at low income levels but increases it at higher income levels.

Lastly, it is often assumed that the substitution effect will dominate for lone
parents, or those in unstable relationships, since they have less opportunity to
share childcare responsibilities, while the income effect will dominate for
couples who are able to reap the gains from specialization. Indeed, US fertility
decreases faster with education among single mothers than married mothers
[@baudin2015fertility].If so, then earnings-increasing genetic variants may
decrease fertility more among single parents than among couples, as in our data.
The same idea could explain our results for lifetime number of sexual partners,
if this is associated with relationship stability.

# Discussion

Previous work has documented natural selection in modern populations on variants
underlying many polygenic scores [@beauchamp2016genetic; @kong2017selection;
@sanjak2018evidence]. We show that natural selection effects are highly
concentrated among specific subgroups of the population, including poorer
people, less educated people, younger parents and those with more lifetime
sexual partners. Indeed, among older parents selection effects are actually
reversed. Furthermore, the size of selection effects correlates with the
association between a PGS and labour market earnings. The economic theory of
fertility provides a parsimonious explanation for this.

An alternative theory is that traits selected for are linked to
externalizing behaviour and low time discount rates, via the channel of
early sexual behaviour [@Mills2020.05.06.081273]. The data here provide
some support for this: scores which might plausibly be linked to
externalizing behaviour, like ADHD and age at smoking initiation, are
selected for. However, this theory is less good at explaining variation
in selection across the full range of scores, including physical measures, e.g.
waist-hip ratio and BMI. Externalizing behaviour also does not explain
why selection effects change sign among older parents. Overall, we believe
that the economic theory is the leading explanation.

PGS which correlate with high (low) earnings and more (less) education are being
selected against (selected for). In addition, many of the phenotypes under
positive selection are linked to poor health, or would be seen as undesirable to
have by many people. For example, most people would probably prefer to have high
educational attainment, a low risk of major depressive disorder, and a low risk
of coronary artery disease, but natural selection is pushing against these
traits. Potentially, this could increase the health burden on modern societies,
but that depends on effect sizes. Our results suggest that naïve estimates can
be affected by sample ascertainment bias. This problem may be less serious in
surveys which aim to be representative (as the UK Biobank does not).
Nevertheless, not all respondents consent to the collection of genetic data: for
instance, completion rates for genotype data in the US Health and Retirement
Study were around 80-85% [@hrs2020]. Researchers should be aware of the risks
of ascertainment when studying modern natural selection.

In addition, we do not know whether estimated effect sizes of natural
selection will increase or decrease as more accurate polygenic scores
are produced. We are also unsure whether genetic variants underlying other
phenotypes will show a similar pattern of natural selection as these ones. 
In short, it is probably too early to tell whether modern natural
selection is substantively important or not.

Research on natural selection is controversial. It brings to mind eugenics,
which has been used to justify mass murder and human rights violations, in the
past and possibly the present [@mao1998chinese; @sleeboom2011genetic;
@zenz2020]. Whilst we claim scientific legitimacy for our empirical results, we
have no special authority to speak on how society should react to
them. This is a matter for open debate, in which scientists take part as
equals among many. We therefore choose not to discuss policy implications here.

\FloatBarrier
# Appendix

## Natural selection by sex

```{r plot-sexes, fig.cap = "Selection effects by sex. Lines show significant differences."}

drake::loadd(res_sex)
n_regs <- as.double(nrow(res_sex))

res_sex_wide <- res_sex %>% 
      select(term, estimate, sex, diff.p.value) %>% 
      pivot_wider(names_from = "sex", values_from = "estimate") %>% 
      mutate(
        min_est = pmin(Female, Male),
        max_est = pmax(Female, Male),
        term = pretty_names(term),
        term = fct_reorder(term, Female)
      )

res_sex %>% 
      mutate(
        term = pretty_names(term),
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, fill = sex, 
            colour = p < 0.05/{{n_regs}})) +
        my_vline +
        geom_linerange(data = res_sex_wide, 
          aes(
            y      = term,
            x      = NULL,
            fill   = NULL,
            xmin   = min_est, 
            xmax   = max_est, 
            colour = diff.p.value < 0.05/{{n_regs}}
          ), 
        group = "") +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 2) +
        my_labs() +
        my_theme()

```

\FloatBarrier
## Controlling for AFLB

```{r plot-age-flb, fig.cap = "Selection effects controlling for age at first live birth (women only). Effect sizes for women without controls are shown for comparison"}

n_regs <- as.double(nrow(res_age_flb))

res_combined %>% 
      mutate(
        score_name = pretty_names(score_name),
        score_name = fct_reorder(score_name, estimate, order_abs(2)),
        Regression = ifelse(reg.type == "raw", "Bivariate", "Controlled for AFLB")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = score_name, x = estimate, fill = Regression, 
        color = p < 0.05/{{n_regs}})) + 
        geom_vline(xintercept = 0, linetype = "dashed", color = "grey20") +
        pval_geom_point() +
        pval_scale() +
        scale_fill_brewer(type = "qual", palette = 3) +
        my_labs() +
        my_theme()
```

\FloatBarrier
## Weighted regressions

Figure \@ref(fig:plot-res-wt) plots effect sizes among the entire
sample, using our three weighting schemes. We include unweighted regressions
for comparison. Table \@ref(tab:tbl-res-wt) gives effect sizes as a
proportion of the unweighted effect size.

```{r tbl-res-wt}

rel_sizes %>% 
      arrange(desc(Geographical)) %>% 
      bind_rows(rel_sizes_summary) %>% 
      rename(PGS = term) %>% 
      as_hux() %>% 
      insert_row("", "Weighting", fill = "") %>% 
      set_caption(
        "Weighted effect sizes as a proportion of unweighted"
      ) %>% 
      merge_cells(1, 2:4) %>% 
      set_align(1, 2, "centre") %>% 
      set_header_rows(1, TRUE) %>% 
      theme_compact() %>% 
      style_headers(bold = TRUE) %>% 
      set_bottom_border(1, 2:4) %>% 
      set_bottom_border(2, everywhere) %>% 
      set_italic(final(2), 1) %>% 
      set_font_size(8) %>% 
      set_width(0.7) %>% # ensures line wrapping below
      set_col_width(c(.34, .22, .22, .22)) %>% 
      add_footnote(paste0(
        "Only consistent and significant (when unweighted) estimates ",
        "are shown.\n",
        "Age/Qual/AFLB as a proportion of unweighted regressions ",
        "for females only."),
        font_size    = 6, 
        top_padding  = 0,
        left_padding = 0
      )

```

```{r plot-res-wt, fig.cap = "Effects of polygenic scores on fertility. Weighted regressions."}

n_regs <- length(score_names)
res_wt_combined %>% 
  mutate(
    Weights  = fct_relevel(Weights, "Geographical", "Age/Qualification",
                "Age/Qual/AFLB", "None"),
    term     = pretty_names(term),
    term     = fct_reorder(term, estimate, order_abs(1))
  ) %>% 
  rename(p = p.value) %>% 
  ggplot(aes(
    x = estimate, 
    y = term, 
    shape = Weights, 
    fill = Weights, 
    colour = p < 0.05/{{n_regs}}
  )) +
  geom_point(size = 1.5, stroke = 0.5) +
  my_vline +
  pval_scale() +
  scale_fill_manual(values = c(
    "None"              = "grey15", 
    "Geographical"      = "dodgerblue4",
    "Age/Qualification" = "steelblue2",
    "Age/Qual/AFLB"      = "darkseagreen4"
  )) +
  scale_shape_manual(values = c(
    "None" = "square filled", 
    "Geographical" = "circle filled",
    "Age/Qualification" = "circle filled", 
    "Age/Qual/AFLB" = "circle filled"
  ))

```

\FloatBarrier
## Parents' generation

Figure \@ref(fig:plot-res-sibs) shows regressions of *number of siblings* on
polygenic scores. For comparison, we show regressions on the respondents'
generation (*number of children*), excluding respondents with no
children.

```{r plot-res-sibs, fig.cap = "Selection effects, respondents' parents vs. respondents"}
drake::loadd(res_sibs_parent_weights)
drake::loadd(res_children_comparison)

res_sibs_kids <-  bind_rows(
                    Siblings = res_sibs_parent_weights, 
                    Children = res_children_comparison,
                    .id = "Dep. var."
                  )

standard_ggplot(res_sibs_kids, `Dep. var.`, n_regs = length(score_names))
```

```{r calc-period-regs}

drake::loadd(res_period_parents)
drake::loadd(res_period_children)

res_period <- bind_rows(res_period_parents, res_period_children)

res_period %<>% 
      tidyr::extract(term, c("period", "score_name"),
        regex="year_split(.*):(.*)")
n_tests <- nrow(res_period)/2

summary_period <- res_period %>% 
      select(-std.error, -statistic) %>% 
      pivot_wider(
        names_from  = period, 
        values_from = c(estimate, conf.low, conf.high, p.value)
      ) %>% 
      group_by(score_name, children) %>% 
      mutate(
        significant = diff.p.value < 0.05/n_tests,
        sign_early  = sign(estimate_early),
        sign_late   = sign(estimate_late),
        size_inc    = abs(estimate_early) < abs(estimate_late),
        change_sign = sign_early != sign_late,
        Change      = case_when(
                        ! significant ~ "Insignificant",
                        change_sign   ~ "Change sign",
                        size_inc      ~ "Size increasing",
                        ! size_inc    ~ "Size decreasing"
                      )
      ) %>% 
      select(score_name, children, Change, significant, change_sign, 
        sign_early, sign_late) %>% 
      mutate(children = ifelse(children, "Respondents", "Parents")) %>% 
      pivot_wider(names_from = children, 
        values_from = c(Change, significant, change_sign, sign_early, 
        sign_late)) %>% 
      mutate(overall = case_when(
              ! change_sign_Parents & ! change_sign_Respondents & 
                sign_late_Parents == sign_early_Parents ~ 
                "Consistent",
              TRUE ~ "Changes direction"
            ))

```

To check whether effect sizes were changing over time, we ran
regressions interacting PGS with birth year, median split at 1950
("early born" versus "late born"). We use both *number of children* and
*number of siblings* as a dependent variable. We weight using
age/qualification cells, and further adjust for selection in the
parents' generation (see above).

Tables \@ref(tab:tbl-siblings-by-period) and
\@ref(tab:tbl-children-by-period) summarize the results. There is no
change in natural selection within the parents' generation. In the
respondents' generation, effect sizes were significantly larger in absolute size among the late-born for five PGS: autism, cognitive ability, EA2, EA3 and
extraversion. These changes are inconsistent with the intergenerational change, 
where effect sizes were if anything larger among the earlier, parents' generation.
Three PGS showed significant changes in sign: alcohol use
(positive to negative); conscientiousness (negative to positive); and
type II diabetes (negative to positive). Overall, there is weak evidence for
change over time, and the clearest result is that the direction of selection is
broadly consistent.

```{r tbl-siblings-by-period}

period_table <- function (tbl, caption) {
  tbl %>% 
      as_hux() %>%
      set_bold(1, everywhere) %>% 
      set_bottom_border(1, everywhere) %>%

      add_footnote(glue("Significance is measured at p < 0.05/{n_tests}"), 
        border = 0.4, font_size = 8) %>% 
      set_contents(1, everywhere, 
        c("Change", "Number of scores")) %>% 
      set_right_border(final(1), 1, 0) %>% 
      set_top_padding(1) %>% 
      set_bottom_padding(1) %>%
      set_left_padding(everywhere, 1, 0) %>% 
      set_right_padding(everywhere, 2, 0) %>% 
      set_caption(caption) %>% 
      set_caption_width(1)
}

summary_period %>% 
      group_by(Change_Parents) %>% 
      count() %>% 
      period_table("Change in effect sizes between parents of early and late born respondents (regressions on number of siblings)")

```

```{r tbl-children-by-period}
summary_period %>% 
      group_by(Change_Respondents) %>% 
      count() %>% 
      period_table("Change in effect sizes between early and late born respondents (regressions on number of children)")
```

\FloatBarrier
## Controlling for principal components

Polygenic scores could capture effects that are really due to population
stratification, although would not change our results for natural
selection of the scores. Figure \@ref(fig:plot-regs-controlled-pcs) shows
selection effects on polygenic scores residualized for the top 100
principal components of the genetic data, calculated within the UK
Biobank population.

```{r calc-regs-controlled}

drake::loadd(res_all)

res_all_controlled <- res_all %>% 
  select(reg.type, dep.var, term, estimate) %>% 
  pivot_wider(names_from = c(reg.type, dep.var), values_from = estimate) %>% 
  mutate(
    consistent = sign(`raw_N siblings`) == sign(`raw_N children`) &
                 sign(`raw_N siblings`) == sign(`controlled_N siblings`) &
                 sign(`raw_N siblings`) == sign(`controlled_N children`)
  )

n_pgs  <- as.double(nrow(res_all_controlled))
prop_consistent_controlled <- mean(res_all_controlled$consistent)
controlled_smaller_sibs <- abs(res_all_controlled[["controlled_N siblings"]]) < 
  abs(res_all_controlled[["raw_N siblings"]])
controlled_smaller_children <- abs(res_all_controlled[["controlled_N children"]]) < 
  abs(res_all_controlled[["raw_N children"]])
median_prop_sibs <- median(abs(res_all_controlled[["controlled_N siblings"]]) /
    abs(res_all_controlled[["raw_N siblings"]]))
median_prop_children <- median(abs(res_all_controlled[["controlled_N children"]]) /
    abs(res_all_controlled[["raw_N children"]]))

```

In siblings regressions, effect sizes are smaller when residualizing for
principal components -- sometimes much smaller, as in the case of
height. `r sum(controlled_smaller_sibs)` out of `r n_pgs` "controlled"
effect sizes have a smaller absolute value than the corresponding "raw"
effect size. The median proportion between raw and controlled effect
sizes is `r median_prop_sibs`. Among the children regressions, this no
longer holds. Effect sizes are barely affected by controlling for
principal components.

Overall, `r prop_consistent_controlled * 100` per cent of effect sizes
are consistently signed across all four regressions (on children and
siblings, and with and without residualization).

```{r plot-regs-controlled-pcs, fig.cap = "Effects of residualized polygenic scores on number of siblings/children."}

drake::loadd(res_all)

n_regs <- as.double(nrow(res_all))
res_all %>% 
      mutate(
        term    = pretty_names(term),
        term    = reorder_within(term, estimate, dep.var, order_abs(1)),
        dep.var = fct_inorder(dep.var) 
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, fill = reg.type, 
        colour = p < 0.05/{{n_regs}})) +
        my_vline + 
        facet_wrap(vars(dep.var), scales = "free_y") +
        pval_geom_point(position = position_dodge(0.4)) +
        my_labs() +
        pval_scale() +
        my_fill_scale(n = 2, guide = guide_legend(title = "")) +
        my_theme() +
        scale_y_reordered()

```

```{r}
drake::loadd(res_pcs)
n_regs <- as.double(nrow(res_pcs))
prop_sig <- res_pcs %>% 
      group_by(dep.var) %>% 
      summarize(sig = sum(p.value <= 0.05/nrow(res_pcs))) %>% 
      pull(sig, name = dep.var)
```

To get a further insight into this we regress *siblings* and *children*
on individual principal components. As Figure \@ref(fig:plot-pc-regs)
shows, effects are larger and more significant in siblings regressions.
`r prop_sig[["N siblings"]]` principal components significantly
predicted number of siblings, while only `r prop_sig[["N children"]]`
significantly predicted number of children.

```{r plot-pc-regs, fig.cap = "Effect of principal components of genetic data on number of siblings/children. Absolute effect sizes are plotted. Each dot represents one bivariate regression. Points are jittered on the Y axis.", fig.height = 2}

res_pcs %>% 
      mutate(
        estimate = abs(estimate),
        dep.var  = fct_relevel(dep.var, "N children")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, dep.var, colour = p < 0.05/{{n_regs}})) + 
        geom_jitter(height = 0.15, width = 0, alpha = 0.9) + 
        scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) +
        labs(x = "Effect size (log scale)", y = "") + 
        pval_scale(aesthetics = "colour")

```

## Selection by number of partners, controlling for age

Figure \@ref(fig:plot-n-partners-controlled) shows selection effects by
lifetime number of sexual partners. Age, age squared, and interactions
of age and age squared with PGS are controlled for.

```{r plot-n-partners-controlled, fig.cap = "Effect sizes on number of siblings interacted with number of sexual partners (median split at 3), respondent's age and age squared."}
drake::loadd(res_partners_controlled)
res_partners_controlled %<>%       
      filter(term != "(Intercept)", term != "lo_partnersTRUE") %>% 
      mutate(`N partners` = ifelse(grepl("TRUE:", term), "<= 3", "> 3"))

n_regs <- as.double(nrow(res_partners_controlled))
res_partners_controlled %>% 
      mutate(
        score_name = pretty_names(score_name)
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(
        y      = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        x      = estimate, 
        fill   = `N partners`, 
        colour = p < 0.05/{{n_regs}}
      )) + 
        my_vline +
        pval_geom_point() +
        facet_wrap(vars(sex), scales = "free_y") +
        my_fill_scale(n = 2) +
        pval_scale() +
        my_theme(legend.position = "bottom") +
        my_labs(x = "", y = "") + 
        scale_y_reordered()

```

## Selection controlling for age at first live birth: respondents' parents

```{r calc-age-birth-parents}

drake::loadd(res_age_birth_parents)
drake::loadd(res_all)

res_sibs_raw <- filter(res_all, dep.var == "N siblings", reg.type == "raw")
compare_abp_raw <- res_age_birth_parents %>% 
      filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
      left_join(res_sibs_raw, by = "term")

cor_raw_fath <- with(
  compare_abp_raw %>% filter(control == "fath_age_birth"), 
  cor(estimate.x, estimate.y))
cor_raw_moth <- with(
  compare_abp_raw %>% filter(control == "moth_age_birth"), 
  cor(estimate.x, estimate.y))
```

Among the parents' generation, we can control for age at first live
birth using the subsets of respondents who reported their mother's or
father's age, and who had no elder siblings. We run regressions on
*number of siblings* on these subsets, controlling for either parent's
age at their birth. Figure \@ref(fig:plot-age-birth-parents) shows the
results. Effect sizes are very similar, whether controlling for father's
or mother's age. Unlike for the respondents' generation, effect sizes
are positively correlated with the effect sizes from bivariate
regressions without the age at birth control (father's age at birth:
$\rho$ `r cor_raw_fath`; mother's age at birth: $\rho$
`r cor_raw_moth`).

```{r plot-age-birth-parents, fig.cap = "Selection effects controlling for parents' age at birth, eldest siblings"}

n_regs <- as.double(nrow(res_age_birth_parents)/2)

res_age_birth_parents %>% 
    filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
    mutate(
      term    = pretty_names(term),
      term    = fct_reorder(term, estimate, order_abs(2)),
      Control = ifelse(control == "fath_age_birth", 
                  "Father's age at birth", 
                  "Mother's age at birth"
                )
    ) %>% 
    rename(p = p.value) %>% 
    ggplot(aes(y = term, x = estimate, fill = Control, 
      colour = p < 0.05/{{n_regs}})) + 
      pval_geom_point(alpha = 0.8) +  
      my_vline +
      my_fill_scale(2) +
      pval_scale() +
      my_labs()

```

## Effects of PGS on age at first live birth

```{r calc-age-flb-dv}

drake::loadd(res_age_flb_dv)
drake::loadd(res_age_birth_parents_dv)

compare_flb_dv <- left_join(
        res_age_flb_dv           %>% select(term, estimate),
        res_age_birth_parents_dv %>% select(term, estimate, dep.var),
        by = "term"
      )

cor_flb_fath <- with(compare_flb_dv %>% filter(dep.var == "fath_age_birth"),
        cor(estimate.x, estimate.y)
      )
cor_flb_moth <- with(compare_flb_dv %>% filter(dep.var == "moth_age_birth"),
        cor(estimate.x, estimate.y)
      )
```

Our results suggest that polygenic scores may directly correlate with
age at first live birth. Figure \@ref(fig:plot-age-flb-dv) plots
estimated effect sizes from bivariate regressions for respondents, and
Figure \@ref(fig:plot-age-birth-parents-dv) does the same for their
parents. Effect sizes are reasonably large. They are also very highly
correlated across generations. Effect sizes of PGS on father's age at
own birth, and on own age at first live birth, have a correlation of
`r cor_flb_fath`; for mother's age and own age it is `r cor_flb_moth`

```{r plot-age-flb-dv, fig.cap = "Effects of polygenic scores on age at first live birth"}

n_regs <- as.double(nrow(res_age_flb_dv))

res_age_flb_dv %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, term, colour = p < 0.05 / {{n_regs}})) + 
      my_vline +
      pval_geom_point(fill = "steelblue4") +
      pval_scale() + 
      my_labs() +
      my_theme()

```

```{r plot-age-birth-parents-dv, fig.cap = "Effects of polygenic scores on parents' age at own birth, eldest siblings"}

n_regs <- as.double(nrow(res_age_birth_parents_dv))

res_age_birth_parents_dv %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1)),
        "Dependent variable" = dplyr::recode(dep.var, 
          "fath_age_birth" = "Father's age", 
          "moth_age_birth" = "Mother's age"
        )
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, term, fill = `Dependent variable`, 
        colour = p < 0.05 / {{n_regs}})) + 
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 2) +
        my_labs()

```

\FloatBarrier
## Controlling for earnings and education

```{r plot-res-earn-educ, fig.cap = "Selection effects controlling for earnings in first job (estimated by median earnings from ASHE 2007 for the SOC2000 job code) and education (left education before 16, 16-18, or after 18). Raw effects are shown for comparison. Lines are 95% confidence intervals uncorrected for multiple testing."}

drake::loadd(res_ee_control)
drake::loadd(res_all)

res_ee_control %<>% filter(term == score_name)

res_all %<>% 
  filter(term == score_name, reg.type == "raw", dep.var == "N children") %>% 
  select(-dep.var, -reg.type)

res_compare <- bind_rows(
                 Raw = res_all, 
                 Controlled = res_ee_control, 
                 .id = "Regression"
               )

standard_ggplot(res_compare, fill_col = Regression, n_regs = 33) +
  geom_linerange(
    aes(
      xmin = estimate - 2 * std.error, 
      xmax = estimate + 2 * std.error
    ), 
    alpha = .5
  )

```

\FloatBarrier
## Genetic correlations with EA3

Another way to examine the "earnings" theory of natural selection is to
compare selection effects of PGS with their genetic correlation with
educational attainment (EA3). Since EA3 strongly predicts earnings, if
earnings drives differences in fertility, we'd expect a correlation
between the two sets of results. Figure
\@ref(fig:plot-rgs-by-effect-size) shows this is so.

TODO: ref source of data.

```{r plot-rgs-by-effect-size, fig.cap = "Effect size on n siblings by genetic correlation with EA3"}

drake::loadd(rgs)
drake::loadd(res_all)

rgs <- left_join(rgs, 
        res_all %>% filter(reg.type == "raw", dep.var == "N children"), 
        by = c("p2" = "term"))

rgs %>% 
      filter(p2 != "EA3_excl_23andMe_UK", p2 != "EA2_noUKB") %>% 
      ggplot(aes(rg, estimate)) + 
        geom_point() +
        my_vline +
        my_hline +
        labs(x = "rG with EA3", y = "Effect on number of children")

```

```{r correlate-effect-sizes, eval = FALSE}
combined_results <- res %>% 
      mutate(PCs = ifelse(reg.type=="raw", "no", "yes")) %>% 
      dplyr::select(-reg.type) %>% 
      {bind_rows(bivariate = ., multivariate = all_res, .id = "reg.type")} %>% 
      arrange(term, reg.type, PCs) %>% 
      dplyr::select(term, reg.type, PCs, estimate) %>% 
      pivot_wider(names_from = c(reg.type, PCs), values_from = estimate)

cor(combined_results[-1], use = "complete")

```

\FloatBarrier
## Number of children

Figure \@ref(fig:plot-n-children-bar) shows the full distribution of
number of children born for different ventiles of the EA3 polygenic
score. The strongest relationship seems to be for having 0 children
versus 1 or more.

```{r plot-n-children-bar, fig.cap = "Number of children by ventiles of EA3 PGS"}
famhist %>% 
      select(n_children, EA3_excl_23andMe_UK, sex) %>% 
      mutate(
        "EA3 Ventile" = chop_equally(EA3_excl_23andMe_UK, 20, labels = 1:20),
        "N Children"  = santoku::chop(n_children, 0:4, labels = c(0:3, "4+")),
        Sex           = factor(sex, levels = 0:1, labels = c("Female", "Male"))
        ) %>% 
      filter(! is.na(`EA3 Ventile`), ! is.na(`N Children`)) %>% 
      ggplot(aes(`EA3 Ventile`, fill = `N Children`, group = `N Children`)) + 
        geom_bar(position = position_fill(reverse = TRUE), alpha = 0.7) + 
        facet_wrap(vars(Sex)) +
        scale_fill_viridis_d(option = "A") +
        scale_y_continuous(n.breaks = 10, labels = scales::percent) +
        theme(legend.position = "bottom")

```

## Causality

Different polygenic scores are correlated. Table
\@ref(tab:tbl-pgs-corrs) shows the top correlations in the sample.
Because of this, bivariate correlations between PGS and number of
children might be driven by other genetic scores. To explore which
polygenic scores are driving negative selection, we run a single omnibus
regression of *number of children* on all the PGS. We exclude EA2,
waist-hip ratio, waist-circumference, and "Hip combined" since they are
highly correlated with other scores, which could make our estimates
unstable. Figure \@ref(fig:plot-together) shows the results.
Interestingly, several PGS remain independently significant, although
effect sizes are reduced.

```{r tbl-pgs-corrs}
cor_tbl <- cor(famhist[score_names], use = "pair") %>% 
      as_tibble() %>% 
      mutate(score_name = score_names, .before = 1) %>% 
      pivot_longer(cols = -score_name, values_to = "Correlation") %>% 
      filter(score_name < name) %>% 
      arrange(desc(Correlation)) %>% 
      slice(1:10) %>% 
      mutate(across(c(score_name, name), pretty_names))

as_hux(cor_tbl, add_colnames = TRUE) %>% 
      set_contents(1, 1:2, c("PGS", "PGS")) %>% 
      theme_basic(header_col = FALSE) %>% 
      set_bottom_border(1, everywhere, 0.4) %>% 
      set_bottom_border(final(1), everywhere, 0.4) %>% 
      set_tb_padding(4) %>% 
      set_caption("Top 10 correlations between polygenic scores")

```

```{r plot-together, fig.cap = "Partial correlations with number of children"}

loadd(res_together)
n_regs <- as.double(nrow(res_together))
res_together %>% 
      mutate(
        term = pretty_names(term),
        dep.var = dplyr::recode(dep.var, 
          "n_children" = "N Children", "n_sibs" = "N Siblings"),
        dep.var = fct_relevel(dep.var, "N Siblings")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = reorder_within(term, estimate, dep.var, order_abs(1)), 
        x = estimate, color = p < 0.05/{{n_regs}}, fill  = PCs)) + 
        pval_geom_point(alpha = 0.8) +
        my_vline +
        facet_wrap(vars(dep.var), scales = "free_y") +
        pval_scale() +
        my_fill_scale(n = 2) +
        my_labs() + 
        scale_y_reordered() +
        theme(legend.position = "bottom")
```

\FloatBarrier

# References