### 2022-01-11 why-natural-selection.Rmd
Tweaks for AJHG: deblind; new cover letter

---
title: "Human capital mediates natural selection in contemporary humans"
author: "David Hugh-Jones \\thanks{Corresponding author. School of Economics, University of East Anglia, Norwich, UK. Email: D.Hugh-Jones@uea.ac.uk}, Abdel Abdellaoui \\thanks{Department of Psychiatry, Amsterdam UMC, University of Amsterdam, Amsterdam, The Netherlands. Email: a.abdellaoui@amsterdamumc.nl}"
date: "`r format(Sys.Date(), '%d %B %Y')`"
abstract: "\\noindent Natural selection has been documented in contemporary humans, but  little is known about the mechanisms behind it. We test for natural selection through the association between 33 polygenic scores and fertility, across two generations, using data from UK Biobank (N = 409,629 British subjects with European ancestry). Consistently over time, polygenic scores associated with lower (higher) earnings, education and health are selected for (against). Selection effects are concentrated among lower SES groups, younger parents, people with more lifetime sexual partners, and people not living with a partner. The direction of natural selection is reversed among older parents, or after controlling for age at first live birth. These patterns are in line with economic theories of fertility, in which earnings-increasing human capital may either increase or decrease fertility via income and substitution effects in the labour market. Studying natural selection can help us understand the genetic architecture of health outcomes: we find evidence in modern day Great Britain for multiple natural selection pressures that vary between subgroups in the direction and strength of their effects, that are strongly related to the socio-economic system, and that may contribute to health inequalities across income groups."
bibliography: "negative-selection.bib"
output: 
  bookdown::pdf_document2:
    toc: false
    latex_engine: xelatex
    number_sections: true
    keep_tex: true
header-includes:
  - \usepackage{subfig}
  - \usepackage{setspace}\doublespacing
  - \usepackage{placeins}
  - \usepackage[format=plain, labelfont={bf,it}, textfont=it]{caption}
  - \usepackage{titlesec}
  - \titleformat*{\section}{\sffamily\LARGE}
  - \titleformat*{\subsection}{\sffamily\itshape\Large}
  - \hypersetup{colorlinks = true, linkcolor = {blue}, linkbordercolor = {white}}
  - \usepackage{amsthm}
  - \theoremstyle{plain}
  - \newtheorem{lem}{\protect\lemmaname}
  - \providecommand{\lemmaname}{Lemma}
  - \usepackage{etoc}
  - \usepackage{lineno}
  - \linenumbers
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
mainfont: Baskerville
mathfont: Baskerville
sansfont: "Gill Sans"
---

```{r setup, include = FALSE}

library(drake)
library(magrittr)
library(dplyr)
library(forcats)
library(ggplot2)
library(tidyr)
library(santoku) # keep this after tidyr as it masks chop
library(purrr)
library(glue)
library(broom)
library(huxtable)
loadNamespace("scales")

# nasty hack to make add_ashe_income work later:
if (! exists("add_ashe_income")) source("~/import-ukbb-data/import-ukbb-data.R")

drake::loadd(rgs)
drake::loadd(score_names)
drake::loadd(famhist)

options(digits = 2)
options(dplyr.summarise.inform = FALSE) # shut up about summarize

knitr::opts_chunk$set(
        echo = FALSE, 
        warning = FALSE, 
        cache = FALSE, 
        error = TRUE,
        fig.height = 3.5
      )

knitr::knit_hooks$set(
  inline = function (x) {
    if (is.numeric(x)) x <- as.character(round(x, getOption("digits")))
    x <- gsub("-", "\u2212", x)
    paste(as.character(x), collapse = ", ")
  }
)

options(huxtable.long_minus = TRUE)
huxtable::set_default_properties(latex_float = "h!")

theme_set(theme_minimal())

my_hline <- geom_hline(yintercept = 0, colour = "grey20", linetype = "dotted")
my_vline <- geom_vline(xintercept = 0, colour = "grey20", linetype = "dotted")
pval_palette <- c("FALSE" = "grey65", "TRUE" = "#FAC32D")

standard_ggplot <- function (
        dfr, 
        fill_col, 
        n_regs, 
        ...,
        score_col = quo(term), 
        order_idx = 1, 
        fill_direction = 1,
        n_cats    = NULL,
        conf_int  = NULL
      ) {
  if (! missing(score_col)) score_col <- enquo(score_col)

  mfc <- missing(fill_col)
  fill_col <- if (mfc) quo(NULL) else enquo(fill_col)
  if (missing(n_cats)) {
    n_cats <- if (mfc) 1 else length(unique(pull(dfr, {{fill_col}})))
  }

  if (is.null(conf_int)) conf_int <- "conf.low" %in% names(dfr)

  n_regs <- as.double(n_regs)

  confint_geom_segment <- if (conf_int) confint_geom_segment() else NULL

  dfr %>% mutate(
          !! score_col := pretty_names(!! score_col),
          !! score_col := fct_reorder(!! score_col, estimate, order_abs(order_idx))
        ) %>% 
        rename(p = p.value) %>% 
        ggplot(aes(estimate, {{score_col}}, yend = {{score_col}}, 
                     fill = {{fill_col}}, ..., colour = p < 0.05/{{n_regs}})) +
          my_vline + 
          confint_geom_segment +
          pval_geom_point(fill = if(rlang::quo_is_null(fill_col)) "steelblue") +
          pval_scale() +
          my_fill_scale(n = n_cats, direction = fill_direction) +
          my_labs() +
          my_theme() +
          scale_linetype_manual(name = "", values = c("95% c.i." = 1))

}

my_theme <- function (...) theme(axis.text.y = element_text(size = 7), ...)

pval_scale <- function (...) scale_color_manual(values = pval_palette, ...)

pval_geom_point <- function (size = 1.5, stroke = 0.5, fill = NULL, ...) {
  if (is.null(fill)) {
    geom_point(shape = 21, size = size, stroke = stroke, ...)
  } else {
    geom_point(shape = 21, size = size, stroke = stroke, fill = fill, ...)
  }
}

confint_geom_segment <- function () {
  geom_segment(aes(x = conf.low, xend = conf.high, linetype = "95% c.i."), 
                    color = "grey60", alpha = 0.25, size = 1.1, 
                    lineend = "round")
}

my_fill_scale <- function (n, aesthetics = "fill", ...) {

  fill_colors <- c("steelblue4", "#D63C3C")

  if (n <= 2) {
    dots <- list(...)
    dots$direction <- NULL
    do.call(scale_colour_manual, c(list(values = fill_colors[1:n],
                                        aesthetics = aesthetics), 
                                   dots))
  } else {
    scale_color_brewer(aesthetics = aesthetics, ...)
  }
}

my_labs <- function (x = "Effect size", y = "", ...) labs(x = x, y = y, ...)

# for use when reordering factors. Order by the nth score in a group
order_abs <- function (n = 1) {
  function (x) x[n] 
}

comma_num <- function (x) prettyNum(x, big.mark = ",")

pretty_names <- function (names) {
  pretty <- c(
    ADHD_2017               = "ADHD",
    age_at_menarche         = "Age at menarche",
    age_at_menopauze        = "Age at menopause",
    agreeableness           = "Agreeableness",
    ai_substance_use        = "Age at smoking initiation",
    alcohol_schumann        = "Alcohol use",
    alzheimer               = "Alzheimer",
    autism_2017             = "Autism",
    bipolar                 = "Bipolar",
    bmi_combined            = "BMI",
    body_fat                = "Body Fat",
    caffeine                = "Caffeine",
    cannabis                = "Cannabis (ever vs. never)",
    cognitive_ability       = "Cognitive Ability",
    conscientiousness       = "Conscientiousness",
    coronary_artery_disease = "Coronary Artery Disease",
    cpd_substance_use       = "Cigarettes per day",
    diagram_T2D             = "Type 2 Diabetes",
    dpw_substance_use       = "Drinks per week",
    EA2_noUKB               = "Educ. attainment 2 (no UKBB)",
    EA3_excl_23andMe_UK     = "Educ. attainment 3 (no UK)",
    eating_disorder         = "Eating disorder",
    extraversion            = "Extraversion",
    height_combined         = "Height",
    hip_combined            = "Hip circumference",
    MDD_PGC2_noUKB          = "Major Depressive Disorder",
    neuroticism             = "Neuroticism",
    openness                = "Openness",
    sc_substance_use        = "Smoking cessation",
    SCZ2                    = "Schizophrenia",
    si_substance_use        = "Smoking initiation",
    wc_combined             = "Waist circumference",
    whr_combined            = "Waist-hip ratio"
  )

  pretty[names]
}

n_in_regs <- function (var, dv = "n_children", data = famhist) {
  fh_subset <- data
  if (dv == "n_children") fh_subset %<>% filter(kids_ss)
  # all PGS have the same NA pattern, so we just use one
  tbl <- table(fh_subset[
                  ! is.na(fh_subset[dv]) & ! is.na(fh_subset$whr_combined), 
                  var
                ])

  c(tbl)
}

add_n <- function (var, famhist_var = var, dv = "n_children", data = famhist) {
  var <- as.factor(var)
  n <- n_in_regs(famhist_var, dv, data)
  levels(var) <- sprintf("%s (N = %s)", levels(var), 
                         prettyNum(n[seq_len(nlevels(var))], big.mark = ","))
  # var_n <- sprintf("%s (N = %s)", var, prettyNum(n[var], big.mark = ","))
  # var_n[is.na(var)] <- NA
  var
}

# from https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R
# for ordering within facets
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}

```

\normalem

```{r NOTES-TODO, eval=FALSE}
"

# TODO response to PNAS

* Only use education, not first job pay, in "controlled" regressions; check
  description is correct - done.
* Add CIs - done.
* Clarify collider bias/endogeneity - done.
* Formalize theory - done.
  - Comment on results where effects are actually reversed. This works
    nicely if you throw in a money cost - done computational example.
* Clarify causality? - done...
  - Helps to deal with the "collider bias" argument of reviewer
  - Might also be a better test of mediation via human capital
  - One other possibility. If we use only people with first profession
    teacher, that (as it happens) shuts down the EA3-income correlation.
    It also shuts down the EA3-fertility relationship, suggesting that
    income is indeed the mediator.
* Clarify "what we learn about genetic links to fertility by using economic theory to interpret relationships between polygenic scores and fertility."
  - Point is that we can explain how natural selection is occurring.
* "Discuss empirical predictions for AFLB grounded in dynamic theory"
  - hmm, tricky, maybe we can include it in discussion of economic results at
  end of formal section.
  - think our theory with 2 periods probably covers this
* Why don't women see stronger effects? But note, they do: not significant
  after correction for multiple testing, but a 25% difference. 
  - I think we leave this for reviewers.
* Quadratic regressions for parents - DONE

# Open questions/TODO

* Limit max no. of children in regressions (use an "in_children" bool)
  - not necessary. Very few have abnormally many.
* Consider doing separate regressions for "any children" vs. "n children | 1 or more"
  - heckit too slow and gives crazy answers, just use plm/ols
  - some variables do look different
* If we weight with the GHS, and plot the means of figure 1 using this w
  weighting, most things stay the same - but the change in EA is reversed! EA
  increases over time. Could this be because the sample includes many
  relatively educated older people? If so, then mean effects are *overestimated*
  by figure 1.
* Reweight other regressions (e.g. the omnibus regression in the appendix)?

* If you take "year of last job started" and see if it is after/before
  the first child, you can figure out whether mothers keep working.
  - better still, look at all jobs started and see if any are within e.g. 5 years of
  first child's birth
  - this could then be another way to get at income vs substitution effects. You'd
  predict negative selection effects among mothers who worked. Of course the
  choice to work is endogenous.
* double check which scores are reverse coded!

# WHAT MODEL FITS YOUR DATA
## ASSUMPTIONS
* Genetics produce human capital h. 
* h and education are compliments in producing earnings
* There are income and substitution effects on fertility:
  - income effects are stronger at high income levels (Becker and Tomes 1976)
    and among couples
* People seek partners when in the labour market. Women match a person, observe
  their quality, and decide whether to have a first child.
* Then they (may) have more children, either with the partner or not 
  (high quality partners stay around more)
* High quality partners increase household income

## RESULTS
* People with low h are more likely to leave education early and enter the labour market
  - because e and h are complements
  - this is part of the substitution effect, as they expect to spend some time
    raising children rather than being in work
* At low values of h, increases in h reduce fertility (substitution effect dominates)
  at higher values, smaller reductions or increases (income effect stronger)
  - because h predicts income
* In 1-parent households, increases in h reduce fertility (sub. effect), in 
  2-parent households, smaller reductions or increases (income effect stronger)
  - because h predicts income and because stronger income effect among couples
* People with low h start having children earlier (b/c leave educ earlier)
  - so part of the substitution effect comes via this
  - controlling for this the income effect dominates
  - hence increases in h reduce fertility less once we control for AFLB

## THOUGHTS
* Should we distinguish between men and women in couples?
  - within couples, male h should increase fertility (via earnings)
  - female h should decrease it (ditto?)
  - see explore-sex-vs-pay.R

## RIVAL MECHANISMS
* Lynn 2011 mentions two:
  - social class gradient on ability to delay gratification
  - modern society protects the weak (Darwin 1871, Spencer 1874)
    NB this latter doesn't predict more fertility among "dysgenic" ppl, just
    equal fertility and then genetic drift does its work
    - examples include modern sanitation, post-1800
    - and monogamy!
"
```

# Introduction

Living organisms evolve through natural selection, in which allele frequencies
change in the population through differential reproduction rates. Studying the
mechanisms behind natural selection can help us better understand how individual
differences in complex traits and disease risk arise [@benton2021influence].
Recent work confirms that natural selection is taking place in modern human
populations, using genome-wide analysis
[@Barban_2016;@beauchamp2016genetic;@conley2016assortative;@kong2017selection;@sanjak2018evidence;@FIEDER202216]. In
particular, genetic variants associated with higher educational attainment are
being selected against, although effect sizes appear small.

As yet we know little about the social mechanisms behind natural selection. A
potential explanation comes from the economic theory of fertility [@becker1960economic]. In this theory, higher potential earnings
have two opposite effects on fertility: a fertility-increasing *income effect*
(higher income makes children more affordable), and a fertility-lowering
*substitution effect* (time spent on childrearing has a higher cost in foregone
earnings). Thus, an individual's *human capital* -- skills and personality
traits which are valuable in labour markets -- can increase or decrease his or
her fertility. This can lead to selection on inherited variants which are linked
to human capital. Furthermore, the economic theory predicts systematic variation
in the relative strength of the income and substitution effects across different
social groups.

This study uses data from UK Biobank [@bycroft2018uk] to learn more about
contemporary natural selection. We test for natural selection on 33 different
polygenic scores by estimating their correlation with fertility. We extend the
analysis over two generations, using data on respondents' number of siblings as
well as their number of children. This is interesting because consistent natural
selection over multiple generations could lead to substantive effects in the
long run. Next, we examine correlations with fertility in different subgroups.
Across the board, selection effects are stronger in groups with lower income and
less education, among younger parents, people not living with a partner, and
people with more lifetime sexual partners. Outside these groups, effects are
weaker and often statistically insignificant. In some subgroups, the direction
of selection is even reversed.

We then show that a simple model of human capital, education and fertility
choices can give rise to these empirical results. To test the theory, we show
that polygenic scores' correlation with fertility is associated with their
correlation with education and earnings. Furthermore, a mediation analysis shows
that part of the correlation with fertility is indeed *via* education. Thus, contemporary natural selection on polygenic scores can be explained by scores' correlation with earnings-increasing human capital.

Lastly, we discuss the effects of natural selection. While measured effects on
mean genetic scores are small, natural selection may substantially increase the
correlation between genetic scores and income, i.e. the "genetic lottery"
[@harden2021genetic].

# Results

```{r calc-sex-ratio}
female_ratio <- mean(famhist$sex == 0)
```

We created polygenic scores for 33 traits in `r comma_num(nrow(famhist))`
individuals of European descent, corrected for ancestry using 100 genetic
principal components (see Materials and Methods). Figure
\@ref(fig:plot-means-over-time) plots mean polygenic scores in the sample by
5-year birth intervals. Several scores show consistent increases or declines
over this 30-year period, of the order of 5% of a standard deviation. These
changes could reflect natural selection within the UK population, but also
emigration, or ascertainment bias within the sample [@10.1093/aje/kwx246].

```{r plot-means-over-time, fig.height = 6, fig.cap = "Mean polygenic scores (PGS) by birth year in UK Biobank. Points are means for 5-year intervals. Lines are 95% confidence intervals. Green triangles show a significant linear increase over time (p < 0.05/33). Red squares show a significant decrease."}

drake::loadd(pgs_over_time)

pgs_over_time %>% 
      mutate(score_name = pretty_names(score_name)) %>% 
      ggplot(aes(YOB, group = score_name, colour = Change, shape = Change)) + 
      geom_pointrange(aes(ymin = score_lo, y = score, ymax = score_hi), 
        fatten = 1.25) +
      facet_wrap(~score_name, ncol = 5) +
      my_hline +
      scale_color_manual(values = c("-" = "red", "+" = "darkgreen", "o" = "black")) +
      scale_shape_manual(values = c("-" = "square", "+" = "triangle", "o" = "circle open")) +
      theme(
        strip.text  = element_text(size = 6), 
        axis.text.x = element_text(size = 7),
        legend.position = "none"
      ) +
      labs(y = "PGS", x = "Birth year (5 year intervals)")

```

To test more directly for natural selection, we regress respondents' number of
children ($N_i$) on their polygenic scores (PGS):

```{=tex}
\begin{equation}
N_i = \alpha + \beta\mathrm{PGS}_i + \varepsilon_i
\label{eq:regression}
\end{equation}
```

The "selection effect", $\beta$, reflects the strength of natural selection
within the sample. In fact, since polygenic scores are normalized, $\beta$ is
the expected polygenic score among children of the sample
[@beauchamp2016genetic].[^normalization] Note that equation
\eqref{eq:regression} does not control for many environmental and genetic
factors that could affect fertility, and as a result, $\beta$ is not an estimate
of the causal effect of a polygenic score on fertility. However, natural
selection is a correlational phenomenon: polygenic scores which correlate with
high fertility are being selected for, whatever the underlying causal
relationship.

[^normalization]: The selection effect $\beta$ equals $Cov(N, PGS)/Var(PGS)$
where $N$ is the number of children. Since PGS are normalized to variance 1 and
mean 0, this reduces to 
$Cov(N, PGS) \equiv E(N\ PGS) - E(N) E(PGS) = E(N\ PGS)$. This is the polygenic 
score weighted by the number of children, which is the average polygenic score 
in the next generation.

```{r calc-res-weighted}

drake::loadd(res_wt_flb_weights)
drake::loadd(res_wt_msoa_weights)
drake::loadd(res_wt_age_qual_weights)
drake::loadd(res_sibs_parent_weights)
drake::loadd(res_all)
drake::loadd(res_sex)

res_wt_combined <- bind_rows(
  None                = res_all %>% 
                          filter(
                            reg.type == "controlled", 
                            dep.var == "N children"
                          ) %>% 
                          select(-reg.type, -dep.var, -score_name),
  Geographical        = res_wt_msoa_weights,
  `Age/Qualification` = res_wt_age_qual_weights,
  `Age/Qual/AFLB`     = res_wt_flb_weights,
  .id = "Weights"
)

res_significance <- res_wt_combined %>% 
                      group_by(term) %>% 
                      summarize(
                        all = all(p.value < 0.05/33), 
                        any = any(p.value < 0.05/33)
                      )
res_sig_all <- sum(res_significance$all)
res_sig_any <- sum(res_significance$any)

res_wt_cb_wide <- res_wt_combined %>% 
                    tidyr::pivot_wider(
                      id_cols     = term,
                      names_from  = Weights,
                      values_from = estimate
                    ) %>% 
                    rowwise() %>% 
                    mutate(
                      consistent = all(c_across(-term) > 0) ||
                        all(c_across(-term) < 0)
                    ) %>% 
                    ungroup() %>% 
                    left_join(
                      res_wt_combined %>% 
                        filter(Weights == "None") %>% 
                        select(term, p.value),
                      by = "term"
                    ) %>% 
                    left_join(
                      res_sex %>% 
                        filter(sex == "Female") %>% 
                        select(term, estimate_females = estimate),
                      by = "term"
                    )

rel_sizes <- res_wt_cb_wide %>% 
                filter(
                  consistent, 
                  p.value < 0.05/length(score_names)
                ) %>% 
                mutate(
                  Geographical = Geographical/None,
                  `Age/Qualification` = `Age/Qualification`/None,
                  `Age/Qual/AFLB` = `Age/Qual/AFLB`/estimate_females
                ) %>% 
                select(-consistent, -None, -p.value, -estimate_females) %>% 
                mutate(
                  term = pretty_names(term)
                )

rel_sizes_summary <- rel_sizes %>% 
                summarize(
                  across(-term, ~ c(mean(.x), median(.x)))
                ) %>% 
                mutate(
                  term = c("Mean", "Median")
                )

mean_rel <- c(rel_sizes_summary[1, 1:3])
```

To correct for ascertainment bias, we weight participants using population data.
We try three alternative weighting schemes: (1) "Geographical" weighting, by
geography, age and presence/absence of a partner; (2) age and highest
educational qualification; and (3) for women only, age, highest qualification,
and age at first live birth (AFLB). Figure \@ref(fig:plot-res-wt) plots
selection effects among the entire sample, estimated with the three weighting
schemes, and also unweighted. Effects for `r res_sig_all` out of 33 polygenic
scores are significant at $p < 0.05/33$ under all four weighting schemes, with
`r res_sig_any` scores reaching significance under at least one scheme.
Weighting makes a large difference to estimates: mean effect sizes across all
polygenic scores are increased by a factor of `r mean_rel$Geographical`
(geographical weighting), `r mean_rel[["Age/Qualification"]]`
(age/qualification) or `r mean_rel[["Age/Qual/AFLB"]]` (age/qualification/AFLB).
Estimates might be further affected by weighting on other demographic
variables.[^weighting]

[^weighting]: To make interpretation straightforward, we do not use weights in 
results below unless noted. However, all our qualitative results are robust to
using weights.

We also check for balancing and diversifying selection by estimating
\@ref(eq:regression) with a quadratic term.[^balance-diverse-def] In particular,
we find diversifying selection for educational attainment polygenic scores: at
higher values of these scores, the negative effect on fertility is smaller
(Appendix Figure \@ref(fig:plot-purifying)).

[^balance-diverse-def]: Balancing selection selects for intermediate values, 
reducing the variance of the trait under selection. Diversifying selection
selects for extreme values and increases the variance.

```{r plot-res-wt, fig.cap = "Selection effects: weighted and unweighted regressions.  Each point represents a single bivariate regression of number of children on a polygenic score."}

n_regs <- as.double(length(score_names))

res_wt_combined %>% 
      mutate(
        Weights  = fct_relevel(Weights, "Geographical", "Age/Qualification",
                    "Age/Qual/AFLB", "None"),
        term     = pretty_names(term),
        term     = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      rename(p = p.value) %>% 
      # can't use standard_ggplot because we have non-standard geom_point()
      ggplot(aes(
        x      = estimate, 
        y      = term, 
        yend   = term,
        shape  = Weights, 
        fill   = Weights, 
        colour = p < 0.05/{{n_regs}}
      )) +
      my_vline +
      geom_point(size = 1.5, stroke = 0.5) +
      confint_geom_segment() +
      pval_scale() +
      scale_fill_manual(values = c(
        "None"              = "grey15", 
        "Geographical"      = "dodgerblue4",
        "Age/Qualification" = "steelblue2",
        "Age/Qual/AFLB"      = "darkseagreen4"
      )) +
      scale_shape_manual(values = c(
        "None" = "square filled", 
        "Geographical" = "circle filled",
        "Age/Qualification" = "circle filled", 
        "Age/Qual/AFLB" = "circle filled"
      )) + 
      labs(x = "Effect size", y = "") +
      scale_linetype_manual(name = "", values = c("95% c.i." = 1)) +
      my_theme()

```

To motivate our economic model, we split the sample by demographic
and social variables, and re-estimate \@ref(eq:regression) within each subgroup.
We examine socio-demographic variables including income, education and sex, and
family structure variables including age at first live birth, presence of a partner, and
lifetime number of sexual partners.
We emphasize that these categories are not necessarily exogenous to the polygenic
scores. For example, both in the data and in our theoretical model, age at first
live birth is a choice variable, which is endogenous to human-capital-related
polygenic scores (Appendix Figure \@ref(fig:plot-age-flb-dv)). Nevertheless,
differences in equation \@ref(eq:regression) across subgroups constrain the set
of possible explanations for natural selection. As we show below, a simple
economic model predicts the differences we find.

```{r plot-income-educ-level, fig.cap = "Selection effects by education and income.", fig.subcap = c("Age left full-time education", "Household income"), fig.ncol = 1}

drake::loadd(res_edu)
n_regs <- as.double(nrow(res_edu))

res_edu %>% 
      mutate(
        "Age left FTE" = fct_relevel(age_fte_cat, "< 16", "16-18", "> 18"),
        "Age left FTE" = add_n(`Age left FTE`, "age_fte_cat")
      ) %>% 
      standard_ggplot(n_regs = n_regs, fill_col = `Age left FTE`, 
                        fill_direction = -1)

drake::loadd(res_income)

n_regs <- as.double(nrow(res_income))
res_income %>% 
      mutate(
        Income = factor(income_cat, 
          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", "> £100K")
        ),
        Income = add_n(Income, "income_cat")
      ) %>% 
      standard_ggplot(n_regs = n_regs, fill_col = `Income`, 
                        fill_direction = -1)
```

```{r calc-income-edu-controlled}

drake::loadd(res_income_controlled, res_edu_controlled)
n_regs_ic <- nrow(res_income_controlled)
n_sig_ic <- sum(res_income_controlled$p.value < 0.05/n_regs_ic)

n_regs_educ <- nrow(res_edu_controlled)
n_sig_educ <- sum(res_edu_controlled$p.value < 0.05/n_regs_educ)
```

Figure \@ref(fig:plot-income-educ-level) plots selection effects for each
polygenic score, grouping respondents by age of completing full-time education,
and by household income. Effects are larger and more significant for the lowest
income category, and for the lowest education category.  These results could be
explained by age, if older respondents have lower income and are less educated,
and also show more natural selection on polygenic scores. However, when we rerun
the regressions, interacting the polygenic score with income category and also
with a quadratic in age, the interaction with income remains significant at
$p < 0.05/`r n_regs_ic`$ for `r n_sig_ic` out of `r n_regs_ic` regressions. Similarly
if we interact the PGS with age of leaving full time education and a quadratic
in age, the interaction with age leaving full time education remains significant
at $p < 0.05/`r n_regs_educ`$ for `r n_sig_educ` out of `r n_regs_educ` regressions.

Selection effects are also different between men and women (Appendix Figure
\@ref(fig:plot-sexes)). Differences are particularly large for educational
attainment, height and MDD. Several polygenic scores for mental illness and
personality traits are more selected for (or less against) among women,
including major depressive disorder (MDD), schizophrenia and neuroticism, while
extraversion is more selected for among men. Scores for waist circumference and
waist-hip ratio are less selected for among women. One
possible reason for these sex differences is that polygenic scores may affect
fertility via success in marriage markets, and men and women may value different
characteristics in these.

```{r calc-age-flb-cross}

drake::loadd(res_age_flb_cross)

cor_age_flb <- res_age_flb_cross %>% 
      mutate(category = gsub(":.*", "", term)) %>% 
      pivot_wider(score_name, names_from = category, values_from = estimate) %>%              {cor(.$`age_flb_cat10-22`, .$`age_flb_cat28-52`)}

```

We next focus on variables related to household type and reproductive
strategy. We split males and females by lifetime number of sexual
partners, at the median value of 3 (Figure \@ref(fig:plot-n-partners)a).
For both sexes, selection effects are larger and more significant among
those with more than 3 lifetime partners. Next we split respondents by whether
they were living with a spouse or partner at the time of interview.
Effects are larger among those not living with a spouse or partner
(Figure \@ref(fig:plot-n-partners)b).

Lastly, we split female respondents by age at first live birth
(AFLB).[^aflb-no-men] There is evidence for genetic effects on AFLB
[@Barban_2016], and there is a close link between this variable and number of
children born. Figure \@ref(fig:plot-age-flb) shows effect sizes estimated
separately for each tercile of AFLB. Several effects are strikingly different
across terciles. ADHD and MDD are selected for amongst the youngest third of
mothers, but selected against among the oldest two-thirds. Educational
attainment is selected for among the oldest two-thirds of mothers, but is not
significantly selected among the youngest third. Similarly, several polygenic
scores for body measurements are selected against only among older mothers. The
correlation between effect sizes for the youngest and oldest terciles is 
`r cor_age_flb`.

[^aflb-no-men]: This information is unavailable for men.

```{r plot-n-partners, fig.cap = "Selection effects by number of sexual partners and household type.", fig.subcap = c("Lifetime number of sexual partners", "Household type"), fig.ncol = 1, fig.align = "center", fig.height = 4}

sexes_theme <-  my_theme(
                  plot.margin          = unit(c(1, 1, 0, 1), "lines"),
                  legend.position      = "bottom", 
                  legend.box.spacing   = unit(0, "lines"),
                  legend.box.margin    = margin(-5, 0, 0, -85),
                  legend.margin        = margin(0, 0, 0, 0),
                  legend.box.just      = "left",
                  legend.justification = "left",
                  legend.spacing       = unit(1, "lines")
                )
sexes_guides <- guides(
                  # fill  = guide_legend(ncol = 1),
                  # color = guide_legend(ncol = 1)
                )

drake::loadd(res_partners)
res_partners %<>%       
      filter(grepl(":", term)) %>% 
      mutate(`N partners` = ifelse(grepl("TRUE:", term), "3 or less", "4 or more"))

n_regs <- as.double(nrow(res_partners))
res_partners %>% 
      mutate(
        score_name = pretty_names(score_name)
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(
        y      = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        yend   = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        x      = estimate, 
        fill   = `N partners`, 
        colour = p < 0.05/{{n_regs}}
      )) + 
        my_vline +
        confint_geom_segment() +
        pval_geom_point() +
        facet_wrap(vars(sex), scales = "free_y") +
        my_fill_scale(n = 2) +
        pval_scale() +
        sexes_theme +
        my_labs(x = "", y = "") + 
        scale_y_reordered() +
        scale_linetype_manual(name = "", values = c("95% c.i." = 1)) +
        sexes_guides

drake::loadd(res_with_partner_sex)
res_with_partner_sex %<>% filter(grepl(":", term))

n_regs <- as.double(nrow(res_with_partner_sex))

res_with_partner_sex %>% 
      mutate(
        score_name = pretty_names(score_name),
        Household = ifelse(grepl("TRUE", term), "With partner", 
                      "Without partner")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(
        y      = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        yend   = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        x      = estimate, 
        fill   = Household, 
        colour = p < 0.05/{{n_regs}}
      )) + 
        my_vline +
        confint_geom_segment() +
        pval_geom_point() +
        facet_wrap(vars(sex), scales = "free_y") +
        my_fill_scale(n = 2) +
        pval_scale() +
        sexes_theme +
        my_labs(x = "", y = "") + 
        scale_y_reordered() +
        scale_linetype_manual(name = "", values = c("95% c.i." = 1)) +
        sexes_guides

```

```{r calc-corr-age-flb}

drake::loadd(res_age_flb)
drake::loadd(res_sex)

res_sex_comparison <- res_sex %>% 
      filter(sex == "Female") %>% 
      mutate(score_name = term) %>% 
      select(score_name, term, estimate:conf.high)

res_age_flb %<>% filter(term != "age_flb")

res_combined <- bind_rows(
        raw      = res_sex_comparison, 
        with_age = res_age_flb %>% select(score_name, term, estimate:conf.high), 
        .id      = "reg.type"
      ) %>% 
      arrange(score_name)

raw_flb_comparison <- res_combined %>% 
      pivot_wider(
        id_cols     = score_name, 
        names_from  = reg.type, 
        values_from = estimate
      )

cor_raw_flb <- cor(raw_flb_comparison$raw, raw_flb_comparison$with_age)
opp_signed <- sum(
        sign(raw_flb_comparison$raw) != sign(raw_flb_comparison$with_age)
      )

```

To investigate this further, we estimate equation \@ref(eq:regression) among
females, *controlling* for AFLB. In `r opp_signed` out of 
`r length(score_names)` cases, effects change sign when controls are added. The
correlation between effect sizes controlling for AFLB, and raw effect sizes, is
`r cor_raw_flb`. Thus, selection effects seem to come through two opposing
channels: an effect on AFLB, and an opposite-signed effect on number of children
after AFLB is controlled for.

```{r plot-age-flb, fig.cap = "Selection effects by age at first live birth terciles (women only).", fig.align = "center"}

drake::loadd(res_age_flb_cross)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_cross))

res_age_flb_cross %>% 
      mutate(
        `Age at first live birth` = gsub("age_flb_cat(.*):.*", "\\1", term),
        `Age at first live birth` = add_n(`Age at first live birth`, "age_flb_cat")
      ) %>% 
      standard_ggplot(score_col = score_name, 
                      fill_col = `Age at first live birth`, n_regs = n_regs,
                      fill_direction = -1)

```

```{r calc-siblings-children}
sib_chn_cor <- cor.test(famhist$n_sibs, famhist$n_children, use = "complete")
stopifnot(sib_chn_cor$p.value < 2e-16)
```

We also examine selection effects among respondents' parents, using information
on respondents' number of siblings. Effect sizes of polygenic scores are highly
correlated across the two generations. Median-splitting respondents by year of
birth, we find little evidence of change in effect sizes among either the
parents' or the respondents' generation. We also check whether selection effects
vary by socio-economic status and AFLB in the parents' generation. We have no
information about parents' income, so we use the 1971 Townsend deprivation score
of respondents' birthplace as a proxy [@townsend1987deprivation]. Results show
the same pattern as for respondents. Effect sizes are larger and more often
significant in the most deprived areas (Appendix Figure
\@ref(fig:plot-siblings-townsend)). Effects are larger among younger parents,
and change sign when controlling for AFLB, among both mothers and fathers
(Appendix Figures \@ref(fig:plot-age-birth-parents-cross), 
\@ref(fig:plot-age-birth-parents)). Lastly, we check for a
"quantity-quality tradeoff" between parents' number of children and number of
grandchildren. We do not find any: in fact, the correlation between *number of
siblings* and *number of children* is positive ($\rho$ = 
`r sib_chn_cor$estimate`, $p < 2 \times 10^{-16}$).

\FloatBarrier

# Human capital and natural selection

These results show that selection effects are weaker, absent, or even reversed
among some subgroups of the population. A possible explanation for this comes 
from the economic theory of fertility 
[@becker1960economic; @willis1973new;@becker1976child]. According to this theory,
increases in a person's wage affect their fertility via two opposing channels.
There is an *income effect* by which children become more affordable, like any
other good. There is also a *substitution effect*: since childrearing has a cost
in time, the opportunity cost of childrearing increases if one's market wage is
higher. The income effect leads higher earners to have more children. The
substitution effect leads them to have fewer.

Suppose that certain genetic variants correlate with human capital -- skills or
other characteristics that affect an individual's earnings in the labour market
[@mincer1958investment; @becker1964human]. These variants may then be associated
with opposing effects on fertility. The income effect will lead to natural
selection in favour of earnings-increasing variants (or variants that are merely
associated with higher earnings). The substitution effect will do the reverse.

To show this, we write down a simple model of fertility choices. $h$ is
an individual's level of human capital. For now, we simply identify this with
his or her wage $W$. Raising a child takes time $b$. People maximize utility $U$
from the number of children $N$ and from income $Y\equiv(1-bN)W$:

\[
U = u(Y)+aN.
\]

Here $a$ captures the strength of preference for children. $u(\cdot)$ captures
the taste for income, and is increasing and concave. We treat $N$ as continuous,
in line with the literature: this can be thought of as the expected number of
children among similar people. The marginal benefit of an extra child is
$\frac{dU}{dN} = -bWu'(Y)+a$. The effect of an increase in human capital on this
marginal benefit is

\[
\frac{d^{2}U}{dNdh}=\underbrace{-bu'(Y)}_{\textrm{Substitution effect}}\underbrace{-bYu''(Y)}_{\textrm{Income effect}}.
\]

The *substitution effect* is negative and reflects that when wages increase,
time devoted to childcare costs more in foregone income. The positive *income
effect* depends on the curvature of the utility function, and reflects that when
income is higher, the marginal loss of income from children is less painful.

To examine education and fertility timing, we extend the model to
two periods. For convenience we ignore time discounting, and assume
that credit markets are imperfect so that agents cannot borrow. Write

\begin{equation}
U(N_{1},N_{2}) = u(Y_{1}) + u(Y_{2}) + aN_{1} + aN_{2}\label{eq:U}
\end{equation}

Instead of identifying human capital with wages, we now allow
individuals to choose a level of education $s \in [0,1]$, which has
a time cost in period 1. Education is
complementary to human capital $h > 0$, and increases period 2 wages, which take 
the  simple functional form $w(s,h) = sh$. We normalize period 1 wages
to 1. Letting $u(\cdot)$ take the constant relative risk aversion form
$u(y)=\frac{y^{1 - \sigma} - 1}{1 - \sigma}$, where $\sigma>0$ captures risk aversion
in income, we examine total fertility $N^{*} = N_{1}^{*} + N_{2}^{*}$ and the 
*fertility-human capital relationship*, $\frac{dN^{*}}{dh}$. For $\sigma < 1$ in
a neighbourhood of $\sigma = 1$, Table \@ref(tab:theory) shows five theoretical
predictions, along with our corresponding empirical results for the correlation
between polygenic scores and fertility.[^one-period-results] The key insight of
the model is that for middling levels of risk aversion, the substitution effect
dominates at low income levels, but is balanced by the income effect as income
increases.

[^one-period-results]: Predictions 1-3 also hold in the one-period model with
constant relative risk aversion. Our empirical results are actually stronger
than prediction 5, in that correlations with fertility are *reversed* at higher
AFLB. This prediction can be accommodated in the model if children have a money
cost as well as a time cost (Appendix Figure \@ref(fig:N-plot-with-m)).

Thus, a simple economic model can explain many of our results.
Other empirical work in economics also supports our mechanism.
@caucutt2002women and @monstad2008education show that education and skills
affect age at first birth and fertility. Income decreases fertility at low
income levels, but increases it at higher income levels [@cohen2013financial].
US fertility decreases faster with education among single mothers than married
mothers [@baudin2015fertility], in line with our prediction 3 and as predicted by
@becker1981treatise.

------------------------------------------------------------------------------------
    **Theory**                                        **Empirical results**
--- ------------------------------------------------- ------------------------------
1.  The fertility-human capital                       Figures \@ref(fig:plot-means-over-time) 
    relationship is negative:                         and \@ref(fig:plot-res-wt).
    $\frac{dN^{*}}{dh} < 0$.

2.  The fertility-human capital                       Figure \@ref(fig:plot-income-educ-level)a. 
    relationship is weaker (closer to zero)           Selection effects are also weaker at higher
    at higher wages and/or levels of                  polygenic scores for 
    human capital.                                    educational attainment (Appendix Figure
                                                      \@ref(fig:plot-purifying)).

3.  The relationship is more negative when            Stronger effects for single parents
    the time burden of children, $b$, is              (Figure \@ref(fig:plot-n-partners)).
    larger.

4.  The relationship is weaker at higher              Figure \@ref(fig:plot-income-educ-level)b.
    levels of education $s$.

5.  The relationship is weaker among those            Effects weaker among those starting 
    who start fertility in period 2                   fertility later 
    ($N_{1}^{*} = 0$) than among those who            (Figure \@ref(fig:plot-age-flb)).
    start fertility in period 1 
    ($N_{1}^{*} > 0$).
----------------------------------------------------------------------------------------

Table: (\#tab:theory) Predictions from the theoretical model and corresponding 
empirical results.

```{r literature, eval = FALSE}
"Other models of the literature

## LITERATURE NOTES

* Lundberg and Pollak JEP 2007:
    - several models where nonmarital fertility is more likely for poorer women
    - Burdett and Ermisch 2002, Willis 1999, Neal 2004
    - Unmarried parents may not be able to credibly commit to each other

* Happel et al. 1984
   - childbirth is chosen to minimize loss of earnings from interrupted work.
   - early childbirth if your skills decay a lot  while childrearing. (No point
     gaining them then losing them.) Late childbirth otherwise.
   - late childbirth if you have acquired skills before marriage.
   - Capital-market imperfections make it wise to delay childbirth (till man has
     high income to cope while wife is at home). More so for men's wages which
     rise faster.
   - empirically: (among married couples) 
     - high-skill occupation women have children later. Higher-earning men have 
     them earlier.

- Caucutt et al. 2001
  - high waged and high educated women (of recent cohorts) more likely to
    delay childbirth

- Cohen et al. 2013
  - analyse effect of Israeli subsidy
  - along the way, find that income affects fertility negatively at low income,
    positively at high income, supporting Becker and Tomes 76

- Becker and Tomes 1976
  - investment in children quality and quantity when quality must be equal
    across children; and some quality comes from an innate endowment.
  - "observed" quantity-of-children income elasticity rises as income rises
  - while "observed" quality-of-children elasticity falls as income rises, 
    b/c more of total quality must be paid for by parents themselves. 
  - the need to pay more for quality results in lower quantity, since total cost is 
    quantity X quality
  - something fishy about argument somehow...

- Lindo 2008
  - men who lose a job have children sooner, but fewer children in total
  - controls for all skills effects
  - so overall effect of income is positive, but timing changes

- Monstad et al. 2008
  - using an education reform to vary education
  - no evidence of a causal effect on fertility, but births start later
  * maybe the usual correlation reflects underlying skills. Your data could
    test this.

- Autor et al. 2017
  - the China shock (exogenous shock to income) 
  - they use Chinese import penetration by industry (instrumented using
    other countries) times industry employment shares; and separate it for
    males and females
  - shocks to male employment decreased fertility and marriage; 
    increased unmarried mothers
  - shocks to female employment *increase* marriage and motherhood
  * this roughly seems to be the consensus
  * Whereas, you find that male and female *skills* (as 'instrumented' by PGS)
    have the same effect on fertility - i.e. less skills = higher fertility

- Baudin et al. 2015
  * look at childlessness vs. fertility (extensive/intensive margin)
  * some people are too poor to afford children
  * 'a husband alleviates the time cost of having children'. But for some groups (see appendix)
    there is a +ve income/fertility gradient
  * fertility decreases with female educn due to subst effect
  * baby boom and bust in US: a decline in childlessness caused by poverty, then an increase
    caused by substn effect ('opportunity cost')
  * fertility decreases with educn both for single mothers and married women
    - slightly (25%) higher decrease per year of educn (Table 2) for single mothers

```

```{r plot-cors-earnings-educ, fig.cap="Selection effects by correlations with earnings and educational attainment. Each point represents one polygenic score."}

drake::loadd(res_all)
drake::loadd(res_age_flb)
drake::loadd(res_cor_educ)
drake::loadd(res_cor_income)

effect_size <-  res_all %>% 
                  filter(dep.var == "N children", reg.type == "controlled") %>% 
                  pull(estimate)

effect_size_flb <-  res_age_flb %>% 
                      filter(term != "age_flb") %>% 
                      pull(estimate)

dfr <- data.frame(
        PGS          = rep(score_names, 4),
        Correlation  = rep(c("Earnings", "Education"), each = 33, 2),
        cor          = rep(c(res_cor_income[,1], res_cor_educ[,1]), 2),
        effect_size  = c(rep(effect_size, 2), rep(effect_size_flb, 2)),
        Controls     = rep(c("None", 
                        "Age at first live birth"), each = 66)
      )

dfr %>% 
      mutate(
        Controls = fct_relevel(Controls, "None"),
        PGS      = ifelse(abs(effect_size) > 0.035, PGS, ""),
        PGS      = sub("(^\\w*?)_.*", "\\1", PGS)
      ) %>% 
      filter(
        Controls == "None"
      ) %>% 
      ggplot(aes(effect_size, cor, colour = Correlation)) + 
        geom_point() +
        geom_text(aes(label = PGS), size = 3, colour = "black", 
                    check_overlap = TRUE, nudge_x = 0.005) +
        facet_wrap( ~ Correlation, scales = "free") + 
        my_vline + 
        my_hline +
        labs(x = "Selection effect", y = "Correlation") +
        coord_cartesian(clip = "off") +
        scale_colour_brewer(type = "qual", palette = 2) +
        theme(
          legend.position = "none",
          panel.spacing = unit(20, "pt")
        )

```

# Testing the theory

We test this explanation in two ways. First, the economic theory predicts that
genetic variants will be selected for (or against) in proportion to their
correlation with human capital. Figure \@ref(fig:plot-cors-earnings-educ) plots
selection effects on each polygenic score against that score's correlation with
two measures of human capital: earnings in a respondent's first job, and
educational attainment. The relationships are strongly negative. Thus, human
capital appears to be relevant to natural selection. The negative relationship
suggests that substitution effects dominate income effects, which fits the known
negative association between income and fertility
[@becker1960economic;@jones2006economic]. The correlations reverse when we
control for age at first live birth, suggesting that within AFLB categories, the
income effect dominates. (See also the Supplementary Animations for the
uncorrected effects per AFLB group, which show the correlation reversing after
age 22.)

```{r}
drake::loadd(res_mediation)

res_mediation %<>% mutate(
                     prop = estimate_ind/estimate_total,
                     # 0.025 not 0.05 because two-sided
                     sig  = abs(statistic_ind) > qnorm(1 - 0.025/33)
                   )

n_pos_sig <- sum(res_mediation$sig & res_mediation$prop > 0)

median_prop <- median(res_mediation$prop[res_mediation$prop > 0 & 
                                           res_mediation$sig])

```

Second, we run a mediation analysis to directly test whether the correlation
between each polygenic score and fertility is mediated by educational attainment
(Appendix Table \@ref(tab:tbl-res-mediation)). For `r n_pos_sig` scores, the
indirect effect of the score on fertility *via* educational attainment is
significantly different from zero ($p < 0.05/33$). Among these scores, the
median proportion of the total effect explained by the indirect effect is 
`r percent(median_prop)`. The educational attainment variable is a relatively
crude measure of human capital: more accurate measures would likely
explain more of the total effect.

```{r calc-fe-fertility}
drake::loadd(res_fe_fertility)
drake::loadd(res_all)

educ_reduction <- with(res_fe_fertility, 
       estimate[Regression == "Controlled" & term == "EA3_excl_23andMe_UK"] / 
       estimate[Regression == "Raw" & term == "EA3_excl_23andMe_UK"]
     )
educ_reduction <- 1 - educ_reduction

res_all %<>% filter(dep.var == "N children", reg.type == "controlled")
res_fe_fertility %<>% filter(Regression == "Raw")
res_all %<>% left_join(res_fe_fertility, by = "term")

lm_pool_w <- tidy(lm(estimate.y ~ estimate.x, res_all))
coef_pool_w <- lm_pool_w[[2, "estimate"]]
se_pool_w <- lm_pool_w[[2, "std.error"]]

```

These results support our theory that natural selection on polygenic scores is
driven by their *correlation* with human capital. We also test whether polygenic
scores *cause* fertility by running within-siblings regressions (Appendix Figure
\@ref(fig:plot-res-fe-fertility)). With a reduced sample size, all
within-sibling effects are insignificant after Bonferroni correction. However,
effect sizes are positively correlated with effect sizes from the pooled model,
and about 50% smaller (regressing within-sibling on pooled effect sizes, $b$ =
`r format(coef_pool_w, digits = 3)`). This attenuation is in line with the ~50%
decrease in heritability in within-sibling GWASs on age at first birth and
educational attainment [@Howe2021.03.05.433935]. We see these results as
providing tentative evidence that polygenic scores cause fertility, with effects
being partly driven by correlations with environmental variation in human
capital. We also reran within-siblings regressions adding a control for
education. Most effect sizes barely change, suggesting that our measure of
education does not in general mediate differences in fertility among siblings.
However, for the educational attainment polygenic score (EA3), effect sizes are
reduced by `r percent(educ_reduction)`.

```{r calc-risk-attitude}

drake::loadd(res_risk_control)
drake::loadd(res_all)

n_risk_sig <- res_risk_control %>% 
                filter(term == 'f.2040.0.0') %>% 
                {sum(.$p.value <= .Machine$double.eps)}

n_pgs_sig <- res_risk_control %>% 
               filter(term == score_name) %>% 
               {sum(.$p.value <= 0.05/33)}

res_compare <- res_all %>% 
      filter(dep.var == "N children", reg.type == "controlled") %>% 
      left_join(res_risk_control %>% filter(term == score_name), 
                  by = "score_name")

ratio <- res_compare$estimate.y/res_compare$estimate.x
n_less_sig <- sum(res_compare$p.value.y <= 0.05/33 & res_compare$p.value.x > 0.05/33)
stopifnot(n_less_sig == 0)
```

We consider three alternative theories that might explain our results. First,
welfare benefits which incentivize child-bearing might be taken up more among
low-income individuals. However, most effect sizes appear unchanged over a large
span of twentieth-century history (Appendix Tables
\@ref(tab:tbl-siblings-by-period) and \@ref(tab:tbl-children-by-period)), during
which government spending on child-related benefits varied considerably
[@hoc1999socsec]. In general, there is only weak evidence that welfare benefits
affect fertility [@gauthier2007impact]. A second alternative theory is that
polygenic scores correlate with the motivation to have children, i.e. parameter
$a$ in the model [cf. @jones2008fertility]. This theory would not explain why
selection effects are smaller at higher incomes and education levels. (Indeed,
in the model, the effect of $a$ on fertility becomes stronger at higher levels
of human capital.) A third alternative is that traits under selection are linked
to externalizing behaviour and risk-seeking. This might be partially captured by
our risk-aversion parameter $\sigma$, which affects fertility by changing the
marginal utility of income, but a more direct channel is risky sexual behaviour
[@mills2021identification]. The data here provide some support for this: scores
which might plausibly be linked to externalizing behaviour, like ADHD and
younger age at smoking initiation, are selected for. However, this theory is
less good at explaining variation in selection across the full range of scores,
including physical measures such as waist-hip ratio and BMI. We test this theory
directly by re-estimating equation \@ref(eq:regression) controlling for a
measure of risk attitude (UK Biobank field 2040).  The median ratio of effect
sizes between regressions with and without controls is `r median(ratio)`; all
scores which are significant at $p < 0.05/33$ in uncontrolled regressions remain
so when controlling for risk attitude. This non-result could simply reflect the
imprecision of the risk attitude measure, which is a single yes/no question.
However, this measure *does* predict the overall number of children, highly
significantly ($p < 2 \times 10^{-16}$ in `r n_risk_sig` out of 33 regressions).
Given that, and the statistical power we get from our sample size, we believe
that the non-result is real: while risk attitude does predict fertility in the
sample, it is not an important channel for natural selection.

# Discussion

Previous work has documented natural selection in modern populations on variants
underlying polygenic traits [@beauchamp2016genetic; @kong2017selection;
@sanjak2018evidence]. We show that correlations between polygenic scores and
fertility are highly concentrated among specific subgroups of the population,
including people with lower income, lower education, younger first parenthood,
and more lifetime sexual partners. Among mothers aged 22+, selection effects
are reversed. Furthermore, the size of selection effects on a polygenic score
correlates with that score's association with labour market earnings. The
economic theory of fertility provides a parsimonious explanation for these findings.
Because of the substitution effect of earnings on fertility, scores are selected
for when they correlate with low human capital, and this effect is stronger at 
lower levels of income and education.

Polygenic scores which correlate with high (low) earnings and more (less)
education are being selected against (selected for). In addition, many of the
phenotypes under positive selection are linked to disease risk. Many people
would probably prefer to have high educational attainment, a low risk of ADHD
and major depressive disorder, and a low risk of coronary artery disease, but
natural selection is pushing against genes associated with these traits.
Potentially, this could increase the health burden on modern populations, but
that depends on effect sizes. Our results suggest that naïve estimates can be
affected by sample ascertainment bias. This problem may be less serious in
surveys which aim to be representative (as the UK Biobank does not). However,
there is still scope for bias, since not all respondents consent to the
collection of genetic data. For instance, completion rates for genotype data in
the US Health and Retirement Study were around 80-85% [@hrs2020]. Researchers
should be aware of the risks of ascertainment when studying modern natural
selection.

We also do not know how estimated effect sizes of natural selection will change
as more accurate measures of genetic variation are produced, or whether genetic
variants underlying other phenotypes will show a similar pattern to those
studied here. In addition, effects of polygenic scores may be inflated in
population-based samples, because of indirect genetic effects, gene-environment
correlations, and/or assortative mating [@lee2018gene;
@selzam2019comparing; @kong2018nature; @Howe2021.03.05.433935], although we do
not expect that this should impact their association with number of offspring,
or the resulting changes in allele frequencies. In short, it is probably too
early to tell whether modern natural selection has a substantively important
effect on population means of phenotypes under selection. Nevertheless, we note
that selection effects on our measured polygenic scores are still relatively
small, even after reweighting to account for ascertainment bias.

```{r calc-inequality}
drake::loadd(res_ineq)

n_increases <- sum(res_ineq$ratio > 1)
median_pct_change <- (median(res_ineq$ratio) - 1) * 100

cor_ea3 <- res_ineq[res_ineq$score == "EA3_excl_23andMe_UK", ]
cor_ea3$increase_pct <- (cor_ea3$ratio - 1) * 100

drake::loadd(res_ineq_ea3)

r2_income_true_psea <- scales::percent(res_ineq_ea3$r2_income_true_psea, 
                                         accuracy = 0.1)
r2_income_true_psea_wt <- scales::percent(res_ineq_ea3$r2_income_true_psea_wt, 
                                            accuracy = 0.1)
```

Because selection effects are concentrated in lower-income groups, they may also
increase inequality with respect to polygenic scores. For example, Figure
\@ref(fig:plot-mean-EA3-by-income) graphs mean polygenic scores for educational
attainment (EA3) among children from households of different income groups. The
blue bars show the actual means, i.e. parents' mean polygenic score weighted by
number of children. The grey bars show the hypothetical means if all households
had equal numbers of children. Natural selection against genes associated with
educational attainment, which lowers the mean, is stronger at the bottom of the
income distribution, and this increases the differences between groups.
Overall, natural selection increases the correlation of polygenic scores with
income for `r n_increases` out of 33 polygenic scores, with a median
percentage increase of `r median_pct_change`% (Appendix Table
\@ref(tab:tbl-inequality-effects)).

If inequalities in polygenic scores are important for understanding social
structure and mobility [@belsky2018genetic;
@Rimfeld_2018; @harden2021genetic], then these increases are substantive. A
back-of-the-envelope calculation, using estimated heritability for educational
attainment of 0.4 [@branigan2013variation], gives the proportion of variance in
household income attributable to the "true" polygenic score for educational
attainment as `r r2_income_true_psea` among respondents. Natural selection would
raise this proportion to `r r2_income_true_psea_wt` among respondents' children,
counterfactually holding environmental conditions constant. Similarly, since
many polygenic scores are predictive of disease risk, they could potentially
increase health inequalities. In general, the evolutionary history of
anatomically modern humans is related to disease risk [@benton2021influence];
understanding the role of contemporary natural selection may aid in mapping the
genetic architecture of current health disparities.

```{r plot-mean-EA3-by-income, fig.align = "center", fig.cap = "Mean polygenic score for educational attainment (EA3) of children by household income group. Blue is actual. Grey is hypothetical in the absence of selection effects. Respondents weighted by Age/Qualification.", fig.width = 4}
income_EA3 <- famhist %>% 
      filter(! is.na(n_children), ! is.na(income_cat)) %>% 
      mutate(
        `Household income` = factor(income_cat, 
                          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", 
                                      "> £100K"))
      ) %>% 
      group_by(`Household income`) %>% 
      summarize(
        `Without selection` = mean(EA3_excl_23andMe_UK, na.rm = TRUE),
        Actual              = weighted.mean(EA3_excl_23andMe_UK, n_children, 
                                              na.rm = TRUE)
      ) %>% 
      tidyr::pivot_longer(-`Household income`) %>% 
      rename(
        `Children's mean EA3` = value
      )

income_EA3 %>% 
              ggplot(aes(`Household income`, `Children's mean EA3`, fill = name)) +
                geom_col(position = "dodge") +
                scale_fill_manual(values = c("navy", "grey60")) +
                theme(
                  legend.background = element_rect(fill = "white", colour = NA),
                  legend.title = element_blank(), 
                  legend.position = c(.25, .85)
                )
```

Existing evidence on human natural selection has led some to "biocosmic
pessimism" [@sarraf2019modernity]. Others are more sanguine, and argue that
natural selection's effects are outweighed by environmental improvements, such
as those underlying the Flynn effect [@flynn1987massive]. The evidence
here may add some nuance to this debate. Patterns of natural selection have been
relatively consistent across the past two generations, but they are not the
outcome of a universal, society-wide phenomenon. Instead they result from 
opposing forces, operating in different parts of society and pulling in
different directions.

Any model of fertility is implicitly a model of natural selection, but so far,
the economic and human genetics literatures have developed in parallel.
Integrating the two could deepen our understanding of natural selection in
modern societies. Economics possesses a range of theoretical models on the
effects of skills, education and income [see @hotz1997economics;
@lundberg2007american]. One perennial problem is how to test these theories in a
world where education, labour and marriage markets all interact. Genetic data,
such as polygenic scores, could help to pin down the direction of causality,
for example via Mendelian randomization [@davey2003mendelian].
Conversely, theory and empirical results from economics can shine a light on the
mechanisms behind natural selection, and thereby on the nature of individual
differences in complex traits and disease risk.

# Materials and methods

We use participant data from UK Biobank [@bycroft2018uk], which has received
ethical approval from the National Health Service North West Centre for Research
Ethics Committee (reference: 11/NW/0382). We limit the sample to white British
participants of European descent, as defined by genetic estimated ancestry and
self-identified ethnic group, giving a sample size of
`r comma_num(nrow(famhist))`. For regressions on number of children we use
participants over 45, since most fertility is completed by this age. This gives
a sample size of `r comma_num(nrow(famhist %>% filter(kids_ss)))`.

Polygenic scores were chosen so as to cover a reasonably broad range of traits,
and based on the availability of a large and powerful GWAS which did not include
UK Biobank. Scores were computed by summing the alleles across ~1.3 million
genetic variants weighted by their effect sizes as estimated in 33 genome-wide
association studies (GWASs) that excluded UK Biobank. To control for population
stratification, we corrected the polygenic scores for 100 principal components
(PCs).

To compute polygenic scores and PCs, the same procedures were followed as 
described in @abdellaoui2019genetic.

Earnings in first job are estimated from mean earnings in the 2007 Annual Survey
of Hours and Earnings, using the SOC 2000 job code (Biobank field 22617).

Population data for weighting is taken from the 2011 UK Census and the 2006
General Household Survey (GHS). Weighting for Age/Qualification and 
Age/Qualification/AFLB weights was done using marginal totals from a linear
model, using the `calibrate()` function in
the R "survey" package [@lumley2020survey]. Geographical weighting was done with
iterative post-stratification using the `rake()` function, on Census Middle Layer
Super Output Areas, sex and presence/absence of a partner.

## Acknowledgements

AA is supported by the Foundation Volksbond Rotterdam and by ZonMw grant 849200011 from The Netherlands Organisation for Health Research and Development. This study was conducted using UK Biobank resources under application numbers 40310 and 19127.

\FloatBarrier
\clearpage
# Appendix

\localtableofcontents
\clearpage

## Natural selection by sex

```{r plot-sexes, fig.cap = "Selection effects by sex. Highlighted lines are significant differences at p < 0.05/33. Highlighted points are significantly different from 0 at p < 0.05/66."}

drake::loadd(res_sex)
n_regs <- as.double(nrow(res_sex))

res_sex_wide <- res_sex %>% 
      select(term, estimate, sex, diff.p.value) %>% 
      pivot_wider(names_from = "sex", values_from = "estimate") %>% 
      mutate(
        min_est = pmin(Female, Male),
        max_est = pmax(Female, Male),
        term = pretty_names(term),
        term = fct_reorder(term, Female)
      ) 
n_comps <-as.double(nrow(res_sex_wide))

res_sex %>% 
      mutate(
        term = pretty_names(term),
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, fill = sex, 
            colour = p < 0.05/{{n_regs}})) +
        my_vline +
        geom_linerange(data = res_sex_wide, 
          aes(
            y      = term,
            x      = NULL,
            fill   = NULL,
            xmin   = min_est, 
            xmax   = max_est, 
            colour = diff.p.value < 0.05/{{n_comps}}
          ), 
        group = "") +
        pval_geom_point() +
        pval_scale(name = "p < 0.05 (corrected)") +
        my_fill_scale(n = 2) +
        my_labs() +
        my_theme()

```

\FloatBarrier
\clearpage
## Weighted regressions

Table \@ref(tab:tbl-res-wt) gives effect sizes as a proportion of the unweighted
effect size, for all polygenic scores which are consistently signed and which
are significantly different from zero in unweighted regressions.

```{r tbl-res-wt}

rel_sizes %>% 
      arrange(desc(Geographical)) %>% 
      bind_rows(rel_sizes_summary) %>% 
      rename(PGS = term) %>% 
      as_hux() %>% 
      insert_row("", "Weighting", fill = "") %>% 
      set_caption(
        "Weighted effect sizes as a proportion of unweighted effect sizes."
      ) %>% 
      merge_cells(1, 2:4) %>% 
      set_align(1, 2, "centre") %>% 
      set_header_rows(1, TRUE) %>% 
      theme_compact() %>% 
      set_number_format(-1, -1, 2) %>% 
      style_headers(bold = TRUE) %>% 
      set_bottom_border(1, 2:4) %>% 
      set_bottom_border(2, everywhere) %>% 
      set_italic(final(2), 1) %>% 
      set_font_size(8) %>% 
      set_width(0.7) %>% # ensures line wrapping below
      set_col_width(c(.34, .22, .22, .22)) %>% 
      add_footnote(paste0(
        "Only consistently-signed and significant (when unweighted) ",
        "estimates are shown. ",
        "Age/Qual/AFLB as a proportion of unweighted regressions ",
        "including females only."),
        top_padding  = 0,
        left_padding = 0
      )

```

\FloatBarrier
\clearpage
## Balancing and diversifying selection

Balancing selection reduces variance in the trait under selection, while
diversifying selection increases variance. To check for these, we rerun equation
\@ref(eq:regression), adding a quadratic term in $PGS_i$, and using our
age/qualification weights. Scores for hip circumference show significant
balancing selection ($p < 0.05/33$, negative coefficient on quadratic term).
Scores for educational attainment (EA2 and EA3) show significant diversifying
selection ($p < 0.05/33$, positive coefficient), which reduces the strength of
selection against educational attainment at very high levels of the PGS. Figure
\@ref(fig:plot-purifying) plots predicted number of children against polygenic
score from these regressions.

We also checked for balancing selection in the parents' generation, using
age/qualification weights multiplied by the inverse of number of siblings.
Scores for EA2 and EA3 again show significant diversifying selection ($p <
0.05/33$, positive coefficient). Other scores including hip circumference were 
not significant.

```{r plot-purifying, fig.align = "center", fig.cap = "Purifying/diversifying selection: predicted number of children by polygenic score.", fig.height = 3}

drake::loadd(res_quadratic)
sig <- res_quadratic %>% 
         filter(grepl("^2", term, fixed = TRUE)) %>% 
         filter(p.value < 0.05/33) %>% 
         pull(score_name)

stopifnot(sig == c("EA2_noUKB", "EA3_excl_23andMe_UK", "hip_combined"))

sig_all_terms <- res_quadratic %>% filter(score_name %in% sig)

drake::loadd(res_sibs_quadratic)
sig_sibs <- res_sibs_quadratic %>% 
         filter(grepl("^2", term, fixed = TRUE)) %>% 
         filter(p.value < 0.05/33) %>% 
         pull(score_name)
stopifnot(sig_sibs == c("EA2_noUKB", "EA3_excl_23andMe_UK"))

op <- par(mfrow = c(1, 3))
for (score in sig) {
  coef <- sig_all_terms %>% filter(score_name == score) %>% pull(estimate)
  curve(
         coef[1] + coef[2] * x + coef[3] * x^2,
         xlim = c(-4, 4), 
         xlab = pretty_names(score),
         ylab = "Predicted N children"
       )
}
par(op)

```

\FloatBarrier
\clearpage
## Parents' generation {#sec-parents}
### Selection effects

```{r calc-res-sibs}
drake::loadd(res_wt_age_qual_weights)
drake::loadd(res_children_comparison)

cmp <- left_join(res_wt_age_qual_weights, res_children_comparison, by = "term")

cor_inc_vs_excl_childless <- cor(cmp$estimate.x, cmp$estimate.y)
n_smaller <- sum(abs(cmp$estimate.y) < abs(cmp$estimate.x))

pct_diff <- abs(cmp$estimate.y)/abs(cmp$estimate.x) - 1
median_pct_diff <- median(pct_diff)

```

The UK Biobank data contains information on respondents' number of
siblings, i.e. their parents' number of children. Since respondents' polygenic 
scores are equal in expectation
to the mean scores of their parents, we can use this to look at selection 
effects in the parents' generation. We estimate equation (\ref{eq:regression})
using *number of siblings* (including the respondent) as the dependent variable.
The parents' generation has an additional source of ascertainment bias: sampling
parents of respondents overweights parents who have many children. For instance,
parents of three children will have, on average, three times more children
represented in UK Biobank than parents of one child. Parents of no children will
by definition not be represented. To compensate, we reweight our preferred
weightings (Age/Qualification) by the inverse of *number of siblings*.

Figure \@ref(fig:plot-res-sibs) shows regressions of *number of siblings* on
polygenic scores. For a clean comparison with the respondents' generation, we
rerun regressions on *number of children* excluding those with no children, and
show results in the figure. Selection effects are highly correlated across the
two generations, and most share the same sign. Absolute effect size estimates
are larger for the parents' generation. We treat this result cautiously, because
when we split respondents up by year of birth, we find few differences in effect
sizes between early- and late-born respondents, for either generation. In other
words, since estimated effect sizes change across "generations", but do not
change over time within either generation, the change may be due to remaining
ascertainment bias within the sample or other effects. In particular, effect
sizes in both generations may depend on polygenic scores' correlation with
childlessness, and we cannot estimate this for the parents' generation.

Excluding childless people in the parents' generation could bias our estimates.
To learn about this, we compare effect sizes excluding and including childless
people in the *current* generation. The correlation
between the two sets of effect sizes is `r cor_inc_vs_excl_childless`. So, patterns across
different scores are broadly similar whether the childless are counted or not. 
However, absolute effect sizes are smaller when the
childless are excluded, for `r n_smaller` out of 33 scores; the median
percentage change is `r percent(median_pct_diff)`.

```{r plot-res-sibs, fig.cap = "Selection effects, respondents' parents vs. respondents. Age/qualification weights. Parental generation weights multiplied by 1/number of siblings. Respondents' regression excludes childless respondents."}
drake::loadd(res_sibs_parent_weights)
drake::loadd(res_children_comparison)

res_sibs_kids <-  bind_rows(
                    Siblings = res_sibs_parent_weights, 
                    Children = res_children_comparison,
                    .id = "Dep. var."
                  )

standard_ggplot(res_sibs_kids, `Dep. var.`, n_regs = length(score_names))
```

```{r calc-period-regs}

drake::loadd(res_period_parents)
drake::loadd(res_period_children)

res_period <- bind_rows(res_period_parents, res_period_children)

res_period %<>% 
      tidyr::extract(term, c("period", "score_name"),
        regex="year_split(.*):(.*)")
n_tests <- nrow(res_period)/2

summary_period <- res_period %>% 
      select(-std.error, -statistic) %>% 
      pivot_wider(
        names_from  = period, 
        values_from = c(estimate, conf.low, conf.high, p.value)
      ) %>% 
      group_by(score_name, children) %>% 
      mutate(
        significant = diff.p.value < 0.05/n_tests,
        sign_early  = sign(estimate_early),
        sign_late   = sign(estimate_late),
        size_inc    = abs(estimate_early) < abs(estimate_late),
        change_sign = sign_early != sign_late,
        Change      = case_when(
                        ! significant ~ "Insignificant",
                        change_sign   ~ "Change sign",
                        size_inc      ~ "Size increasing",
                        ! size_inc    ~ "Size decreasing"
                      )
      ) %>% 
      select(score_name, children, Change, significant, change_sign, 
        sign_early, sign_late) %>% 
      mutate(children = ifelse(children, "Respondents", "Parents")) %>% 
      pivot_wider(names_from = children, 
        values_from = c(Change, significant, change_sign, sign_early, 
        sign_late)) %>% 
      mutate(overall = case_when(
              ! change_sign_Parents & ! change_sign_Respondents & 
                sign_late_Parents == sign_early_Parents ~ 
                "Consistent",
              TRUE ~ "Changes direction"
            ))

stopifnot(sum(summary_period$Change_Parents == "Insignificant") == 33)
stopifnot(sum(summary_period$Change_Respondents == "Size increasing") == 4)
stopifnot(sum(summary_period$Change_Respondents == "Change sign") == 1)
stopifnot(sum(summary_period$Change_Respondents == "Insignificant") == 28)
```

The fact that childless people have such a strong effect on estimates makes it
hard to compare total effect sizes across generations. In particular, since the
parents' generation has a different distribution of numbers of children,
childless people may have had more or less effect in that generation. Another
issue is that we are estimating parents' polygenic scores by the
scores of their children. This introduces noise into our independent
variable, which might lead to errors-in-variables and bias coefficients towards
zero.

As an alternative approach, we run regressions interacting polygenic scores with
birth year, median split at 1950 ("early born" versus "late born"). We use both
*number of children* and *number of siblings* as a dependent variable. We weight
using age/qualification cells, and further adjust for selection in the parents'
generation (see above).

Tables \@ref(tab:tbl-siblings-by-period) and
\@ref(tab:tbl-children-by-period) summarize the results. There is no
evidence for changes in selection effects within the parents' generation. In the
respondents' generation, effect sizes were significantly larger in absolute size 
among the later-born for four polygenic scores: cognitive ability, EA2, EA3 and
extraversion. These changes are inconsistent with the intergenerational change, 
where estimated effect sizes were larger among the earlier, parents' generation.
One score, conscientiousness, showed a significant change in sign, from 
negative to positive effects on fertility.

Overall, there is weak evidence for change over time. The clearest results are
that (a) the direction of selection, and (b) the pattern of relative effect sizes
across scores, are broadly consistent over time.

```{r tbl-siblings-by-period}

period_table <- function (tbl, caption) {
  tbl %>% 
      as_hux() %>%
      set_bold(1, everywhere) %>% 
      set_bottom_border(1, everywhere) %>%

      add_footnote(glue("Significance is measured at p < 0.05/{n_tests}"), 
        border = 0.4, font_size = 8) %>% 
      set_contents(1, everywhere, 
        c("Change", "Number of scores")) %>% 
      set_right_border(final(1), 1, 0) %>% 
      set_top_padding(1) %>% 
      set_bottom_padding(1) %>%
      set_left_padding(everywhere, 1, 0) %>% 
      set_right_padding(everywhere, 2, 0) %>% 
      set_caption(caption) %>% 
      set_caption_width(1)
}

summary_period %>% 
      group_by(Change_Parents) %>% 
      count() %>% 
      period_table("Change in selection effects between parents of early and late born respondents (regressions on number of siblings). Weights: Age/Qualification times 1/number of siblings.")

```

```{r tbl-children-by-period}
summary_period %>% 
      group_by(Change_Respondents) %>% 
      count() %>% 
      period_table("Change in selection effects between early and late born respondents (regressions on number of children). Weights: Age/Qualification.")
```

\clearpage
\FloatBarrier
### Area deprivation

Figure \@ref(fig:plot-siblings-townsend) plots effects on *number of siblings* by
Townsend deprivation quintile of birth area.

```{r plot-siblings-townsend, fig.cap = "Selection effects in the parents' generation by Townsend deprivation quintile of birth area. Higher = more deprived. Estimates weighted by 1/number of siblings.", fig.align = "center"}

drake::loadd(res_townsend_parents)
n_regs <- nrow(res_townsend_parents)

res_townsend_parents %>% 
    mutate(
        n = prettyNum(n, big.mark = ","),
        `Townsend quintile` = paste0(quintile, " (N = ", n, ")")
      ) %>% 
  standard_ggplot(fill_col    = `Townsend quintile`,
                    n_regs    = n_regs, 
                    score_col = score_name, 
                    order_idx = 5
                  )

```

\clearpage
\FloatBarrier
### Age at first live birth

```{r calc-age-birth-parents}

drake::loadd(res_age_birth_parents)
drake::loadd(res_all)

res_sibs_unc <- filter(res_all, dep.var == "N siblings", reg.type == "controlled")
compare_abp_unc <- res_age_birth_parents %>% 
      filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
      left_join(res_sibs_unc, by = "term")

cor_unc_fath <- with(
  compare_abp_unc %>% filter(control == "fath_age_birth"), 
  cor(estimate.x, estimate.y))
cor_unc_moth <- with(
  compare_abp_unc %>% filter(control == "moth_age_birth"), 
  cor(estimate.x, estimate.y))
```

Among the parents' generation, we can control for age at first live birth using
the subsets of respondents who reported their mother's or father's age, and who
had no elder siblings. We run regressions on *number of siblings* on these
subsets. Figure \@ref(fig:plot-age-birth-parents-cross) shows selection effects
by terciles of age at first live birth, for mothers and fathers. As in the
respondents' generation, effect sizes are smaller, or even oppositely signed, for older parents. Importantly, this holds for both sexes.

```{r plot-age-birth-parents-cross, fig.cap = "Selection effects (parents' generation) by age at first live birth terciles.  Estimates weighted by 1/number of siblings.", fig.subcap = c("Mothers", "Fathers"), fig.ncol = 1}

drake::loadd(res_age_flb_mothers_cross)
drake::loadd(res_age_flb_fathers_cross)

famhist_bo1 <- famhist %>% 
                 filter(birth_order == 1) %>%
                 select(n_sibs, whr_combined, moth_age_birth_cat, 
                          fath_age_birth_cat)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_mothers_cross))

res_age_flb_mothers_cross %>% 
      mutate(
        `Age at first live birth` = gsub("moth_age_birth_cat(.*):.*", "\\1", term),
        `Age at first live birth` = add_n(`Age at first live birth`, 
                                          "moth_age_birth_cat", dv = "n_sibs",
                                          data = famhist_bo1)
      ) %>% 
      standard_ggplot(score_col = score_name, 
                      fill_col = `Age at first live birth`, n_regs = n_regs,
                      fill_direction = -1)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_fathers_cross))

res_age_flb_fathers_cross %>% 
      mutate(
        `Age at first live birth` = gsub("fath_age_birth_cat(.*):.*", "\\1", term),
        `Age at first live birth` = add_n(`Age at first live birth`, 
                                          "fath_age_birth_cat", dv = "n_sibs",
                                          data = famhist_bo1)
      ) %>% 
      standard_ggplot(score_col = score_name, 
                      fill_col = `Age at first live birth`, n_regs = n_regs,
                      fill_direction = -1)

rm(famhist_bo1)

```

Figure \@ref(fig:plot-age-birth-parents) shows the regressions controlling for
either parent's age at their birth. Effect sizes are very similar, whether
controlling for father's or mother's age. As in the respondents' generation,
effect sizes are negatively correlated with the effect sizes from bivariate
regressions without the age at birth control (father's age at birth: $\rho$ 
`r cor_unc_fath`; mother's age at birth: $\rho$ `r cor_unc_moth`).

```{r plot-age-birth-parents, fig.cap = "Selection effects (parents' generation) among eldest siblings, controlling for parents' age at birth. Estimates weighted by 1/number of siblings."}

n_regs <- as.double(nrow(res_age_birth_parents)/2)

res_age_birth_parents %>% 
    filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
    mutate(
      Control = ifelse(control == "fath_age_birth", 
                  "Father's age at birth", 
                  "Mother's age at birth"
                )
    ) %>% 
    standard_ggplot(fill_col = Control, n_regs = n_regs, order_idx = 2)

```

\clearpage
\FloatBarrier
## Effects of polygenic scores on age at first live birth

```{r calc-age-flb-dv}

drake::loadd(res_age_flb_dv)
drake::loadd(res_age_birth_parents_dv)

compare_flb_dv <- left_join(
        res_age_flb_dv           %>% select(term, estimate),
        res_age_birth_parents_dv %>% select(term, estimate, dep.var),
        by = "term"
      )

cor_flb_fath <- with(compare_flb_dv %>% filter(dep.var == "fath_age_birth"),
        cor(estimate.x, estimate.y)
      )
cor_flb_moth <- with(compare_flb_dv %>% filter(dep.var == "moth_age_birth"),
        cor(estimate.x, estimate.y)
      )
```

Our results suggest that polygenic scores may directly correlate with age at
first live birth. Figure \@ref(fig:plot-age-flb-dv) plots estimated effect sizes
from bivariate regressions for respondents. Figure
\@ref(fig:plot-age-birth-parents-dv) does the same for their parents, using only
eldest siblings.[^only-eldest] Effect sizes are reasonably large. They are also
highly correlated across generations. Effect sizes of polygenic scores on
father's age at own birth, and on own age at first live birth, have a
correlation of `r cor_flb_fath`; for mother's age and own age it is 
`r cor_flb_moth`.

[^only-eldest]: Parental AFLB can only be calculated for this group.

```{r plot-age-flb-dv, fig.cap = "Effects of polygenic scores on age at first live birth."}

n_regs <- as.double(nrow(res_age_flb_dv))

res_age_flb_dv %>% standard_ggplot(n_regs = n_regs)

```

```{r plot-age-birth-parents-dv, fig.cap = "Effects of polygenic scores on parents' age at respondent's birth, eldest siblings. Estimates weighted by 1/number of siblings."}

n_regs <- as.double(nrow(res_age_birth_parents_dv))

res_age_birth_parents_dv %>% 
      mutate(
        "Dependent variable" = dplyr::recode(dep.var, 
          "fath_age_birth" = "Father's AFLB", 
          "moth_age_birth" = "Mother's AFLB"
        )
      ) %>% 
      standard_ggplot(fill_col = `Dependent variable`, n_regs = n_regs)

```

\FloatBarrier
\clearpage
## Mediation analysis

We run a standard mediation analysis in the framework of @baron1986moderator. We estimate 
\begin{align}
N_i & = \alpha + \beta PGS_i + \gamma EA_i + X_i\mu + \varepsilon_i \\
EA_i & = \delta + \zeta PGS_i + X_i\mu + \eta_i
\end{align}
where $N_i$ is fertility, $PGS_i$ is each polygenic score, $EA_i$ is educational
attainment (age of leaving fulltime education), and $X_i$ is a vector of
controls. The total effect of $PGS$ on number of children is $\beta + \gamma
\zeta$. The "indirect effect" mediated by $EA$ is $\gamma \zeta$. The standard
error of the indirect effect can be calculated as

\[
\sqrt{\hat\gamma^2 \hat\sigma_\zeta^2 + \hat\zeta^2\hat\sigma_\gamma^2}
\]

where $\hat\sigma_\zeta$ is the standard error of $\hat\zeta$, etc.

We are not concerned about endogeneity of polygenic scores, since our theory
requires only that these correlate with human capital. However, endogeneity of
educational attainment is a concern, i.e., educational attainment might
correlate with unobserved characteristics that affect one's fertility (and which
are not related to human capital). To partially address this, we include
controls for age, sex, fluid IQ, height and risk attitude in $X$.

```{r tbl-res-mediation}

drake::loadd(res_mediation)

res_mediation %<>% mutate(
                     prop = estimate_ind/estimate_total,
                     # 0.025 not 0.05 because two-sided
                     sig  = abs(statistic_ind) > qnorm(1 - 0.025/33),
                     star = ifelse(sig, "*", "")
                   )

n_pos_sig <- sum(res_mediation$sig & res_mediation$prop > 0)

res_mediation %>%
      arrange(desc(estimate_total)) %>%
      transmute(
        PGS = pretty_names(term),
        `Total effect` = estimate_total,
        `Indirect effect` = estimate_ind,
        `95% c.i.` = glue::glue(
                     "[{estimate_ind - 1.96 * se_ind}, {estimate_ind + 1.96 * se_ind}]"
                     ),
        `Indirect/total` = prop,
        ` ` = star
      ) %>%
      as_hux() %>%
      set_align(everywhere, -1, "right") %>% 
      set_number_format(-1, 2:4, 4) %>% 
      set_number_format(-1, 5, fmt_percent()) %>%
      set_caption("Mediation analysis") %>% 
      theme_compact() %>% 
      set_width(0.9) %>%
      set_col_width(c(.3, .125, .125, .3, .1, .05)) %>% 
      set_bottom_border(final(), everywhere) %>%
      add_footnote("* indirect effect p < 0.05/33", font_size = 8)
```

Table \@ref(tab:tbl-res-mediation) shows results. For `r n_pos_sig` scores, the
indirect effect on fertility via human capital is significantly greater than 0 
at $p = 0.05/33$.

\FloatBarrier
\clearpage
## Within-siblings regressions

```{r plot-res-fe-fertility, fig.cap = "Selection effects controlling for sibling-group fixed effects, with and without a control for education (left education before 16, 16-18, or after 18). Each set of results is from a single regression of \\emph{n children} on 29 polygenic scores. Standard errors clustered by sibling group."}

drake::loadd(res_fe_fertility)
n_regs <- nrow(res_fe_fertility)/2 # Raw and Controlled treated separately

standard_ggplot(res_fe_fertility, fill_col = Regression, n_regs = n_regs,
                  fill_direction = -1)

n_groups <- attr(res_fe_fertility, "n_groups")
n <- attr(res_fe_fertility, "n")

```

We run a single regression on 29 polygenic scores within `r n_groups` sibling
groups (N = `r n`). Thus, we control both for environmental confounds (since
scores are randomly allocated within sib-groups by meiosis), and for genetic
confounds captured by our polygenic scores. We remove four scores which
correlate highly with other scores (educational attainment 2, hip circumference,
waist circumference and waist-hip ratio). Figure \@ref(fig:plot-res-fe-fertility)
shows the results.

\clearpage
\FloatBarrier
## Effects on inequality

Table \@ref(tab:tbl-inequality-effects) shows correlations between children's
polygenic scores and household income (UKB data field 738). Column "With
selection" uses respondents' scores weighted by Age/Qualification, times number
of children. Column "Without selection" uses scores weighted by
Age/Qualification only, i.e. the counterfactual correlation if all respondents
had the same number of children.

```{r tbl-inequality-effects}

res_ineq %>% 
      select(
        PGS                      = score, 
        `Cor. with selection`    = actual, 
        `Cor. without selection` = cf,
        `Ratio`                  = ratio
      ) %>% 
      mutate(
        PGS = pretty_names(PGS)
      ) %>% 
      arrange(desc(`Cor. with selection`)) %>% 
      as_hux() %>% 
      set_align(everywhere, -1, "right") %>% 
      set_number_format(everywhere, 2:3, 3) %>% 
      set_number_format(everywhere, 4, 2) %>% 
      set_caption("Correlations of polygenic scores with income group.") %>% 
      theme_compact() %>% 
      set_width(0.75) %>% 
      set_col_width(c(.4, .2, .2, .2)) %>% 
      set_bottom_border(final(1), everywhere) %>% 
      set_position("center")
```

## Estimating the effect of natural selection on the contribution of PSEA to inequality

```{r calc-load-ineq-ea3}

res <- drake::readd(res_ineq_ea3)

prop_effect_selection <- res$r2_income_ea3_wt/res$r2_income_ea3 - 1
prop_effect_selection <- scales::percent(prop_effect_selection, accuracy = 0.1)

prop_wt_fe <- res$r2_income_ea3_fe_wt/res$r2_income_ea3_fe - 1
prop_wt_fe <- scales::percent(prop_wt_fe, accuracy = 0.1)

res <- lapply(res, format, digits = 3)

```

From the meta-analysis of twin studies in @branigan2013variation, we take the
heritability of educational attainment to be around 0.4. This is the variance of
educational attainment explained by the "true" polygenic score for educational
attainment $p^{*}$, i.e. the best linear unbiased predictor of educational
attainment from genetic data. We write the measured polygenic
score of educational attainment, "EA3" from @lee2018gene, as
\[
p = p^* + \eta.
\]

For realized educational attainment $EA$, write:
\[
EA = \alpha+\beta p^* + \varepsilon = \alpha + \beta(p - \eta) + \varepsilon
\]
We assume the classical errors-in-variables model, i.e. that $\eta$
is independent of both $p$ and $\varepsilon$. Under these conditions,
the ratio between the estimated $R^{2}$ in a regression of $p$ on
$EA$, and the true $R^{2}$ of $p^{*}$ on $EA$, satisfies: 
\begin{equation}
\frac{\textrm{plim}\hat{R}^{2}}{R^{2}} 
= \frac{\sigma_{p}^{2} - \sigma_{\eta}^{2}}{\sigma_{p}^{2}}
= \frac{\sigma_{p^*}^{2}}{\sigma_{p}^{2}}
\equiv \lambda, \label{eq:r2-ratio}
\end{equation}
as we show below. The right hand side is the "reliability ratio".
We regress educational attainment on EA3 in our sample, weighting
by our Age/Qualification weights, to give 
$\textrm{plim}\hat{R}^2 = `r res$r2_edu_ea3`$. Plugging this
into the above, with $R^2 = h^2\approx0.4$, gives $\lambda=`r res$lambda`$.

We now consider the $R^2$ of $p$ on income, that is, the proportion
of variance in income that is explained by the true polygenic score
of educational attainment. Applying (\ref{eq:r2-ratio}) again gives
that
\begin{equation}
\frac{\textrm{plim}\hat{R}_{\textrm{income}}^{2}}{R_{\textrm{income}}^{2}}=\lambda.\label{eq:r2-ratio-income}
\end{equation}

We compute $\hat{R}_{\textrm{income}}^{2}$by regressing income group on EA3,
first weighting by age/qualification, and then by age/qualification weights
times respondent's number of children. The estimated $R^2$ are 
`r res$r2_income_ea3` and `r res$r2_income_ea3_wt` respectively. Applying
(\ref{eq:r2-ratio-income}) to both estimates gives the proportion of variance in
income explained by true polygenic score for education attainment (a) among
respondents as `r res$r2_income_true_psea`; (b) accounting for natural selection
as `r res$r2_income_true_psea_wt`. The second figure gives the counterfactual
estimate of $R_{\textrm{income}}^{2}$ after one generation of natural selection,
holding environmental conditions unchanged.

The errors-in-variables conditions are quite restrictive. In particular, they
rule out that the unmeasured part of PSEA ($\eta$) correlates with the error 
term $\varepsilon$, e.g. with environmental causes of educational attainment. 
If unmeasured PSEA correlates positively with $\varepsilon$, then the
reliability coefficient $\lambda$ will be lower than calculated. However, the
proportional effect of natural selection on the PSEA-income relationship will
remain the same, so long as $\lambda$ does not differ between the two sets of
weighted regressions.

The estimate above uses a pooled regression of income on EA3, which may include
confounding environmental variation such as genetic nurture. Within-siblings
regression would give a smaller estimate of the effect of EA3. However, this
would also affect the reliability of the regression [@griliches1986errors],
making it hard to back out estimates of the true $R^2$. Nevertheless, we
can calculate the proportional increase in $R^2$ due to natural selection, again 
so long as the reliability coefficient is unaffected by
the weighting scheme. Running within-siblings regressions of EA3 on income group
gives within-$R^2$ estimates of `r res$r2_income_ea3_fe` (weighted by
age and qualification) and `r res$r2_income_ea3_fe_wt` (multiplying weights by number of children). Thus, this procedure estimates the proportional change in $R^2_{income}$ due to natural selection as `r prop_wt_fe`.

### Proof of \@ref(eq:r2-ratio)

Write 
\begin{align}
\hat{R}^{2} & =1-\frac{\sum_{i=1}^{N}(\hat{\alpha}+\hat{\beta}p_{i}-EA_{i})^{2}}{\sum_{i=1}^{N}(EA_{i}-\bar{EA})^{2}}\label{eq:r-squared}
\end{align}

Using the fact that $\hat{\alpha}=\bar{EA}-\hat{\beta}\bar{p}$, where bars signify
sample averages, we can simplify the numerator in the above:

\[
\sum_{i=1}^{N}(\hat{\alpha}+\hat{\beta}p_{i}-EA_{i})^{2}=\sum_{i=1}^{N}(\bar{EA}-\hat{\beta}\bar{p}+\hat{\beta}p_{i}-EA_{i})^{2}=\sum_{i=1}^{N}(\hat{\beta}(p_{i}-\bar{p})-(EA_{i}-\bar{EA}))^{2}
\]

Now, we can compute the following probability limit:

\begin{align}
 & \text{plim}\frac{\sum_{i=1}^{N}(\hat{\beta}(p_{i}-\bar{p})-(EA_{i}-\bar{EA}))^{2}}{N}\nonumber \\
= & \text{plim}\frac{\sum_{i=1}^{N}\hat{\beta}^{2}(p_{i}-\bar{p})^{2}+(EA_{i}-\bar{EA})^{2}-2\hat{\beta}(p_{i}-\bar{p})(EA_{i}-\bar{EA})}{N}\nonumber \\
= & \beta^{2}\lambda^{2}\sigma_{p}^{2}+\sigma_{EA}^{2}-2\beta\lambda\sigma_{p,EA},\label{eq:plim}
\end{align}

using the standard errors-in-variables formula $\textrm{plim}\hat{\beta}=\lambda\beta$,
and where $\sigma_{p,EA}$ is the covariance between $p$ and $EA$.

Write
\begin{align*}
\sigma_{p,EA} & =\textrm{Cov}(p^{*}+\eta,\alpha+\beta p^{*}+\varepsilon)\\
 & =0+\beta\sigma_{p^{*}}^{2}+\beta\textrm{Cov}(\eta,p^{*})+\textrm{Cov}(\eta,\varepsilon)\\
 & =\beta\sigma_{p^{*}}^{2}
\end{align*}

using the errors-in-variables assumption that $\eta$ is independent
of $p^{*}$ and $\varepsilon$. Plug this into (\ref{eq:plim}) to
give
\begin{align*}
 & \text{plim}\frac{\sum_{i=1}^{N}(\hat{\beta}(p_{i}-\bar{p})-(EA_{i}-\bar{EA}))^{2}}{N}\\
= & \beta^{2}\lambda^{2}\sigma_{p}^{2}+\sigma_{EA}^{2}-2\beta^{2}\lambda\sigma_{p^{*}}^{2}\\
= & \beta^{2}\lambda\sigma_{p^{*}}^{2}+\sigma_{EA}^{2}-2\beta^{2}\lambda\sigma_{p^{*}}^{2}\textrm{, using \ensuremath{\sigma_{p^{*}}^{2}=\lambda\sigma_{p}^{2}},}\\
= & \sigma_{EA}^{2}-\beta^{2}\lambda\sigma_{p^{*}}^{2}
\end{align*}

Inserting this into (\ref{eq:r-squared}) gives
\begin{align*}
\textrm{plim}\hat{R}^{2} & =1-\frac{\sigma_{EA}^{2}}{\sigma_{EA}^{2}}+\frac{\beta^{2}\lambda\sigma_{p^{*}}^{2}}{\sigma_{EA}^{2}}\\
 & =\lambda\frac{\beta^{2}\sigma_{p^{*}}^{2}}{\sigma_{EA}^{2}}\\
 & =\lambda R^{2}
\end{align*}
since for bivariate regression $R^{2}$ is the square of the correlation
between $EA$ and $p^{*}$, which can be written $r=$$\beta\frac{\sigma_{p^{*}}^{2}}{\sigma_{EA}\sigma_{p^{*}}}$.
We thank an anonymous commenter on https://stats.stackexchange.com/questions/238878/how-do-errors-in-variables-affect-the-r2
for much of this proof.

\clearpage
\FloatBarrier
## Further results
### Selection effects on raw polygenic scores

```{r calc-regs-controlled}

drake::loadd(res_all)

res_all_controlled <- res_all %>% 
  select(reg.type, dep.var, term, estimate) %>% 
  pivot_wider(names_from = c(reg.type, dep.var), values_from = estimate) %>% 
  mutate(
    consistent = sign(`raw_N siblings`) == sign(`raw_N children`) &
                 sign(`raw_N siblings`) == sign(`controlled_N siblings`) &
                 sign(`raw_N siblings`) == sign(`controlled_N children`)
  )

n_pgs  <- as.double(nrow(res_all_controlled))
prop_consistent_controlled <- mean(res_all_controlled$consistent)
controlled_smaller_sibs <- abs(res_all_controlled[["controlled_N siblings"]]) < 
  abs(res_all_controlled[["raw_N siblings"]])
controlled_smaller_children <- abs(res_all_controlled[["controlled_N children"]]) < 
  abs(res_all_controlled[["raw_N children"]])
median_prop_sibs <- median(abs(res_all_controlled[["controlled_N siblings"]]) /
    abs(res_all_controlled[["raw_N siblings"]]))
median_prop_children <- median(abs(res_all_controlled[["controlled_N children"]]) /
    abs(res_all_controlled[["raw_N children"]]))

```

Figure \@ref(fig:plot-regs-controlled-pcs) compares selection effects on
polygenic scores residualized for the top 100 principal components of the
genetic data, to selection effects on raw, unresidualized polygenic scores. In
siblings regressions, effect sizes are larger for raw scores  -- sometimes much
larger, as in the case of height. `r sum(controlled_smaller_sibs)` out of 
`r n_pgs` "raw" effect sizes have a larger absolute value than the corresponding
"residualized" effect size. The median proportion between raw and controlled
effect sizes is `r median_prop_sibs`. Among the children regressions, this no
longer holds. Effect sizes are barely affected by controlling for principal
components.

Overall, `r prop_consistent_controlled * 100` per cent of effect sizes
are consistently signed across all four regressions (on children and
siblings, and with and without residualization).

```{r plot-regs-controlled-pcs, fig.cap = "Selection effects using unresidualized polygenic scores on number of siblings/children. Sibling estimates weighted by 1/number of siblings."}

drake::loadd(res_all)

n_regs <- as.double(nrow(res_all))
res_all %>% 
      mutate(
        term     = pretty_names(term),
        term     = reorder_within(term, estimate, dep.var, order_abs(1)),
        dep.var  = fct_inorder(dep.var),
        reg.type = fct_recode(reg.type, 
                               `Resid. for 100 PCs`  = "controlled", 
                               Unresidualized        = "raw"
                             )
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, fill = reg.type, 
        colour = p < 0.05/{{n_regs}})) +
        my_vline + 
        facet_wrap(vars(dep.var), scales = "free") +
        pval_geom_point(position = position_dodge(0.4), alpha = 0.8) +
        my_labs() +
        pval_scale() +
        my_fill_scale(n = 2, guide = guide_legend(title = "")) +
        my_theme() +
        scale_y_reordered() +
        theme(
          axis.text.x  = element_text(size = 7),
          legend.text  = element_text(size = 8),
          legend.title = element_text(size = 9)
        )

```

To get a further insight into this we regress *n siblings* and *n children*
on individual principal components. Figure \@ref(fig:plot-pc-regs)
shows the results. Labels show the top principal components. These
have larger effect sizes in siblings regressions. One possibility is that 
the parents' generation was less geographically mobile, and so geographic
patterns of childrearing were more correlated with principal components, which
partly capture the location of people's ancestors.

```{r plot-pc-regs, fig.cap = "Selection effects of 100 principal components of genetic data. Each dot represents one bivariate regression. Sibling estimates weighted by 1/number of siblings. Absolute effect sizes are plotted. Points are jittered on the Y axis. Top principal components are labelled."}

drake::loadd(res_pcs)
n_regs <- as.double(nrow(res_pcs))

pj <- position_jitter(height = 0.15, width = 0, seed = 12345)

res_pcs %>% 
      mutate(
        estimate = abs(estimate),
        dep.var  = fct_relevel(dep.var, "N children"),
        top_pcs  = ifelse(as.numeric(pc_name) <= 5, term, "")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, dep.var, colour = p < 0.05/{{n_regs}})) + 
        geom_point(alpha = 0.8, position = pj) +
        geom_text(aes(estimate, dep.var, label = top_pcs), position = pj, 
                    hjust = "right", color = "black", size = 3) +
        scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) +
        labs(x = "Effect size (log scale)", y = "") + 
        pval_scale(aesthetics = "colour")

```

\FloatBarrier
\clearpage

### Genetic correlations with EA3

```{r plot-rgs-by-effect-size, fig.cap = "Selection effects plotted against genetic correlation with EA3."}

# TODO: ref source of data.

drake::loadd(rgs)
drake::loadd(res_all)

rgs <- left_join(rgs, 
        res_all %>% filter(reg.type == "controlled", dep.var == "N children"), 
        by = c("p2" = "term"))

rgs_cor <- cor.test(~ rg + estimate, 
                      data = rgs %>%
                               filter(
                                 p2 != "EA3_excl_23andMe_UK", 
                                 p2 != "EA2_noUKB"
                               )
                    )
rgs %>% 
      filter(p2 != "EA3_excl_23andMe_UK", p2 != "EA2_noUKB") %>% 
      ggplot(aes(rg, estimate)) + 
        geom_point() +
        my_vline +
        my_hline +
        labs(x = "Genetic correlation with EA3", y = "Effect on number of children")

```

Another way to examine the "earnings" theory of natural selection is to compare
selection effects of polygenic scores with their genetic correlation with educational
attainment (EA3). Since EA3 strongly predicts earnings, if earnings drives
differences in fertility, we'd expect a correlation between the two sets of
results. Figure \@ref(fig:plot-rgs-by-effect-size) shows this is so: the 
correlation, after excluding EA2, is `r rgs_cor$estimate`. Genetic
correlations were calculated using LD score regression from GWAS summary
statistics.

\clearpage
\FloatBarrier

## Model proofs

```{r include-model-appendix, child="model-appendix.Rmd"}

```

\clearpage
\FloatBarrier

# References