### 2022-01-19 why-natural-selection.Rmd
Add Age/Qualification weights to all regressions in this branch

---
title: "Human capital mediates natural selection in contemporary humans"
author: "David Hugh-Jones \\thanks{Corresponding author. School of Economics, University of East Anglia, Norwich, UK. Email: D.Hugh-Jones@uea.ac.uk}, Abdel Abdellaoui \\thanks{Department of Psychiatry, Amsterdam UMC, University of Amsterdam, Amsterdam, The Netherlands. Email: a.abdellaoui@amsterdamumc.nl}"
date: "`r format(Sys.Date(), '%d %B %Y')`"
abstract: "\\noindent Natural selection has been documented in contemporary humans, but  little is known about the mechanisms behind it. We test for natural selection through the association between 33 polygenic scores and fertility, across two generations, using data from UK Biobank (N = 409,629 British subjects with European ancestry). Consistently over time, polygenic scores that predict lower earnings, education and health also predict lower fertility. Selection effects are concentrated among lower SES groups, younger parents, people with more lifetime sexual partners, and people not living with a partner. The direction of natural selection is reversed among older parents, or after controlling for age at first live birth. These patterns are in line with economic theories of fertility, in which earnings-increasing human capital may either increase or decrease fertility via income and substitution effects in the labour market. Studying natural selection can help us understand the genetic architecture of health outcomes: we find evidence in modern day Great Britain for multiple natural selection pressures that vary between subgroups in the direction and strength of their effects, that are strongly related to the socio-economic system, and that may contribute to health inequalities across income groups."
bibliography: "why-natural-selection.bib"
output: 
  bookdown::pdf_document2:
    toc: false
    latex_engine: xelatex
    number_sections: true
    keep_tex: true
header-includes:
  - \usepackage{subfig}
  - \usepackage{setspace}\doublespacing
  - \usepackage{placeins}
  - \usepackage[format=plain, labelfont={bf,it}, textfont=it]{caption}
  - \usepackage{titlesec}
  - \titleformat*{\section}{\sffamily\LARGE}
  - \titleformat*{\subsection}{\sffamily\itshape\Large}
  - \hypersetup{colorlinks = true, linkcolor = {blue}, linkbordercolor = {white}}
  - \usepackage{amsthm}
  - \theoremstyle{plain}
  - \newtheorem{lem}{\protect\lemmaname}
  - \providecommand{\lemmaname}{Lemma}
  - \usepackage{etoc}
  - \usepackage{lineno}
  - \linenumbers
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
mainfont: Baskerville
mathfont: Baskerville
sansfont: "Gill Sans"
---

```{r setup, include = FALSE}

library(drake)
library(magrittr)
library(dplyr)
library(forcats)
library(ggplot2)
library(tidyr)
library(santoku) # keep this after tidyr as it masks chop
library(purrr)
library(glue)
library(broom)
library(huxtable)
loadNamespace("scales")
loadNamespace("shades")

# nasty hack to make add_ashe_income work later:
if (! exists("add_ashe_income")) source("~/import-ukbb-data/import-ukbb-data.R")

drake::loadd(rgs)
drake::loadd(score_names)
drake::loadd(famhist)

options(digits = 2)
options(dplyr.summarise.inform = FALSE) # shut up about summarize

knitr::opts_chunk$set(
        echo       = FALSE, 
        warning    = FALSE, 
        cache      = FALSE, 
        error      = FALSE,
        fig.height = 3.5
      )

knitr::knit_hooks$set(
  inline = function (x) {
    if (is.numeric(x)) x <- as.character(round(x, getOption("digits")))
    x <- gsub("-", "\u2212", x)
    paste(as.character(x), collapse = ", ")
  }
)

options(huxtable.long_minus = TRUE)
huxtable::set_default_properties(latex_float = "h!")

theme_set(theme_minimal())

my_hline <- geom_hline(yintercept = 0, colour = "grey20", linetype = "dotted")
my_vline <- geom_vline(xintercept = 0, colour = "grey20", linetype = "dotted")
pval_palette <- c("FALSE" = "grey65", "TRUE" = "#f3b372")

standard_ggplot <- function (
  dfr, 
  fill_col, 
  n_regs, 
  ...,
  score_col = quo(term), 
  order_idx = 1, 
  fill_direction = 1,
  n_cats    = NULL,
  conf_int  = NULL
) {
  if (! missing(score_col)) score_col <- enquo(score_col)

  mfc <- missing(fill_col)
  fill_col <- if (mfc) quo(NULL) else enquo(fill_col)
  if (missing(n_cats)) {
    n_cats <- if (mfc) 1 else length(unique(pull(dfr, {{fill_col}})))
  }

  if (is.null(conf_int)) conf_int <- "conf.low" %in% names(dfr)
  confint_geom_segment <- if (conf_int) confint_geom_segment() else NULL

  n_regs <- as.double(n_regs)

  dfr %>% mutate(
    !! score_col := pretty_names(!! score_col),
    !! score_col := fct_reorder(!! score_col, estimate, order_abs(order_idx))
  ) %>% 
    rename(p = p.value) %>% 
    ggplot(aes(estimate, {{score_col}}, yend = {{score_col}}, 
               shape = {{fill_col}}, ...)) +
    my_vline + 
    confint_geom_segment +
    geom_point(size = 1.8, aes(color = {{fill_col}})) +
    geom_point(size = 0.7, color = "white", aes(alpha = p < 0.05/33)) +
    my_fill_scale(aesthetics = "color", n = n_cats, direction = fill_direction) +
    scale_shape_manual(values = 1:n_cats + 14) +
    my_labs() +
    my_theme() +
    scale_linetype_manual(name = "", values = c("95% c.i." = 1)) +
    scale_alpha_discrete(range = c(1, 0),
                         guide = guide_legend(
                           override.aes = list(
                             shape = c(1, 16),
                             color = "black",
                             alpha = 1,
                             size = 1.8
                           )
                         ))
}

my_theme <- function (...) theme(axis.text.y = element_text(size = 7), ...)

pval_scale <- function (...) scale_color_manual(values = pval_palette, ...)

pval_geom_point <- function (size = 1.6, stroke = 0.8, fill = NULL, ...) {
  if (is.null(fill)) {
    geom_point(shape = 21, size = size, stroke = stroke, ...)
  } else {
    geom_point(shape = 21, size = size, stroke = stroke, fill = fill, ...)
  }
}

confint_geom_segment <- function () {
  geom_segment(aes(x = conf.low, xend = conf.high, linetype = "95% c.i. uncorrected"), 
                    color = "grey60", alpha = 0.25, size = 1.1, 
                    lineend = "round")
}

my_fill_scale <- function (n, aesthetics = "fill", ...) {

  fill_colors <- c("steelblue4", "darkred")

  if (n <= 2) {
    dots <- list(...)
    dots$direction <- NULL
    do.call(scale_colour_manual, c(list(values = fill_colors[1:n],
                                        aesthetics = aesthetics), 
                                   dots))
  } else {
    shades::lightness(scale_color_brewer(aesthetics = aesthetics, ...), 
                        shades::scalefac(0.85))
  }
}

my_labs <- function (x = "Effect size", y = "", ...) labs(x = x, y = y, ...)

# for use when reordering factors. Order by the nth score in a group
order_abs <- function (n = 1) {
  function (x) x[n] 
}

comma_num <- function (x) prettyNum(x, big.mark = ",")

pretty_names <- function (names) {
  pretty <- c(
    ADHD_2017               = "ADHD",
    age_at_menarche         = "Age at menarche",
    age_at_menopauze        = "Age at menopause",
    agreeableness           = "Agreeableness",
    ai_substance_use        = "Age at smoking initiation",
    alcohol_schumann        = "Alcohol use",
    alzheimer               = "Alzheimer",
    autism_2017             = "Autism",
    bipolar                 = "Bipolar",
    bmi_combined            = "BMI",
    body_fat                = "Body Fat",
    caffeine                = "Caffeine",
    cannabis                = "Cannabis (ever vs. never)",
    cognitive_ability       = "Cognitive Ability",
    conscientiousness       = "Conscientiousness",
    coronary_artery_disease = "Coronary Artery Disease",
    cpd_substance_use       = "Cigarettes per day",
    diagram_T2D             = "Type 2 Diabetes",
    dpw_substance_use       = "Drinks per week",
    EA2_noUKB               = "Educ. attainment 2 (no UKBB)",
    EA3_excl_23andMe_UK     = "Educ. attainment 3 (no UK)",
    eating_disorder         = "Eating disorder",
    extraversion            = "Extraversion",
    height_combined         = "Height",
    hip_combined            = "Hip circumference",
    MDD_PGC2_noUKB          = "Major Depressive Disorder",
    neuroticism             = "Neuroticism",
    openness                = "Openness",
    sc_substance_use        = "Smoking cessation",
    SCZ2                    = "Schizophrenia",
    si_substance_use        = "Smoking initiation",
    wc_combined             = "Waist circumference",
    whr_combined            = "Waist-hip ratio"
  )

  pretty[names]
}

n_in_regs <- function (var, dv = "RLRS", data = famhist) {
  fh_subset <- data
  if (dv == "RLRS") fh_subset %<>% filter(kids_ss)
  # all PGS have the same NA pattern, so we just use one
  tbl <- table(fh_subset[
                  ! is.na(fh_subset[dv]) & ! is.na(fh_subset$whr_combined), 
                  var
                ])

  c(tbl)
}

add_n <- function (var, famhist_var = var, dv = "RLRS", data = famhist) {
  var <- as.factor(var)
  n <- n_in_regs(famhist_var, dv, data)
  levels(var) <- sprintf("%s (N = %s)", levels(var), 
                         prettyNum(n[seq_len(nlevels(var))], big.mark = ","))
  # var_n <- sprintf("%s (N = %s)", var, prettyNum(n[var], big.mark = ","))
  # var_n[is.na(var)] <- NA
  var
}

# from https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R
# for ordering within facets
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}

```

\normalem

# Introduction

Living organisms evolve through natural selection, in which allele frequencies
change in the population through differential reproduction rates. Studying the
mechanisms behind natural selection can help us better understand how individual
differences in complex traits and disease risk arise [@benton2021influence].
Recent work confirms that natural selection is taking place in modern human
populations, using genome-wide analysis
[@Barban_2016;@beauchamp2016genetic;@conley2016assortative;@kong2017selection;@sanjak2018evidence;@FIEDER202216]. In
particular, genetic variants associated with higher educational attainment are
being selected against, although effect sizes appear small.

As yet we know little about the social mechanisms behind natural selection. A
potential explanation comes from the economic theory of fertility [@becker1960economic]. In this theory, higher potential earnings
have two opposite effects on fertility: a fertility-increasing *income effect*
(higher income makes children more affordable), and a fertility-lowering
*substitution effect* (time spent on childrearing has a higher cost in foregone
earnings). Thus, an individual's *human capital* -- skills and personality
traits which are valuable in labour markets -- can increase or decrease his or
her fertility. This can lead to selection on inherited variants which are linked
to human capital. Furthermore, the economic theory predicts systematic variation
in the relative strength of the income and substitution effects across different
social groups.

This study uses data from UK Biobank [@bycroft2018uk] to learn more about
contemporary natural selection. We test for natural selection on 33 different
polygenic scores by estimating their correlation with fertility. We extend the
analysis over two generations, using data on respondents' number of siblings as
well as their number of children. This is interesting because consistent natural
selection over multiple generations could lead to substantive effects in the
long run. Next, we examine correlations with fertility in different subgroups.
Across the board, selection effects are stronger in groups with lower income and
less education, among younger parents, people not living with a partner, and
people with more lifetime sexual partners. Outside these groups, effects are
weaker and often statistically insignificant. In some subgroups, the direction
of selection is even reversed.

We then show that a simple model of human capital, education and fertility
choices can give rise to these empirical results. To test the theory, we show
that polygenic scores' correlation with fertility is associated with their
correlation with education and earnings. Furthermore, a mediation analysis shows
that part of the correlation with fertility is indeed *via* education. Thus, contemporary natural selection on polygenic scores can be explained by scores' correlation with earnings-increasing human capital.

Lastly, we discuss the effects of natural selection. While measured effects on
mean genetic scores are small, natural selection may substantially increase the
correlation between genetic scores and income, increasing genetic differences 
between different social groups, and thus making the "genetic lottery"
[@harden2021genetic] more unfair.

# Results

We created polygenic scores for 33 traits in `r comma_num(nrow(famhist))`
individuals of European descent, corrected for ancestry using 100 genetic
principal components (see Materials and Methods). Figure
\@ref(fig:plot-means-over-time) plots mean polygenic scores in the sample by
5-year birth intervals. Several scores show consistent increases or declines
over this 30-year period, of the order of 5% of a standard deviation. These
changes could reflect natural selection within the UK population, but also
emigration, or ascertainment bias within the sample [@10.1093/aje/kwx246].

```{r plot-means-over-time, fig.height = 6, fig.cap = "Mean polygenic scores (PGS) by birth year in UK Biobank. Circles are means for 5-year intervals. Lines are 95% confidence intervals. Solid circles show a significant linear increase (green) or decrease (red) over time (p < 0.05/33)."}

drake::loadd(pgs_over_time)

pgs_over_time %>% 
      mutate(score_name = pretty_names(score_name)) %>% 
      ggplot(aes(YOB, group = score_name, colour = Change, shape = Change)) + 
      geom_linerange(aes(ymin = score_lo, ymax = score_hi), color = "grey60", 
                       size = 3) +
      geom_point(aes(y = score), size = 2) +
      facet_wrap(~score_name, ncol = 6) +
      my_hline +
      scale_color_manual(values = c("-" = "darkred", "+" = "darkgreen", "o" = "black")) +
      scale_shape_manual(values = c("-" = "circle", "+" = "circle", "o" = "circle open")) +
      theme(
        line        = element_line(size = 0.5, lineend = "round"),
        strip.text  = element_text(size = 6), 
        axis.text.x = element_text(size = 7),
        legend.position = "none"
      ) +
      labs(y = "PGS", x = "Birth year (5 year intervals)")

```

To test more directly for natural selection, we regress respondents' relative 
lifetime reproductive success (RLRS) on their polygenic scores (PGS):

```{=tex}
\begin{equation}
\mathrm{RLRS}_i = \alpha + \beta\mathrm{PGS}_i + \varepsilon_i
\label{eq:regression}
\end{equation}
```

RLRS is defined as respondents' number of children, divided by the mean number
of children of people born in the same year. The "selection effect", $\beta$,
reflects the strength of natural selection within the sample. In fact, since
polygenic scores are normalized, $\beta$ is the expected polygenic score among
children of the sample [@beauchamp2016genetic].[^normalization] Note that
equation \eqref{eq:regression} does not control for many environmental and genetic
factors that could affect fertility, and as a result, $\beta$ is not an estimate
of the causal effect of a polygenic score on fertility. However, natural
selection is a correlational phenomenon: polygenic scores which correlate with
high fertility are being selected for, whatever the underlying causal
relationship.

[^normalization]:The selection effect $\beta$ equals $Cov(RLRS, PGS)/Var(PGS)$.
Since PGS are normalized to variance 1 and mean 0, this reduces to $Cov(RLRS,
PGS) = E(RLRS\times PGS) - E(RLRS) E(PGS) = E(RLRS \times PGS)$. This is the
polygenic score weighted by relative lifetime reproductive success, which is the
average polygenic score in the next generation [@robertson1966mathematical].

```{r calc-res-weighted}

drake::loadd(res_wt_flb_weights)
drake::loadd(res_wt_msoa_weights)
drake::loadd(res_wt_age_qual_weights)
drake::loadd(res_sibs_parent_weights)
drake::loadd(res_all)
drake::loadd(res_sex)

res_wt_combined <- bind_rows(
  None                = res_all %>% 
                          filter(
                            reg.type == "controlled", 
                            dep.var == "RLRS"
                          ) %>% 
                          select(-reg.type, -dep.var, -score_name),
  Geographical        = res_wt_msoa_weights,
  `Age/Qualification` = res_wt_age_qual_weights,
  `Age/Qual/AFLB`     = res_wt_flb_weights,
  .id = "Weights"
)

res_significance <- res_wt_combined %>% 
                      group_by(term) %>% 
                      summarize(
                        all = all(p.value < 0.05/33), 
                        any = any(p.value < 0.05/33)
                      )
res_sig_all <- sum(res_significance$all)
res_sig_any <- sum(res_significance$any)

res_wt_cb_wide <- res_wt_combined %>% 
                    tidyr::pivot_wider(
                      id_cols     = term,
                      names_from  = Weights,
                      values_from = estimate
                    ) %>% 
                    rowwise() %>% 
                    mutate(
                      consistent = all(c_across(-term) > 0) ||
                        all(c_across(-term) < 0)
                    ) %>% 
                    ungroup() %>% 
                    left_join(
                      res_wt_combined %>% 
                        filter(Weights == "None") %>% 
                        select(term, p.value),
                      by = "term"
                    ) %>% 
                    left_join(
                      res_sex %>% 
                        filter(sex == "Female") %>% 
                        select(term, estimate_females = estimate),
                      by = "term"
                    )

rel_sizes <- res_wt_cb_wide %>% 
                filter(
                  consistent, 
                  p.value < 0.05/length(score_names)
                ) %>% 
                mutate(
                  Geographical = Geographical/None,
                  `Age/Qualification` = `Age/Qualification`/None,
                  `Age/Qual/AFLB` = `Age/Qual/AFLB`/estimate_females
                ) %>% 
                select(-consistent, -None, -p.value, -estimate_females) %>% 
                mutate(
                  term = pretty_names(term)
                )

rel_sizes_summary <- rel_sizes %>% 
                summarize(
                  across(-term, ~ c(mean(.x), median(.x)))
                ) %>% 
                mutate(
                  term = c("Mean", "Median")
                )

mean_rel <- c(rel_sizes_summary[1, 1:3])
```

To correct for ascertainment bias, we weight participants using population data.
We try three alternative weighting schemes: (1) "Geographical" weighting, by
geography, age and presence/absence of a partner; (2) age and highest
educational qualification; and (3) for women only, age, highest qualification,
and age at first live birth (AFLB). Figure \@ref(fig:plot-res-wt) plots
selection effects among the entire sample, estimated with the three weighting
schemes, and also unweighted. Effects for `r res_sig_all` out of 33 polygenic
scores are significant at $p < 0.05/33$ under all four weighting schemes, with
`r res_sig_any` scores reaching significance under at least one scheme.[^balance-diverse-def]
Weighting makes a large difference to estimates: mean effect sizes across all
polygenic scores are increased by a factor of `r mean_rel$Geographical`
(geographical weighting), `r mean_rel[["Age/Qualification"]]`
(age/qualification) or `r mean_rel[["Age/Qual/AFLB"]]` (age/qualification/AFLB).
Estimates might be further affected by weighting on other demographic
variables.[^weighting]

[^balance-diverse-def]: We also check for stabilizing and disruptive selection by
estimating \@ref(eq:regression) with a quadratic term. Stabilizing selection
selects for intermediate values, while disruptive selection selects for extreme 
values. In particular, we find disruptive selection for educational attainment
polygenic scores: at higher values of these scores, the negative effect on
fertility is smaller (Appendix Figure \@ref(fig:plot-purifying)).

[^weighting]: To make interpretation straightforward, we do not use weights in 
results below unless noted. However, all our qualitative results are robust to
using weights.

```{r plot-res-wt, fig.cap = "Selection effects: weighted and unweighted regressions.  Each point represents a single bivariate regression of number of children on a polygenic score. P value threshold is 0.05, Bonferroni-corrected for multiple comparisons. Confidence intervals are uncorrected."}

n_regs <- as.double(length(score_names))

res_wt_combined %>% 
      mutate(
        Weights  = fct_relevel(Weights, "Geographical", "Age/Qualification",
                    "Age/Qual/AFLB", "None"),
        term     = pretty_names(term),
        term     = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      rename(p = p.value) %>% 
      # can't use standard_ggplot because we have non-standard geom_point()
      ggplot(aes(
        x      = estimate, 
        y      = term, 
        yend   = term,
        shape  = Weights, 
        fill   = Weights, 
        colour = p < 0.05/{{n_regs}}
      )) +
      my_vline +
      geom_point(size = 1.6, stroke = 0.8, alpha = 0.9) +
      confint_geom_segment() +
      pval_scale() +
      scale_fill_manual(values = c(
        "None"              = "grey15", 
        "Geographical"      = "dodgerblue4",
        "Age/Qualification" = "steelblue2",
        "Age/Qual/AFLB"      = "darkseagreen4"
      )) +
      scale_shape_manual(values = c(
        "None" = "square filled", 
        "Geographical" = "circle filled",
        "Age/Qualification" = "circle filled", 
        "Age/Qual/AFLB" = "circle filled"
      )) + 
      labs(x = "Effect size", y = "") +
      scale_linetype_manual(name = "", values = c("95% c.i. uncorrected" = 1)) +
      my_theme()

```

We now show the empirical puzzles which motivate our economic model. Each
concerns differences in the strength of natural selection across different
subgroups in the sample. Splitting the sample by demographic and social
variables, we re-estimate \@ref(eq:regression) within each subgroup. We use
socio-economic variables including income and education, and family structure
variables including age at first live birth, presence of a partner, and lifetime
number of sexual partners.

Figure \@ref(fig:plot-income-educ-level) plots selection effects for each
polygenic score, grouping respondents by age of completing full-time education,
and by household income. Effects are larger and more significant for the lowest
income category, and for the lowest education category. These results are robust
to controlling for respondents' age (Appendix section \@ref(sec:age-control)).

```{r plot-income-educ-level, fig.cap = "Selection effects by education and income.", fig.subcap = c("Age left full-time education", "Household income"), fig.ncol = 1}

drake::loadd(res_edu)
n_regs <- as.double(nrow(res_edu))

res_edu %>% 
      mutate(
        "Age left FTE" = fct_relevel(age_fte_cat, "< 16", "16-18", "> 18"),
        "Age left FTE" = add_n(`Age left FTE`, "age_fte_cat")
      ) %>% 
      standard_ggplot(n_regs = n_regs, fill_col = `Age left FTE`, 
                        fill_direction = -1)

drake::loadd(res_income)

n_regs <- as.double(nrow(res_income))
res_income %>% 
      mutate(
        Income = factor(income_cat, 
          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", "> £100K")
        ),
        Income = add_n(Income, "income_cat")
      ) %>% 
      standard_ggplot(n_regs = n_regs, fill_col = `Income`, 
                        fill_direction = -1)
```

Turning to family structure, we split respondents by lifetime number of sexual
partners, at the median value of 3 (Figure \@ref(fig:plot-n-partners)a).
Selection effects are larger and more significant among
those with more than 3 lifetime partners. Next we split respondents by whether
they were living with a spouse or partner at the time of interview.
Effects are larger among those not living with a spouse or partner
(Figure \@ref(fig:plot-n-partners)b).[^sex]

[^sex]: The same pattern holds if we analyse men and women separately (Appendix Figure
\@ref(fig:plot-n-partners-sex)). We also directly compared selection effects 
between men and women (Appendix Figure \@ref(fig:plot-sexes)).

```{r plot-n-partners, fig.cap = "Selection effects by number of sexual partners and presence of a partner.", fig.subcap = c("Lifetime number of sexual partners", "Presence of a partner"), fig.ncol = 1, fig.align = "center", fig.height = 4}

drake::loadd(res_partners_joint)
res_partners_joint %<>%       
      filter(grepl(":", term)) %>% 
      mutate(`N partners` = ifelse(grepl("TRUE:", term), "3 or less", "4 or more"))

n_regs <- as.double(nrow(res_partners_joint))
res_partners_joint %>% 
      standard_ggplot(fill_col = `N partners`, n_regs = n_regs, 
                        score_col = score_name)

drake::loadd(res_with_partner)
res_with_partner %<>% filter(grepl(":", term))

n_regs <- as.double(nrow(res_with_partner))

res_with_partner %>% 
      mutate(
        Household = ifelse(grepl("TRUE", term), "With partner", 
                      "Without partner")
      ) %>% 
      standard_ggplot(fill_col = Household, n_regs = n_regs, 
                        score_col = score_name)
```

```{r calc-age-flb-cross}

drake::loadd(res_age_flb_cross)

cor_age_flb <- res_age_flb_cross %>% 
      mutate(category = gsub(":.*", "", term)) %>% 
      pivot_wider(score_name, names_from = category, values_from = estimate) %>%              {cor(.$`age_flb_cat10-22`, .$`age_flb_cat28-52`)}

```

```{r calc-corr-age-flb}

drake::loadd(res_age_flb)
drake::loadd(res_sex)

res_sex_comparison <- res_sex %>% 
      filter(sex == "Female") %>% 
      mutate(score_name = term) %>% 
      select(score_name, term, estimate:conf.high)

res_age_flb %<>% filter(term != "age_flb")

res_combined <- bind_rows(
        raw      = res_sex_comparison, 
        with_age = res_age_flb %>% select(score_name, term, estimate:conf.high), 
        .id      = "reg.type"
      ) %>% 
      arrange(score_name)

raw_flb_comparison <- res_combined %>% 
      pivot_wider(
        id_cols     = score_name, 
        names_from  = reg.type, 
        values_from = estimate
      )

cor_raw_flb <- cor(raw_flb_comparison$raw, raw_flb_comparison$with_age)
opp_signed <- sum(
        sign(raw_flb_comparison$raw) != sign(raw_flb_comparison$with_age)
      )

```

Lastly, we split female respondents by age at first live birth
(AFLB).[^aflb-no-men] There is evidence for genetic effects on AFLB
[@Barban_2016], and there is a close link between this variable and number of
children born. Figure \@ref(fig:plot-age-flb) shows effect sizes estimated
separately for each tercile of AFLB. Effects are strikingly different
across terciles. Educational attainment, ADHD and MDD are selected for amongst
the youngest third of mothers, but selected against among the oldest two-thirds.
Similarly, several polygenic scores for body measurements are selected against
only among older mothers. The correlation between effect sizes for the youngest
and oldest terciles is `r cor_age_flb`. To investigate this further, we estimate
equation \@ref(eq:regression) among females, *controlling* for AFLB. In 
`r opp_signed` out of `r length(score_names)` cases, effects change sign when
controls are added. The correlation between effect sizes controlling for AFLB,
and raw effect sizes, is `r cor_raw_flb`. Thus, selection effects seem to come
through two opposing channels: a correlation with AFLB, and an opposite-signed
correlation with number of children after AFLB is controlled for.

[^aflb-no-men]: AFLB is unavailable for men.

```{r plot-age-flb, fig.cap = "Selection effects by age at first live birth terciles (women only).", fig.align = "center"}

drake::loadd(res_age_flb_cross)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_cross))

res_age_flb_cross %>% 
      mutate(
        `Age at first live birth` = gsub("age_flb_cat(.*):.*", "\\1", term),
        `Age at first live birth` = add_n(`Age at first live birth`, "age_flb_cat")
      ) %>% 
      standard_ggplot(score_col = score_name, 
                      fill_col = `Age at first live birth`, n_regs = n_regs,
                      fill_direction = -1)

```

```{r calc-siblings-children}
sib_chn_cor <- cor.test(famhist$RLRS_parents, famhist$RLRS, use = "complete")
stopifnot(sib_chn_cor$p.value < 2e-16)
```

We emphasize that our categories are not necessarily exogenous to polygenic
scores. For example -- both in the data (Appendix Figure
\@ref(fig:plot-age-flb-dv)) and in our model -- age at first live birth is a
choice variable, which is endogenous to an individual's human capital and
related polygenic scores. Nevertheless, observed differences in equation
\@ref(eq:regression) across subgroups constrain the set of possible
theories; a good theory of contemporary natural selection needs to explain
how these results arise. As we show below, a model based on the economic
theory of fertility can do just that.

We also examine selection effects among respondents' parents, using information
on respondents' number of siblings to calculate parents' RLRS. Effect sizes of
polygenic scores are highly correlated across the two generations (Appendix
Figure \@ref(fig:plot-res-sibs)). Median-splitting respondents by year of birth,
we find little evidence of change in effect sizes among the parents' generation.
There is some evidence that selection effect sizes are increasing in the 
respondents' generation, with 9 polygenic scores showing a significant increase. 
We also check whether selection effects vary by
socio-economic status and AFLB in the parents' generation. We have no
information about parents' income, so we use the 1971 Townsend deprivation score
of respondents' birthplace as a proxy [@townsend1987deprivation]. Results show
the same pattern as for respondents. Effect sizes are larger and more often
significant in the most deprived areas (Appendix Figure
\@ref(fig:plot-siblings-townsend)). Effects are larger among younger parents,
and change sign when controlling for AFLB, among both mothers and fathers
(Appendix Figures \@ref(fig:plot-age-birth-parents-cross), 
\@ref(fig:plot-age-birth-parents)). Lastly, we check for a
"quantity-quality tradeoff" between parents' number of children and number of
grandchildren. We do not find any: in fact, the correlation between respondents' 
and parents' RLRS is positive ($\rho$ = `r sib_chn_cor$estimate`, 
$p < 2 \times 10^{-16}$).

\FloatBarrier

# Human capital and natural selection

These results show that selection effects are weaker, absent, or even reversed
among some subgroups of the population. A possible explanation for this comes 
from the economic theory of fertility 
[@becker1960economic; @willis1973new;@becker1976child]. According to this theory,
increases in a person's wage affect their fertility via two opposing channels.
There is an *income effect* by which children become more affordable, like any
other good. There is also a *substitution effect*: since childrearing has a cost
in time, the opportunity cost of childrearing increases if one's market wage is
higher. The income effect leads higher earners to have more children. The
substitution effect leads them to have fewer.

Suppose that certain genetic variants correlate with *human capital*: skills or
other characteristics that affect an individual's earnings in the labour market
[@mincer1958investment; @becker1964human]. These variants may then be associated
with opposing effects on fertility. The income effect will lead to natural
selection in favour of earnings-increasing variants (or variants that are merely
associated with higher earnings). The substitution effect will do the reverse.

To show this, we write down a simple model of fertility choices. $h$ is
an individual's level of human capital. For now, we simply identify this with
his or her wage $W$. Raising a child takes time $b$. People maximize utility $U$
from the number of children $N$ and from income $Y\equiv(1-bN)W$:

\[
U = u(Y)+aN.
\]

Here $a$ captures the strength of preference for children. $u(\cdot)$ captures
the taste for income, and is increasing and concave. We treat $N$ as continuous,
in line with the literature: this can be thought of as the expected number of
children among people with a given $a$, $b$ and $W$. The marginal
benefit of an extra child is $\frac{dU}{dN} = -bWu'(Y)+a$. The effect of an
increase in human capital on this marginal benefit is

\[
\frac{d^{2}U}{dNdh}=\underbrace{-bu'(Y)}_{\textrm{Substitution effect}}\underbrace{-bYu''(Y)}_{\textrm{Income effect}}.
\]

The *substitution effect* is negative and reflects that when wages increase,
time devoted to childcare costs more in foregone income. The positive *income
effect* depends on the curvature of the utility function, and reflects that when
income is higher, the marginal loss of income from children is less painful.

To examine education and fertility timing, we extend the model to
two periods. For convenience we ignore time discounting, and assume
that credit markets are imperfect so that agents cannot borrow. Write

\begin{equation}
U(N_{1},N_{2}) = u(Y_{1}) + u(Y_{2}) + aN_{1} + aN_{2}\label{eq:U}
\end{equation}

Instead of identifying human capital with wages, we now allow individuals to
choose a level of education $s \in [0,1]$, which has a time cost in period 1.
Education is complementary to human capital $h > 0$, and increases period 2
wages, which take the  simple functional form $w(s,h) = sh$. We normalize period
1 wages to 1, and let $u(\cdot)$ take the constant relative risk aversion form
$u(y)=\frac{y^{1 - \sigma} - 1}{1 - \sigma}$. $\sigma > 0$ measures the
curvature of the utility function, i.e. the decline in marginal utility of
income as income increases. We examine total fertility 
$N^{*} = N_{1}^{*} + N_{2}^{*}$ and the *fertility-human capital relationship*,
$\frac{dN^{*}}{dh}$. For $\sigma < 1$ in a neighbourhood of $\sigma = 1$, Table
\@ref(tab:theory) shows five theoretical predictions, along with our
corresponding empirical results for the correlation between polygenic scores and
RLRS.[^one-period-results] The key insight of the model is that for middling
levels of $\sigma$, the substitution effect dominates at low income levels, but
is balanced by the income effect as income increases.

[^one-period-results]: Predictions 1-3 also hold in the one-period model with
constant relative risk aversion. Our empirical results are actually stronger
than prediction 5, in that correlations with fertility are *reversed* at higher
AFLB. This prediction can be accommodated in the model if children have a money
cost as well as a time cost (Appendix Figure \@ref(fig:N-plot-with-m)).

------------------------------------------------------------------------------------
    **Theory: the fertility-human capital         **Empirical results**
    relationship is...**                                      
--- --------------------------------------------- ----------------------------------
1.  Negative: $\frac{dN^{*}}{dh} < 0$.            Figures \@ref(fig:plot-means-over-time) 
                                                  and \@ref(fig:plot-res-wt).

2.  Weaker (closer to zero) at higher wages       Figure 
    and/or levels of human capital.               \@ref(fig:plot-income-educ-level)a.
                                                  Selection effects are also weaker at
                                                  higher polygenic scores for 
                                                  educational attainment 
                                                  (Appendix Figure
                                                  \@ref(fig:plot-purifying)).

3.  More negative when the time burden            Stronger effects for single parents
    of children $b$ is larger.                    (Figure \@ref(fig:plot-n-partners)).

4.  Weaker at higher levels of education $s$.     Figure \@ref(fig:plot-income-educ-level)b.

5.  Weaker among those who start fertility        Effects weaker among those starting 
    in period 2 ($N_{1}^{*} = 0$) than among      fertility later 
    those who start fertility in period 1         (Figure \@ref(fig:plot-age-flb)).
    ($N_{1}^{*} > 0$).

------------------------------------------------------------------------------------

Table: (\#tab:theory) Predictions from the theoretical model and corresponding 
empirical results.

Thus, a simple economic model can explain many of our results.
Other empirical work in economics also supports our mechanism.
@caucutt2002women and @monstad2008education show that education and skills
affect age at first birth and fertility. Income decreases fertility at low
income levels, but increases it at higher income levels [@cohen2013financial].
US fertility decreases faster with education among single mothers than married
mothers [@baudin2015fertility], in line with our prediction 3 and as predicted by
@becker1981treatise. A related literature shows negative correlations
between IQ and fertility [e.g. @lynn2004new; @reeve2018systematic].

```{r plot-cors-earnings-educ, fig.cap="Selection effects by correlations with earnings and educational attainment. Each point represents one polygenic score."}

drake::loadd(res_all)
drake::loadd(res_age_flb)
drake::loadd(res_cor_educ)
drake::loadd(res_cor_income)

effect_size <-  res_all %>% 
                  filter(dep.var == "RLRS", reg.type == "controlled") %>% 
                  pull(estimate)

effect_size_flb <-  res_age_flb %>% 
                      filter(term != "age_flb") %>% 
                      pull(estimate)

dfr <- data.frame(
        PGS          = rep(score_names, 4),
        Correlation  = rep(c("Earnings", "Education"), each = 33, 2),
        cor          = rep(c(res_cor_income[,1], res_cor_educ[,1]), 2),
        effect_size  = c(rep(effect_size, 2), rep(effect_size_flb, 2)),
        Controls     = rep(c("None", 
                        "Age at first live birth"), each = 66)
      )

dfr %>% 
      mutate(
        PGS      = ifelse(abs(effect_size) > 0.02, PGS, ""),
        PGS      = sub("(^\\w*?)_.*", "\\1", PGS)
      ) %>% 
      filter(
        Controls == "None"
      ) %>% 
      ggplot(aes(effect_size, cor, colour = Correlation)) + 
        geom_point() +
        geom_text(aes(label = PGS), size = 3, colour = "black", 
                    check_overlap = TRUE, nudge_x = 0.002) +
        facet_wrap( ~ Correlation, scales = "free") + 
        my_vline + 
        my_hline +
        labs(x = "Selection effect", y = "Correlation") +
        coord_cartesian(clip = "off") +
        scale_colour_brewer(type = "qual", palette = 2) +
        theme(
          legend.position = "none",
          panel.spacing = unit(20, "pt")
        )

```

# Testing the theory

We test this explanation in two ways. First, the economic theory predicts that
genetic variants will be selected for (or against) in proportion to their
correlation with human capital. Figure \@ref(fig:plot-cors-earnings-educ) plots
selection effects on each polygenic score against that score's correlation with
two measures of human capital: earnings in a respondent's first job, and
educational attainment. The relationships are strongly negative. Thus, human
capital appears to be relevant to natural selection. The negative relationship
suggests that substitution effects dominate income effects, which fits the known
negative association between income and fertility
[@becker1960economic;@jones2006economic]. The correlations reverse when we
control for age at first live birth, suggesting that within AFLB categories, the
income effect dominates. (See also the Supplementary Animations for the
uncorrected effects per AFLB group, which show the correlation reversing after
age 22.)

```{r calc-mediation}
drake::loadd(res_mediation)

res_mediation %<>% mutate(
                     # 0.025 not 0.05 because two-sided
                     sig     = abs(statistic_ind) > qnorm(1 - 0.025/33),
                     pos_sig = sig & prop_ind > 0
                   )

n_pos_sig <- sum(res_mediation$pos_sig)
median_prop <- median(res_mediation$prop_ind[res_mediation$pos_sig])

```

Second, for the `r nrow(res_mediation)` scores where the bivariate correlation in
\eqref{eq:regression} is significant at $p < 0.05/33$, we run a mediation
analysis to directly test whether the correlation between each polygenic score
and fertility is mediated by educational attainment (Appendix Table
\@ref(tab:tbl-res-mediation)). Figure \@ref(fig:plot-mediation) shows estimated
proportions explained along with bootstrap 95% confidence intervals
(uncorrected; 100 bootstraps). For `r n_pos_sig` scores, the indirect effect of
the score on fertility *via* educational attainment takes the same sign as the
overall effect, and is significantly different from zero ($p < 0.05/33$). Among
these scores, the median proportion of the total effect explained by the
indirect effect is `r percent(median_prop)`. The educational attainment variable
is a relatively crude measure of human capital: more accurate measures would
likely explain more of the total effect.

```{r plot-mediation, fig.align = "center", fig.cap = "Proportion of selection effect mediated by educational attainment, polygenic scores with significant selection effects"}

res_mediation %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, prop_ind, order_abs(1)),
        `Indirect effect p < 0.05/33` = sig 
      ) %>% 
      ggplot(aes(y = term)) + 
        geom_col(aes(x = prop_ind, fill = `Indirect effect p < 0.05/33`), 
                   width = .5) + 
        geom_errorbar(aes(
                          xmin = prop_ind_conf_low, 
                          xmax = prop_ind_conf_high,
                          linetype = "Bootstrap 95% c.i.",
                        ),
                        width = 0.3,
                        color = "grey50"
                      ) +
        scale_fill_manual(values = pval_palette) +
        my_vline +
        scale_x_continuous(labels = scales::percent, n.breaks = 7) +
        coord_cartesian(xlim = c(-0.5, 1)) +
        labs(x = "Proportion mediated", y = "") +
        scale_linetype_manual(name = "", values = c("Bootstrap 95% c.i." = 1))

```

```{r calc-risk-attitude}

drake::loadd(res_risk_control)
drake::loadd(res_all)

n_risk_sig <- res_risk_control %>% 
                filter(term == 'f.2040.0.0') %>% 
                {sum(.$p.value <= .Machine$double.eps)}

n_pgs_sig <- res_risk_control %>% 
               filter(term == score_name) %>% 
               {sum(.$p.value <= 0.05/33)}

res_compare <- res_all %>% 
      filter(dep.var == "RLRS", reg.type == "controlled") %>% 
      left_join(res_risk_control %>% filter(term == score_name), 
                  by = "score_name")

ratio <- res_compare$estimate.y/res_compare$estimate.x
n_less_sig <- sum(res_compare$p.value.y <= 0.05/33 & res_compare$p.value.x > 0.05/33)
stopifnot(n_less_sig == 0)
```

We consider three alternative theories that might explain our results. First,
welfare benefits which incentivize child-bearing might be taken up more among
low-income individuals. However, the majority of effect sizes appear unchanged
over a large span of twentieth-century history (Appendix Table
\@ref(tab:tbl-change-by-period)), during
which government spending on child-related benefits varied considerably
[@hoc1999socsec]. In general, there is only weak evidence that welfare benefits
affect fertility [@gauthier2007impact]. Future
work could test this theory more explicitly. A second alternative theory is that
polygenic scores correlate with the motivation to have children, i.e. parameter
$a$ in the model [cf. @jones2008fertility]. This theory would not explain why
selection effects are smaller at higher incomes and education levels. Indeed,
in the model, the effect of $a$ on fertility becomes stronger at higher levels
of human capital. A third alternative is that traits under selection are linked
to externalizing behaviour and risk-seeking. This might be partially captured by
our parameter $\sigma$, which can be interpreted as a measure of risk aversion
over income. However, a more direct channel is risky sexual behaviour
[@mills2021identification]. The data here provide some support for this: scores
which might plausibly be linked to externalizing behaviour, like ADHD and
younger age at smoking initiation, are selected for. However, this theory is
less good at explaining variation in selection across the full range of scores,
including physical measures such as waist-hip ratio and BMI. We test this theory
directly by re-estimating equation \@ref(eq:regression) controlling for a
measure of risk attitude (UK Biobank field 2040).  The median ratio of effect
sizes between regressions with and without controls is `r median(ratio)`; all
scores which are significant at $p < 0.05/33$ in uncontrolled regressions remain
so when controlling for risk attitude. This non-result could simply reflect the
imprecision of the risk attitude measure, which is a single yes/no question.
However, this measure *does* predict the overall number of children, highly
significantly ($p < 2 \times 10^{-16}$ in `r n_risk_sig` out of 33 regressions).
Given that, and the statistical power we get from our sample size, we believe
that the non-result is real: while risk attitude does predict fertility in the
sample, it is not an important channel for natural selection.

# Discussion

Previous work has documented natural selection in modern populations on variants
underlying polygenic traits [@beauchamp2016genetic; @kong2017selection;
@sanjak2018evidence]. We show that correlations between polygenic scores and
fertility are highly concentrated among specific subgroups of the population,
including people with lower income, lower education, younger first parenthood,
and more lifetime sexual partners. Among mothers aged 22+, selection effects
are reversed. Furthermore, the size of selection effects on a polygenic score
correlates with that score's association with labour market earnings. Strikingly,
some of these results were predicted by @fisher1930genetical, pp. 253-254.

The economic theory of fertility provides a parsimonious explanation for these
findings. Because of the substitution effect of earnings on fertility, scores
are selected for when they correlate with low human capital, and this effect is
stronger at lower levels of income and education.

Polygenic scores which correlate with high (low) earnings and more (less)
education are being selected against (selected for). In addition, many of the
phenotypes under positive selection are linked to disease risk. Many people
would probably prefer to have high educational attainment, a low risk of ADHD
and major depressive disorder, and a low risk of coronary artery disease, but
natural selection is pushing against genes associated with these traits.
Potentially, this could increase the health burden on modern populations, but
that depends on effect sizes. Our results suggest that naïve estimates can be
affected by sample ascertainment bias. There may be remaining sources of
ascertainment bias after our weighting; if so, we expect that, like the sources
of ascertainment we have controlled for, they probably bias our results towards
zero. This problem may be less serious in surveys which aim to be representative
(as the UK Biobank does not). However, there is still scope for bias, since not
all respondents consent to the collection of genetic data. For instance,
completion rates for genotype data in the US Health and Retirement Study were
around 80-85% [@hrs2020]. Researchers should be aware of the risks of
ascertainment when studying modern natural selection.

We also do not know how estimated effect sizes of natural selection will change
as more accurate measures of genetic variation are produced, or whether genetic
variants underlying other phenotypes will show a similar pattern to those
studied here. In addition, effects of polygenic scores may be inflated in
population-based samples, because of indirect genetic effects, gene-environment
correlations, and/or assortative mating [@lee2018gene;
@selzam2019comparing; @kong2018nature; @Howe2021.03.05.433935], although we do
not expect that this should impact their association with number of offspring,
or the resulting changes in allele frequencies. In short, it is probably too
early to tell whether modern natural selection has a substantively important
effect on population means of phenotypes under selection. Nevertheless, we note
that selection effects on our measured polygenic scores are still relatively
small, even after reweighting to account for ascertainment bias.

```{r calc-inequality}
drake::loadd(res_ineq)

n_increases <- sum(res_ineq$ratio > 1)
median_pct_change <- (median(res_ineq$ratio) - 1) * 100

```

Because selection effects are concentrated in lower-income groups, they may also
increase inequality with respect to polygenic scores. For example, Figure
\@ref(fig:plot-mean-EA3-by-income) graphs mean polygenic scores for educational
attainment (EA3) among children from households of different income groups. The
blue bars show the actual means, i.e. parents' mean polygenic score weighted by
number of children. The grey bars show the hypothetical means if all households
had equal numbers of children. Natural selection against genes associated with
educational attainment, which lowers the mean, is stronger at the bottom of the
income distribution, and this increases the differences between groups.
Overall, natural selection increases the correlation of polygenic scores with
income for `r n_increases` out of 33 polygenic scores, with a median
percentage increase of `r median_pct_change`% (Appendix Table
\@ref(tab:tbl-inequality-effects)).  If inequalities in polygenic scores are
important for understanding social structure and mobility 
[@belsky2018genetic; @Rimfeld_2018; @harden2021genetic], then these increases
are substantive. Similarly, since many polygenic scores are predictive of
disease risk, they could potentially increase health inequalities. In general,
the evolutionary history of anatomically modern humans is related to disease
risk [@benton2021influence]; understanding the role of contemporary natural
selection may aid in mapping the genetic architecture of current health
disparities.

```{r plot-mean-EA3-by-income, fig.align = "center", fig.cap = "Mean polygenic score for educational attainment (EA3) of children by household income group. Blue is actual. Grey is hypothetical in the absence of selection effects. Respondents weighted by Age/Qualification.", fig.width = 4}
income_EA3 <- famhist %>% 
      filter(! is.na(n_children), ! is.na(income_cat)) %>% 
      mutate(
        `Household income` = factor(income_cat, 
                          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", 
                                      "> £100K"))
      ) %>% 
      group_by(`Household income`) %>% 
      summarize(
        `Without selection` = mean(EA3_excl_23andMe_UK, na.rm = TRUE),
        Actual              = weighted.mean(EA3_excl_23andMe_UK, n_children, 
                                              na.rm = TRUE)
      ) %>% 
      tidyr::pivot_longer(-`Household income`) %>% 
      rename(
        `Children's mean EA3` = value
      )

income_EA3 %>% 
              ggplot(aes(`Household income`, `Children's mean EA3`, fill = name)) +
                geom_col(position = "dodge") +
                scale_fill_manual(values = c("navy", "grey60")) +
                theme(
                  legend.background = element_rect(fill = "white", colour = NA),
                  legend.title = element_blank(), 
                  legend.position = c(.25, .85)
                )
```

Existing evidence on human natural selection has led some to "biocosmic
pessimism" [@sarraf2019modernity]. Others are more sanguine, and argue that
natural selection's effects are outweighed by environmental improvements, such
as those underlying the Flynn effect [@flynn1987massive]. The evidence
here may add some nuance to this debate. Patterns of natural selection have been
relatively consistent across the past two generations, but they are not the
outcome of a universal, society-wide phenomenon. Instead they result from 
opposing forces, operating in different parts of society and pulling in
different directions.

Any model of fertility is implicitly a model of natural selection, but so far,
the economic and human genetics literatures have developed in parallel.
Integrating the two could deepen our understanding of natural selection in
modern societies. Economics possesses a range of theoretical models on the
effects of skills, education and income [see @hotz1997economics;
@lundberg2007american]. One perennial problem is how to test these theories in a
world where education, labour and marriage markets all interact. Genetic data,
such as polygenic scores, could help to pin down the direction of causality,
for example via Mendelian randomization [@davey2003mendelian].
Conversely, theory and empirical results from economics can shine a light on the
mechanisms behind natural selection, and thereby on the nature of individual
differences in complex traits and disease risk.

# Materials and methods

We use participant data from UK Biobank [@bycroft2018uk], which has received
ethical approval from the National Health Service North West Centre for Research
Ethics Committee (reference: 11/NW/0382). We limit the sample to white British
participants of European descent, as defined by genetic estimated ancestry and
self-identified ethnic group, giving a sample size of
`r comma_num(nrow(famhist))`. For regressions on number of children we use
participants over 50 (males)/45 (females), since most fertility is completed by this age. This gives
a sample size of `r comma_num(nrow(famhist %>% filter(kids_ss)))`.

Polygenic scores were chosen so as to cover a reasonably broad range of traits,
and based on the availability of a large and powerful GWAS which did not include
UK Biobank. Scores were computed by summing the alleles across ~1.3 million
genetic variants weighted by their effect sizes as estimated in 33 genome-wide
association studies (GWASs) that excluded UK Biobank. To control for population
stratification, we corrected the polygenic scores for 100 principal components
(PCs).

To compute polygenic scores and PCs, the same procedures were followed as 
described in @abdellaoui2019genetic.

Earnings in first job are estimated from mean earnings in the 2007 Annual Survey
of Hours and Earnings, using the SOC 2000 job code (Biobank field 22617).

Population data for weighting is taken from the 2011 UK Census and the 2006
General Household Survey (GHS). Weighting for Age/Qualification and 
Age/Qualification/AFLB weights was done using marginal totals from a linear
model, using the `calibrate()` function in
the R "survey" package [@lumley2020survey]. Geographical weighting was done with
iterative post-stratification using the `rake()` function, on Census Middle Layer
Super Output Areas, sex and presence/absence of a partner.

Code for this paper is available at 
https://github.com/hughjonesd/why-natural-selection.

## Acknowledgements

AA is supported by the Foundation Volksbond Rotterdam and by ZonMw grant 849200011 from The Netherlands Organisation for Health Research and Development. This study was conducted using UK Biobank resources under application numbers 40310 and 19127.

\FloatBarrier
\clearpage
# Appendix

\localtableofcontents
\clearpage

## Selection effects by sex

Figure \@ref(fig:plot-sexes) plots selection effects by sex. Differences are
particularly large for educational attainment, height and MDD. Several polygenic
scores for mental illness and personality traits are more selected for (or less
against) among women, including major depressive disorder (MDD), schizophrenia
and neuroticism, while extraversion is more selected for among men. Scores for
waist circumference and waist-hip ratio are less selected for among women. One
possible reason for these sex differences is that polygenic scores may affect
fertility via success in marriage markets, and men and women may value different
characteristics in these.

```{r plot-sexes, fig.cap = "Selection effects by sex. Highlighted lines are significant differences at p < 0.05/33. Highlighted points are significantly different from 0 at p < 0.05/66."}

drake::loadd(res_sex)
n_regs <- as.double(nrow(res_sex))

res_sex_wide <- res_sex %>% 
      select(term, estimate, sex, diff.p.value) %>% 
      pivot_wider(names_from = "sex", values_from = "estimate") %>% 
      mutate(
        min_est = pmin(Female, Male),
        max_est = pmax(Female, Male),
        term = pretty_names(term),
        term = fct_reorder(term, Female)
      ) 
n_comps <-as.double(nrow(res_sex_wide))

res_sex %>% 
      mutate(
        term = pretty_names(term),
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, fill = sex, 
            colour = p < 0.05/{{n_regs}})) +
        my_vline +
        geom_linerange(data = res_sex_wide, 
          aes(
            y      = term,
            x      = NULL,
            fill   = NULL,
            xmin   = min_est, 
            xmax   = max_est, 
            colour = diff.p.value < 0.05/{{n_comps}}
          ), 
        group = "") +
        pval_geom_point() +
        pval_scale(name = "p < 0.05 (corrected)") +
        my_fill_scale(n = 2) +
        my_labs() +
        my_theme()

```

\FloatBarrier
\clearpage
## Weighted regressions

Table \@ref(tab:tbl-res-wt) gives effect sizes as a proportion of the unweighted
effect size, for all polygenic scores which are consistently signed and which
are significantly different from zero in unweighted regressions.

```{r tbl-res-wt}

rel_sizes %>% 
      arrange(desc(Geographical)) %>% 
      bind_rows(rel_sizes_summary) %>% 
      rename(PGS = term) %>% 
      as_hux() %>% 
      insert_row("", "Weighting", fill = "") %>% 
      set_caption(
        "Weighted effect sizes as a proportion of unweighted effect sizes."
      ) %>% 
      merge_cells(1, 2:4) %>% 
      set_align(1, 2, "centre") %>% 
      set_header_rows(1, TRUE) %>% 
      theme_compact() %>% 
      set_number_format(-1, -1, 2) %>% 
      style_headers(bold = TRUE) %>% 
      set_bottom_border(1, 2:4) %>% 
      set_bottom_border(2, everywhere) %>% 
      set_italic(final(2), 1) %>% 
      set_font_size(8) %>% 
      set_width(0.7) %>% # ensures line wrapping below
      set_col_width(c(.34, .22, .22, .22)) %>% 
      add_footnote(paste0(
        "Only consistently-signed and significant (when unweighted) ",
        "estimates are shown. ",
        "Age/Qual/AFLB as a proportion of unweighted regressions ",
        "including females only."),
        top_padding  = 0,
        left_padding = 0
      )

```

\FloatBarrier
\clearpage
## Stabillizing and disruptive selection

Stabilizing selection reduces variance in the trait under selection, while
disruptive selection increases variance. To check for these, we rerun equation
\@ref(eq:regression), adding a quadratic term in $PGS_i$, and using our
age/qualification weights. Scores for hip circumference show significant
stabilizing selection ($p < 0.05/33$, negative coefficient on quadratic term).
The EA2 score for educational attainment shows significant disruptive selection
($p < 0.05/33$, positive coefficient), which reduces the strength of selection
against educational attainment at very high levels of the PGS. (The quadratic on
the EA3 score has a similar coefficient but is not significant at $p < 0.05/33$.)
Figure \@ref(fig:plot-purifying) plots predicted number of children
against polygenic score from these regressions.

We also checked for stabilizing selection in the parents' generation, using
age/qualification weights multiplied by the inverse of number of siblings.
Scores for EA2 and EA3 show significant disruptive selection ($p < 0.05/33$, 
positive coefficient on quadratic). Other scores including hip circumference 
were not significant.

*Weighted regressions only*: in addition to the above, 
major depressive disorder shows diversifying selection
and drinks per week show balancing selection.

```{r plot-purifying, fig.align = "center", fig.cap = "Stabilizing/disruptive selection: predicted number of children by polygenic score.", fig.height = 3}

drake::loadd(res_quadratic)
sig <- res_quadratic %>% 
         filter(grepl("^2", term, fixed = TRUE)) %>% 
         filter(p.value < 0.05/33) %>% 
         pull(score_name)

stopifnot(sig == c("EA2_noUKB", "hip_combined"))

sig_all_terms <- res_quadratic %>% filter(score_name %in% sig)

drake::loadd(res_sibs_quadratic)
sig_sibs <- res_sibs_quadratic %>% 
         filter(grepl("^2", term, fixed = TRUE)) %>% 
         filter(p.value < 0.05/33) %>% 
         pull(score_name)
stopifnot(sig_sibs == c("EA2_noUKB", "EA3_excl_23andMe_UK"))

op <- par(mfrow = c(1, 2))

for (score in sig) {
  coef <- sig_all_terms %>% filter(score_name == score) %>% pull(estimate)
  curve(
         coef[1] + coef[2] * x + coef[3] * x^2,
         xlim = c(-4, 4), 
         xlab = pretty_names(score),
         ylab = "Predicted RLRS"
       )
}
par(op)

```

\FloatBarrier
\clearpage
## Controlling for age {#sec:age-control}

```{r calc-income-edu-controlled}

drake::loadd(res_income_controlled, res_edu_controlled)
n_regs_ic <- nrow(res_income_controlled)
n_sig_ic <- sum(res_income_controlled$p.value < 0.05/n_regs_ic)

n_regs_educ <- nrow(res_edu_controlled)
n_sig_educ <- sum(res_edu_controlled$p.value < 0.05/n_regs_educ)
```

Results in Figure \@ref(fig:plot-income-educ-level) could be
explained by age, if older respondents have lower income and are less educated,
and also show more natural selection on polygenic scores. However, when we rerun
the regressions, interacting the polygenic score with income category and also
with a quadratic in age, the interaction with income remains significant at
$p < 0.05/`r n_regs_ic`$ for `r n_sig_ic` out of `r n_regs_ic` regressions. Similarly
if we interact the PGS with age of leaving full time education and a quadratic
in age, the interaction with age leaving full time education remains significant
at $p < 0.05/`r n_regs_educ`$ for `r n_sig_educ` out of `r n_regs_educ` regressions.

\FloatBarrier
\clearpage
## Number of partners and presence of partner by sex

Figure \@ref(fig:plot-n-partners-sex) splits up Figure \@ref(fig:plot-n-partners)
by sex. The pattern of results is the same in both sexes: selection effects are stronger among those with more lifetime sexual partners, and among those not
currently living with a partner.

```{r plot-n-partners-sex, fig.cap = "Selection effects by number of sexual partners and presence of a partner, for men and women separately.", fig.subcap = c("Lifetime number of sexual partners", "Presence of a partner"), fig.ncol = 1, fig.align = "center", fig.height = 4}

sexes_theme <-  my_theme(
                  plot.margin          = unit(c(1, 1, 0, 1), "lines"),
                  legend.position      = "bottom", 
                  #legend.box.spacing   = unit(0, "lines"),
                  legend.box.margin    = margin(-5, 0, 0, -85),
                  legend.margin        = margin(0, 0, 0, 0),
                  legend.box.just      = "left",
                  legend.justification = "left",
                  legend.spacing.y     = unit(0, "lines"),
                  #legend.spacing       = unit(0.5, "lines"),
                  legend.title         = element_text(size = 10)
                  #legend.text          = element_text(size = 8)
                )
sexes_guides <- guides(
                  fill  = guide_legend(ncol = 1),
                  color = guide_legend(ncol = 1)
                )

drake::loadd(res_partners)
res_partners %<>%       
      filter(grepl(":", term)) %>% 
      mutate(`N partners` = ifelse(grepl("TRUE:", term), "3 or less", "4 or more"))

n_regs <- as.double(nrow(res_partners))
res_partners %>% 
      mutate(
        score_name = pretty_names(score_name)
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(
        y      = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        yend   = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        x      = estimate, 
        fill   = `N partners`, 
        colour = p < 0.05/{{n_regs}}
      )) + 
        my_vline +
        confint_geom_segment() +
        pval_geom_point() +
        facet_wrap(vars(sex), scales = "free_y") +
        my_fill_scale(n = 2) +
        pval_scale() +
        sexes_theme +
        my_labs(x = "", y = "") + 
        scale_y_reordered() +
        scale_linetype_manual(name = "", values = c("95% c.i. uncorrected" = 1)) +
        sexes_guides

drake::loadd(res_with_partner_sex)
res_with_partner_sex %<>% filter(grepl(":", term))

n_regs <- as.double(nrow(res_with_partner_sex))

res_with_partner_sex %>% 
      mutate(
        score_name = pretty_names(score_name),
        Household = ifelse(grepl("TRUE", term), "With partner", 
                      "Without partner")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(
        y      = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        yend   = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        x      = estimate, 
        fill   = Household, 
        colour = p < 0.05/{{n_regs}}
      )) + 
        my_vline +
        confint_geom_segment() +
        pval_geom_point() +
        facet_wrap(vars(sex), scales = "free_y") +
        my_fill_scale(n = 2) +
        pval_scale() +
        sexes_theme +
        my_labs(x = "", y = "") + 
        scale_y_reordered() +
        scale_linetype_manual(
                                name   = "", 
                                values = c("95% c.i. uncorrected" = 1)
                              ) +
        sexes_guides

```

\FloatBarrier
\clearpage
## Parents' generation {#sec-parents}
### Selection effects and change over time

```{r calc-res-sibs}
drake::loadd(res_wt_age_qual_weights)
drake::loadd(res_children_comparison)

cmp <- left_join(res_wt_age_qual_weights, res_children_comparison, by = "term")

cor_inc_vs_excl_childless <- cor(cmp$estimate.x, cmp$estimate.y)
n_smaller <- sum(abs(cmp$estimate.y) < abs(cmp$estimate.x))

pct_diff <- abs(cmp$estimate.y)/abs(cmp$estimate.x) - 1
median_pct_diff <- median(pct_diff)

```

The UK Biobank data contains information on respondents' number of siblings
(including them), i.e. their parents' number of children. Since respondents'
polygenic scores are equal in expectation to the mean scores of their parents,
we can use this to look at selection effects in the parents' generation. We
estimate equation (\ref{eq:regression}) using parents' RLRS as the dependent
variable.[^parent-RLRS] The parents' generation has an additional source of ascertainment
bias: sampling parents of respondents overweights parents who have many
children. For instance, parents of three children will have, on average, three
times more children represented in UK Biobank than parents of one child. Parents
of no children will by definition not be represented. To compensate, we reweight
our preferred weightings (Age/Qualification) by the inverse of *number of
siblings*.

[^parent-RLRS]: We don't have data on parents' year of birth for most 
respondents. To create parents' RLRS, we divide respondents' number of siblings 
by the average number of siblings of all respondents born in the same year,
weighting the average by respondents' inverse of number of siblings to 
compensate for ascertainment bias.

Figure \@ref(fig:plot-res-sibs) shows regressions of parents' RLRS on
polygenic scores. For a clean comparison with the respondents' generation, we
rerun regressions on respondents' RLRS excluding those with no children, and
show results in the figure. Selection effects are highly correlated across the
two generations, and most share the same sign. Absolute effect size estimates
are larger for the parents' generation. We treat this result cautiously, because
effect sizes in both generations may depend on polygenic scores' correlation with
childlessness, and we cannot estimate this for the parents' generation.

To learn more about this, we compare effect sizes excluding and including
childless people in the *current* generation. The correlation between the two
sets of effect sizes is `r cor_inc_vs_excl_childless`. So, patterns across
different scores are broadly similar whether the childless are counted or not.
However, absolute effect sizes are smaller when the childless are excluded, for
`r n_smaller` out of 33 scores; the median percentage change is 
`r percent(median_pct_diff)`.

```{r plot-res-sibs, fig.cap = "Selection effects, respondents' parents vs. respondents. Age/qualification weights. Parental generation weights multiplied by 1/number of siblings. Respondents' regression excludes childless respondents."}
drake::loadd(res_sibs_parent_weights)
drake::loadd(res_children_comparison)

res_sibs_kids <-  bind_rows(
                    Siblings = res_sibs_parent_weights, 
                    Children = res_children_comparison,
                    .id = "Dep. var."
                  )

standard_ggplot(res_sibs_kids, `Dep. var.`, n_regs = length(score_names))
```

```{r calc-period-regs}

drake::loadd(res_period_parents)
drake::loadd(res_period_children)

res_period <- bind_rows(res_period_parents, res_period_children)

res_period %<>% 
      tidyr::extract(term, c("period", "score_name"),
        regex="year_split(.*):(.*)")
n_tests <- nrow(res_period)/2

summary_period <- res_period %>% 
      select(-std.error, -statistic) %>% 
      pivot_wider(
        names_from  = period, 
        values_from = c(estimate, conf.low, conf.high, p.value)
      ) %>% 
      group_by(score_name, children) %>% 
      mutate(
        significant = diff.p.value < 0.05/n_tests,
        sign_early  = sign(estimate_early),
        sign_late   = sign(estimate_late),
        size_inc    = abs(estimate_early) < abs(estimate_late),
        change_sign = sign_early != sign_late,
        Change      = case_when(
                        ! significant ~ "Insignificant",
                        change_sign   ~ "Change sign",
                        size_inc      ~ "Size increasing",
                        ! size_inc    ~ "Size decreasing"
                      )
      ) %>% 
      select(score_name, children, Change, significant, change_sign, 
        sign_early, sign_late) %>% 
      mutate(children = ifelse(children, "Respondents", "Parents")) %>% 
      pivot_wider(names_from = children, 
        values_from = c(Change, significant, change_sign, sign_early, 
        sign_late)) %>% 
      mutate(overall = case_when(
              ! change_sign_Parents & ! change_sign_Respondents & 
                sign_late_Parents == sign_early_Parents ~ 
                "Consistent",
              TRUE ~ "Changes direction"
            ))

stopifnot(sum(summary_period$Change_Parents == "Insignificant") == 32)
stopifnot(sum(summary_period$Change_Respondents == "Size increasing") == 9)
stopifnot(sum(summary_period$Change_Respondents == "Change sign") == 1)
stopifnot(sum(summary_period$Change_Respondents == "Insignificant") == 23)
```

The fact that childless people have such a strong effect on estimates makes it
hard to compare total effect sizes across generations. In particular, since the
parents' generation has a different distribution of numbers of children,
childless people may have had more or less effect in that generation. Another
issue is that we are estimating parents' polygenic scores by the
scores of their children. This introduces noise into our independent
variable, which might lead to errors-in-variables and bias coefficients towards
zero.

As an alternative approach, we run regressions interacting polygenic scores with
birth year, median split at 1950 ("early born" versus "late born"). We use both
respondents' RLRS and parents' RLRS as a dependent variable. We weight
using age/qualification cells, and further adjust for selection in the parents'
generation (see above).

Table \@ref(tab:tbl-change-by-period) 
summarizes the results. We report the number of scores showing significant
changes over time (i.e. a significant interaction between polygenic score and
the "late born" dummy): either a significant change in sign, a significant
increase in size, or a significant decrease in size. There is little evidence
for changes in selection effects within the parents' generation, with just one
score showing a significant decrease in size. In the respondents' generation,
effect sizes were significantly larger in absolute size among the later-born for
nine polygenic scores: ADHD, age at menopause, caffeine consumption, cognitive
ability, Coronary Artery Disease, EA2, EA3, extraversion and Major Depressive
Disorder. These changes are inconsistent with the intergenerational change,
where estimated effect sizes were larger among the earlier, parents' generation.
One score, for alzheimer's disease, showed a significant change in sign, from
negative to positive effects on fertility.

Overall, there is some evidence for an increase in the strength of selection in
recent history. The clearest result is that the pattern of relative effect sizes
across scores is broadly consistent over time.

```{r tbl-change-by-period}

tbl_resp <- summary_period %>% 
              group_by(Change_Respondents) %>% 
              count()

tbl_par <- summary_period %>% 
             group_by(Change_Parents) %>% 
             count()

tbl_both <- full_join(tbl_par, tbl_resp, 
                        by = c("Change_Parents" = "Change_Respondents"))

tbl_both %>% 
            rename(
              Change              = Change_Parents, 
              `Parents' RLRS`     = n.x, 
              `Respondents' RLRS` = n.y
            ) %>% 
            as_hux() %>%
            set_bold(1, everywhere) %>% 
            set_bottom_border(1, everywhere) %>% 
            add_footnote(glue("Significance is measured at p < 0.05/{n_tests}."), 
              border = 0.4, font_size = 8) %>% 
            set_top_padding(1) %>% 
            set_bottom_padding(1) %>%
            set_left_padding(everywhere, 1, 0) %>% 
            set_right_padding(everywhere, 3, 0) %>% 
            set_caption("Numbers of polygenic scores showing changes in selection effects between early and late born. 
      Weights for respondents: Age/Qualification. Weights for parents: Age/Qualification times 1/number of siblings.") %>% 
            set_caption_width(1)

```

\clearpage
\FloatBarrier
### Area deprivation

Figure \@ref(fig:plot-siblings-townsend) plots effects on parents' RLRS by
Townsend deprivation quintile of birth area.

```{r plot-siblings-townsend, fig.cap = "Selection effects (parents' RLRS) by Townsend deprivation quintile of birth area. Higher = more deprived. Estimates weighted by 1/number of siblings.", fig.align = "center"}

drake::loadd(res_townsend_parents)
n_regs <- nrow(res_townsend_parents)

res_townsend_parents %>% 
    mutate(
        n = prettyNum(n, big.mark = ","),
        `Townsend quintile` = paste0(quintile, " (N = ", n, ")")
      ) %>% 
  standard_ggplot(fill_col    = `Townsend quintile`,
                    n_regs    = n_regs, 
                    score_col = score_name, 
                    order_idx = 5
                  )

```

For comparison, Figure \@ref(fig:plot-townsend) plots effects on 
respondents' RLRS by Townsend deprivation quintile of birth area.

```{r plot-townsend, fig.cap = "Selection effects in the respondents' generation by Townsend deprivation quintile of birth area. Higher = more deprived.", fig.align = "center"}

drake::loadd(res_townsend)
n_regs <- nrow(res_townsend)

res_townsend %>% 
    mutate(
        n = prettyNum(n, big.mark = ","),
        `Townsend quintile` = paste0(quintile, " (N = ", n, ")")
      ) %>% 
  standard_ggplot(fill_col    = `Townsend quintile`,
                    n_regs    = n_regs, 
                    score_col = score_name, 
                    order_idx = 5
                  )

```

\clearpage
\FloatBarrier
### Age at first live birth

```{r calc-age-birth-parents}

drake::loadd(res_age_birth_parents)
drake::loadd(res_all)

res_sibs_unc <- filter(res_all, dep.var == "RLRS_parents", reg.type == "controlled")
compare_abp_unc <- res_age_birth_parents %>% 
      filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
      left_join(res_sibs_unc, by = "term")

cor_unc_fath <- with(
  compare_abp_unc %>% filter(control == "fath_age_birth"), 
  cor(estimate.x, estimate.y))
cor_unc_moth <- with(
  compare_abp_unc %>% filter(control == "moth_age_birth"), 
  cor(estimate.x, estimate.y))
```

Among the parents' generation, we can control for age at first live birth using
the subsets of respondents who reported their mother's or father's age, and who
had no elder siblings. We run regressions on parents' RLRS on these
subsets. Figure \@ref(fig:plot-age-birth-parents-cross) shows selection effects
by terciles of age at first live birth, for mothers and fathers. As in the
respondents' generation, effect sizes are smaller, or even oppositely signed, for older parents. Importantly, this holds for both sexes.

```{r plot-age-birth-parents-cross, fig.cap = "Selection effects (parents' RLRS) by age at first live birth terciles.  Estimates weighted by 1/number of siblings.", fig.subcap = c("Mothers", "Fathers"), fig.ncol = 1}

drake::loadd(res_age_flb_mothers_cross)
drake::loadd(res_age_flb_fathers_cross)

famhist_bo1 <- famhist %>% 
                 filter(birth_order == 1) %>%
                 select(RLRS_parents, whr_combined, moth_age_birth_cat, 
                          fath_age_birth_cat)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_mothers_cross))

res_age_flb_mothers_cross %>% 
      mutate(
        `Age at first live birth` = gsub("moth_age_birth_cat(.*):.*", "\\1", term),
        `Age at first live birth` = add_n(`Age at first live birth`, 
                                            "moth_age_birth_cat", 
                                            dv   = "RLRS_parents",
                                            data = famhist_bo1
                                          )
      ) %>% 
      standard_ggplot(score_col = score_name, 
                      fill_col = `Age at first live birth`, n_regs = n_regs,
                      fill_direction = -1)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_fathers_cross))

res_age_flb_fathers_cross %>% 
      mutate(
        `Age at first live birth` = gsub("fath_age_birth_cat(.*):.*", "\\1", term),
        `Age at first live birth` = add_n(`Age at first live birth`, 
                                            "fath_age_birth_cat", 
                                            dv   = "RLRS_parents",
                                            data = famhist_bo1
                                          )
      ) %>% 
      standard_ggplot(score_col = score_name, 
                      fill_col = `Age at first live birth`, n_regs = n_regs,
                      fill_direction = -1)

rm(famhist_bo1)

```

Figure \@ref(fig:plot-age-birth-parents) shows the regressions controlling for
either parent's age at their birth. Effect sizes are very similar, whether
controlling for father's or mother's age. As in the respondents' generation,
effect sizes are negatively correlated with the effect sizes from bivariate
regressions without the age at birth control (father's age at birth: $\rho$ 
`r cor_unc_fath`; mother's age at birth: $\rho$ `r cor_unc_moth`).

```{r plot-age-birth-parents, fig.cap = "Selection effects (parents' generation) among eldest siblings, controlling for parents' age at birth. Estimates weighted by 1/number of siblings."}

n_regs <- as.double(nrow(res_age_birth_parents)/2)

res_age_birth_parents %>% 
    filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
    mutate(
      Control = ifelse(control == "fath_age_birth", 
                  "Father's age at birth", 
                  "Mother's age at birth"
                )
    ) %>% 
    standard_ggplot(fill_col = Control, n_regs = n_regs, order_idx = 2)

```

\clearpage
\FloatBarrier
## Effects of polygenic scores on age at first live birth

```{r calc-age-flb-dv}

drake::loadd(res_age_flb_dv)
drake::loadd(res_age_birth_parents_dv)

compare_flb_dv <- left_join(
        res_age_flb_dv           %>% select(term, estimate),
        res_age_birth_parents_dv %>% select(term, estimate, dep.var),
        by = "term"
      )

cor_flb_fath <- with(compare_flb_dv %>% filter(dep.var == "fath_age_birth"),
        cor(estimate.x, estimate.y)
      )
cor_flb_moth <- with(compare_flb_dv %>% filter(dep.var == "moth_age_birth"),
        cor(estimate.x, estimate.y)
      )
```

Our results suggest that polygenic scores may directly correlate with age at
first live birth. Figure \@ref(fig:plot-age-flb-dv) plots estimated effect sizes
from bivariate regressions for respondents. Figure
\@ref(fig:plot-age-birth-parents-dv) does the same for their parents, using only
eldest siblings.[^only-eldest] Effect sizes are reasonably large. They are also
highly correlated across generations. Effect sizes of polygenic scores on
father's age at own birth, and on own age at first live birth, have a
correlation of `r cor_flb_fath`; for mother's age and own age it is 
`r cor_flb_moth`.

[^only-eldest]: Parental AFLB can only be calculated for this group.

```{r plot-age-flb-dv, fig.cap = "Effects of polygenic scores on age at first live birth."}

n_regs <- as.double(nrow(res_age_flb_dv))

res_age_flb_dv %>% standard_ggplot(n_regs = n_regs)

```

```{r plot-age-birth-parents-dv, fig.cap = "Effects of polygenic scores on parents' age at respondent's birth, eldest siblings. Estimates weighted by 1/number of siblings."}

n_regs <- as.double(nrow(res_age_birth_parents_dv))

res_age_birth_parents_dv %>% 
      mutate(
        "Dependent variable" = dplyr::recode(dep.var, 
          "fath_age_birth" = "Father's AFLB", 
          "moth_age_birth" = "Mother's AFLB"
        )
      ) %>% 
      standard_ggplot(fill_col = `Dependent variable`, n_regs = n_regs)

```

\FloatBarrier
\clearpage
## Mediation analysis

We run a standard mediation analysis in the framework of @baron1986moderator.
For each polygenic score where the bivariate correlation with RLRS is 
significant at $p$ < 0.05/33, we estimate 
\begin{align}
RLRS_i & = \alpha + \beta PGS_i + \gamma EA_i + X_i\mu + \varepsilon_i \\
EA_i & = \delta + \zeta PGS_i + X_i\mu + \eta_i
\end{align}
where $RLRS_i$ is relative lifetime reproductive success, $PGS_i$ is the
polygenic score, $EA_i$ is educational attainment (age of leaving fulltime
education), and $X_i$ is a vector of controls. The total effect of $PGS$ on
number of children is $\beta + \gamma \zeta$. The "indirect effect" mediated by
$EA$ is $\gamma \zeta$. The standard error of the indirect effect can be
calculated as

\[
\sqrt{\hat\gamma^2 \hat\sigma_\zeta^2 + \hat\zeta^2\hat\sigma_\gamma^2}
\]

where $\hat\sigma_\zeta$ is the standard error of $\hat\zeta$, etc.

We are not concerned about endogeneity of polygenic scores, since our theory
requires only that these correlate with human capital. However, endogeneity of
educational attainment is a concern, i.e., educational attainment might
correlate with unobserved characteristics that affect one's fertility (and which
are not related to human capital). To partially address this, we include
controls for age, sex, fluid IQ, height and risk attitude in $X$. Note that
controlling for fluid IQ is conservative, in the sense that IQ might also measure
human capital; nevertheless, we control for it because fluid IQ might affect
fertility decisions via other channels than human capital.

```{r tbl-res-mediation}

drake::loadd(res_mediation)

res_mediation %<>% mutate(
                     # 0.025 not 0.05 because two-sided
                     sig  = abs(statistic_ind) > qnorm(1 - 0.025/33),
                     star = ifelse(sig, "*", "")
                   )

n_pos_sig <- sum(res_mediation$sig & res_mediation$prop_ind > 0)

res_mediation %>%
      arrange(desc(estimate_total)) %>%
      transmute(
        PGS = pretty_names(term),
        `Total effect`    = estimate_total,
        `Indirect effect` = paste(estimate_ind, star),
        `Proportion (%)`      = prop_ind * 100,
        `Proportion 95% c.i. (%)` = glue::glue(
                     "[{prop_ind_conf_low * 100}, {prop_ind_conf_high * 100}]"
                     )
      ) %>%
      as_hux() %>%
      set_align(everywhere, -1, "right") %>% 
      set_number_format(-1, 2:3, 4) %>% 
      set_number_format(-1, 4:5, 1) %>% 
      set_caption("Mediation analysis") %>% 
      theme_compact() %>% 
      set_width(0.95) %>%
      set_col_width(c(.25, .15, .15, .15, .25)) %>% 
      set_bottom_border(final(), everywhere) %>%
      set_font_size(9) %>% 
      add_footnote("* p < 0.05/33.", font_size = 8)
```

Table \@ref(tab:tbl-res-mediation) shows results. For `r n_pos_sig` out of
`r nrow(res_mediation)` scores, the indirect effect on fertility via human 
capital is significantly different from 0 at $p = 0.05/33$ and has the
same sign as the total effect. We also calculate
the proportion of the total effect that is mediated via the indirect effect,
along with uncorrected 95% confidence intervals (100 bootstraps).

\FloatBarrier
\clearpage
## Within-siblings regressions

```{r calc-fe-fertility}
drake::loadd(res_fe_fertility)
drake::loadd(res_all)

educ_reduction <- with(res_fe_fertility, 
       estimate[Regression == "Controlled" & term == "EA3_excl_23andMe_UK"] / 
       estimate[Regression == "Raw" & term == "EA3_excl_23andMe_UK"]
     )
educ_reduction <- 1 - educ_reduction

res_all %<>% filter(dep.var == "RLRS", reg.type == "controlled")
res_fe_fertility %<>% filter(Regression == "Raw")
res_all %<>% left_join(res_fe_fertility, by = "term")

lm_pool_w <- tidy(lm(estimate.y ~ estimate.x, res_all))
coef_pool_w <- lm_pool_w[[2, "estimate"]]
se_pool_w <- lm_pool_w[[2, "std.error"]]

```

Results in the main text support our theory that natural selection on polygenic scores is
driven by their *correlation* with human capital. Here, we test whether polygenic
scores *cause* fertility by running within-siblings regressions (Figure
\@ref(fig:plot-res-fe-fertility)). With a reduced sample size, all
within-sibling effects are insignificant after Bonferroni correction. However,
effect sizes are positively correlated with effect sizes from the pooled model,
and about 50% smaller (regressing within-sibling on pooled effect sizes, $b$ =
`r format(coef_pool_w, digits = 3)`). This attenuation is in line with the ~50%
decrease in heritability in within-sibling GWASs on age at first birth and
educational attainment [@Howe2021.03.05.433935]. We see these results as
providing tentative evidence that polygenic scores cause fertility, with effects
being partly driven by correlations with environmental variation in human
capital. We also reran within-siblings regressions adding a control for
education. Most effect sizes barely change, suggesting that our measure of
education does not in general mediate differences in fertility among siblings.

```{r plot-res-fe-fertility, fig.cap = "Selection effects controlling for sibling-group fixed effects, with and without a control for education (left education before 16, 16-18, or after 18). Each set of 29 results is from a single regression of RLRS on 29 polygenic scores. Standard errors clustered by sibling group."}

drake::loadd(res_fe_fertility)
n_regs <- nrow(res_fe_fertility)/2 # Raw and Controlled treated separately

standard_ggplot(res_fe_fertility, fill_col = Regression, n_regs = n_regs,
                  fill_direction = -1) +
  guides(alpha = guide_legend(
                   override.aes = list(
                             shape = 1,
                             color = "black",
                             alpha = 1,
                             size = 1.8
                   )
                 )
  )

n_groups <- attr(res_fe_fertility, "n_groups")
n <- attr(res_fe_fertility, "n")

```

We run a single regression on 29 polygenic scores within `r n_groups` sibling
groups (N = `r n`). Thus, we control both for environmental confounds (since
scores are randomly allocated within sib-groups by meiosis), and for genetic
confounds captured by our polygenic scores. We remove four scores which
correlate highly with other scores (educational attainment 2, hip circumference,
waist circumference and waist-hip ratio). Figure \@ref(fig:plot-res-fe-fertility)
shows the results.

\clearpage
\FloatBarrier
## Effects on inequality

Table \@ref(tab:tbl-inequality-effects) shows correlations between children's
polygenic scores and household income (UKB data field 738). Column "With
selection" uses respondents' scores weighted by Age/Qualification, times number
of children. Column "Without selection" uses scores weighted by
Age/Qualification only, i.e. the counterfactual correlation if all respondents
had the same number of children.

```{r tbl-inequality-effects}

res_ineq %>% 
      select(
        PGS                      = score, 
        `Cor. with selection`    = actual, 
        `Cor. without selection` = cf,
        `Ratio`                  = ratio
      ) %>% 
      mutate(
        PGS = pretty_names(PGS)
      ) %>% 
      arrange(desc(`Cor. with selection`)) %>% 
      as_hux() %>% 
      set_align(everywhere, -1, "right") %>% 
      set_number_format(everywhere, 2:3, 3) %>% 
      set_number_format(everywhere, 4, 2) %>% 
      set_caption("Correlations of polygenic scores with income group.") %>% 
      theme_compact() %>% 
      set_width(0.75) %>% 
      set_col_width(c(.4, .2, .2, .2)) %>% 
      set_bottom_border(final(1), everywhere) %>% 
      set_position("center")
```

\clearpage
\FloatBarrier
## Further results
### Selection effects on raw polygenic scores

```{r calc-regs-controlled}

drake::loadd(res_all)

res_all_controlled <- res_all %>% 
  select(reg.type, dep.var, term, estimate) %>% 
  pivot_wider(names_from = c(reg.type, dep.var), values_from = estimate) %>% 
  mutate(
    consistent = sign(`raw_RLRS_parents`) == sign(`raw_RLRS`) &
                 sign(`raw_RLRS_parents`) == sign(`controlled_RLRS_parents`) &
                 sign(`raw_RLRS_parents`) == sign(`controlled_RLRS`)
  )

n_pgs  <- as.double(nrow(res_all_controlled))
prop_consistent_controlled <- mean(res_all_controlled$consistent)
controlled_smaller_sibs <- abs(res_all_controlled[["controlled_RLRS_parents"]]) < 
  abs(res_all_controlled[["raw_RLRS"]])
controlled_smaller_children <- abs(res_all_controlled[["controlled_RLRS"]]) < 
  abs(res_all_controlled[["raw_RLRS"]])
median_prop_sibs <- median(abs(res_all_controlled[["controlled_RLRS_parents"]]) /
    abs(res_all_controlled[["raw_RLRS_parents"]]))
median_prop_children <- median(abs(res_all_controlled[["controlled_RLRS"]]) /
    abs(res_all_controlled[["raw_RLRS"]]))

```

Figure \@ref(fig:plot-regs-controlled-pcs) compares selection effects on
polygenic scores residualized for the top 100 principal components of the
genetic data, to selection effects on raw, unresidualized polygenic scores. In
siblings regressions, effect sizes are larger for raw scores  -- sometimes much
larger, as in the case of height. `r sum(controlled_smaller_sibs)` out of 
`r n_pgs` "raw" effect sizes have a larger absolute value than the corresponding
"residualized" effect size. The median proportion between raw and controlled
effect sizes is `r median_prop_sibs`. Among the children regressions, this no
longer holds. Effect sizes are barely affected by controlling for principal
components.

Overall, `r prop_consistent_controlled * 100` per cent of effect sizes
are consistently signed across all four regressions (on children and
siblings, and with and without residualization).

```{r plot-regs-controlled-pcs, fig.cap = "Selection effects using unresidualized polygenic scores on number of siblings/children. Sibling estimates weighted by 1/number of siblings.", fig.height = 6}

drake::loadd(res_all)

n_regs <- as.double(nrow(res_all))
res_all %>% 
      mutate(
        term     = pretty_names(term),
        term     = reorder_within(term, estimate, dep.var, order_abs(1)),
        dep.var  = fct_inorder(dep.var),
        dep.var  = fct_recode(dep.var, 
                                `Respondents' RLRS` = "RLRS",
                                `Parents' RLRS` = "RLRS_parents" 
                              ),
        reg.type = fct_recode(reg.type, 
                               `Resid. for 100 PCs`  = "controlled", 
                               Unresidualized        = "raw"
                             )
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(y = term, x = estimate, fill = reg.type, 
        colour = p < 0.05/{{n_regs}})) +
        my_vline + 
        facet_grid(rows = vars(dep.var), scales = "free") +
        pval_geom_point(position = position_dodge(0.4), alpha = 0.8) +
        my_labs() +
        pval_scale() +
        my_fill_scale(n = 2, guide = guide_legend(title = "")) +
        my_theme() +
        scale_y_reordered() +
        theme(
          axis.text.x  = element_text(size = 7),
          legend.text  = element_text(size = 8),
          legend.title = element_text(size = 9)
        )

```

To get a further insight into this we regress respondents' and parents' RLRS
on individual principal components. Figure \@ref(fig:plot-pc-regs)
shows the results. Labels show the top principal components. These
have larger effect sizes in siblings regressions. One possibility is that 
the parents' generation was less geographically mobile, and so geographic
patterns of childrearing were more correlated with principal components, which
partly capture the location of people's ancestors.

```{r plot-pc-regs, fig.cap = "Selection effects of 100 principal components of genetic data. Each dot represents one bivariate regression. Sibling estimates weighted by 1/number of siblings. Absolute effect sizes are plotted. Points are jittered on the Y axis. Top principal components are labelled."}

drake::loadd(res_pcs)
n_regs <- as.double(nrow(res_pcs))

pj <- position_jitter(height = 0.15, width = 0, seed = 12345)

res_pcs %>% 
      mutate(
        estimate = abs(estimate),
        dep.var  = fct_relevel(dep.var, "RLRS"),
        top_pcs  = ifelse(as.numeric(pc_name) <= 5, term, "")
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(estimate, dep.var, colour = p < 0.05/{{n_regs}})) + 
        geom_point(alpha = 0.8, position = pj) +
        geom_text(aes(estimate, dep.var, label = top_pcs), position = pj, 
                    hjust = "right", color = "black", size = 3) +
        scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) +
        labs(x = "Effect size (log scale)", y = "") + 
        pval_scale(aesthetics = "colour")

```

\FloatBarrier
\clearpage

### Genetic correlations with EA3

```{r plot-rgs-by-effect-size, fig.cap = "Selection effects plotted against genetic correlation with EA3."}

# TODO: ref source of data.

drake::loadd(rgs)
drake::loadd(res_all)

rgs <- left_join(rgs, 
        res_all %>% filter(reg.type == "controlled", dep.var == "RLRS"), 
        by = c("p2" = "term"))

rgs_cor <- cor.test(~ rg + estimate, 
                      data = rgs %>%
                               filter(
                                 p2 != "EA3_excl_23andMe_UK", 
                                 p2 != "EA2_noUKB"
                               )
                    )
rgs %>% 
      filter(p2 != "EA3_excl_23andMe_UK", p2 != "EA2_noUKB") %>% 
      ggplot(aes(rg, estimate)) + 
        geom_point() +
        my_vline +
        my_hline +
        labs(x = "Genetic correlation with EA3", y = "Effect on RLRS")

```

Another way to examine the "earnings" theory of natural selection is to compare
selection effects of polygenic scores with their genetic correlation with educational
attainment (EA3). Since EA3 strongly predicts earnings, if earnings drives
differences in fertility, we'd expect a correlation between the two sets of
results. Figure \@ref(fig:plot-rgs-by-effect-size) shows this is so: the 
correlation, after excluding EA2, is `r rgs_cor$estimate`. Genetic
correlations were calculated using LD score regression from GWAS summary
statistics.

\clearpage
\FloatBarrier

## Model proofs

```{r include-model-appendix, child="model-appendix.Rmd"}

```

\clearpage
\FloatBarrier

# References